### 2022-11-09 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@LouisDesdoigts}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2022-11-10 src/scripts/gen_model_data.py
Adding optics script

@@ -1,500 +1,0 @@
-# Core jax
-import jax
-import jax.numpy as np
-import jax.random as jr
-
-# Optimisation
-import equinox as eqx
-import optax
-
-# Optics
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
-from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
-import paths
-
-# Pickle
-import pickle as p
-
-# Plotting/visualisation
-import matplotlib.pyplot as plt
-
-plt.rcParams['image.cmap'] = 'inferno'
-plt.rcParams["font.family"] = "serif"
-plt.rcParams["image.origin"] = 'lower'
-plt.rcParams['figure.dpi'] = 120
-
-
-
-
-'''Create Optics, Detector, Source & Instrument'''
-# Define wavelengths
-wavels = 1e-9 * np.linspace(545, 645, 3)
-
-# Basic Optical Parameters
-aperture = 0.5
-wf_npix = 512
-
-# Detector Parameters
-det_npix = 1024
-sampling_rate = 20
-det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
-
-# Load mask
-raw_mask = dl.utils.phase_to_opd(np.load("mask.npy"), wavels.mean())
-mask = dl.utils.scale_array(raw_mask, wf_npix, 0)
-
-# Zernike Basis
-zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
-coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), [len(zern_basis)])
-
-# Define Optical Configuration
-optical_layers = [
-    dl.CreateWavefront    (wf_npix, aperture),
-    dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
-    dl.ApplyBasisOPD      (zern_basis, coeffs),
-    dl.AddOPD             (mask),
-    dl.NormaliseWavefront (),
-    dl.AngularMFT         (det_npix, det_pixsize)]
-
-# Create Optics object
-optics = dl.Optics(optical_layers)
-
-# Pixel response
-pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(0), [det_npix, det_npix])
-
-# Create Detector object
-detector = dl.Detector([dl.ApplyPixelResponse(pix_response)])
-
-# Observation stratergy, define dithers
[…]

### 2022-11-10 src/scripts/gen_optics.py

@@ -1,0 +1,138 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+import pickle as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+
+
+'''Create Optics, Detector, Source & Instrument'''
+# Define wavelengths
+wavels = 1e-9 * np.linspace(545, 645, 3)
+
+# Basic Optical Parameters
+aperture = 0.5
+wf_npix = 512
+
+# Detector Parameters
+det_npix = 1024
+sampling_rate = 20
+det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
+
+# Load mask
+raw_mask = dl.utils.phase_to_opd(np.load("mask.npy"), wavels.mean())
+mask = dl.utils.scale_array(raw_mask, wf_npix, 0)
+
+# Zernike Basis
+zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
+coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), [len(zern_basis)])
+
+# Define Optical Configuration
+optical_layers = [
+    dl.CreateWavefront    (wf_npix, aperture),
+    dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
+    dl.ApplyBasisOPD      (zern_basis, coeffs),
+    dl.AddOPD             (mask),
+    dl.NormaliseWavefront (),
+    dl.AngularMFT         (det_npix, det_pixsize)]
+
+# Create Optics object
+optics = dl.Optics(optical_layers)
+
+# Pixel response
+pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(0), [det_npix, det_npix])
+
+# Create Detector object
+detector = dl.Detector([dl.ApplyPixelResponse(pix_response)])
+
+# Observation stratergy, define dithers
[…]

### 2022-11-10 src/scripts/gen_data.py
Adding the data script

@@ -1,0 +1,440 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+import pickle as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+tel = p.load(open(paths.data / "instrument.p"))
+
+# Observe!
+psfs = tel.observe()
+
+# Observe!
+psfs = tel.observe()
+
+'''Figure 1'''
+from matplotlib.cm import get_cmap
+opd = tel.ApplyBasisOPD.get_total_opd()
+psf = tel.model(source=dl.PointSource(wavelengths=wavels))
+throughput = tel.CompoundAperture.get_aperture(npixels=wf_npix)
+pupil = opd.at[np.where(throughput==0.)].set(np.nan)
+
+FF = tel.ApplyPixelResponse.pixel_response
+
+cmap = get_cmap("inferno")
+cmap.set_bad('k',1.)
+
+plt.figure(figsize=(15, 4))
+plt.subplot(1, 3, 1)
+plt.title("Pupil OPD")
+plt.imshow(pupil * 1e9, cmap=cmap)
+plt.xticks([])
+plt.yticks([])
+cbar = plt.colorbar()
+cbar.set_label("OPD (nm)")
+
+plt.subplot(1, 3, 2)
+plt.title("PSF")
+plt.imshow(psf)
+plt.xlabel("Pixels")
+plt.ylabel("Pixels")
+cbar = plt.colorbar()
+cbar.set_label("Probability")
+
+plt.subplot(1, 3, 3)
+plt.title("Pixel Response")
+plt.hist(FF.flatten(), bins=25)
+plt.ylabel("Counts")
+plt.xlabel("Relative Sensitivity")
+
+plt.tight_layout()
+plt.savefig(paths.figures / "optics.pdf", dpi=300)
+plt.show()
+
+
+
+
+
+'''Figure 2'''
+# Apply some noise to the PSF Background noise
[…]

### 2022-11-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[]{spie}
+\usepackage[utf8]{inputenc}
 \usepackage{showyourwork}
+\usepackage{xspace}
+%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
+%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
+
+\renewcommand{\baselinestretch}{1.0} % Change to 1.65 for double spacing
+\usepackage{amsmath,amsfonts,amssymb}
+\usepackage{graphicx}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+
+\usepackage{relsize}
+
+\usepackage{acro}
+\acsetup{
+  single    = false,
+  list/sort  = true,
+  cite/group = true,
+  cite/group/cmd = \cite,
+  cite/group/pre = {},
+}
+
+\DeclareAcronym{prf}{
+  short = PRF,
+  long = Pixel Response Function
+}
+
+\DeclareAcronym{psf}{
+  short = PSF,
+  long = Point Spread Function
+}
+
+\DeclareAcronym{gpu}{
+  short = GPU,
+  long = Graphics Processing Unit
+}
+
+\DeclareAcronym{hmc}{
+  short = HMC,
+  long = Hamiltonian Monte Carlo,
+  cite = {Betancourt2017}
+}
+
+\DeclareAcronym{mcmc}{
+  short = MCMC,
+  long = Markov Chain Monte Carlo,
+  cite = {Metropolis1953}
+}
+
+\DeclareAcronym{gp}{
+  short = GP,
+  long = Gaussian Process,
+  cite = {rasmussen:williams:2006}
+}
+
+\DeclareAcronym{snr}{
+  short = SNR,
+  long = signal-to-noise ratio
+}
+
+\newcommand\jax{\textsc{Jax}\xspace}
+\newcommand\dlux{$\partial$Lux\xspace}
+
+\title{Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase with Automatic Differentiation}
+
+\author[a]{Louis C. Desdoigts}
+\author[b]{Benjamin J. S. Pope}
+\author[a]{Peter G. Tuthill}
+
+\affil[a]{Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW~2006, Australia}
+\affil[b]{School of Mathematics and Physics, University of Queensland, St Lucia, QLD~4072, Australia}
+\affil[c]{Centre for Astrophysics, University of Southern Queensland, West Street, Toowoomba, QLD~4350, Australia}
 
 % Begin!
+\authorinfo{Further author information: (Send correspondence to L. C. D.)\\L. C. D.: E-mail: louis.desdoigts@sydney.edu.au}
+
+% Option to view page numbers
+\pagestyle{empty} % change to \pagestyle{plain} for page numbers   
+\setcounter{page}{301} % Set start page numbering at e.g. 301
+
 \begin{document} 
+\maketitle
 
 % Title
-\title{An open source scientific article}
 
 % Author list
-\author{@LouisDesdoigts}
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+Stuff about \dlux
+
 \end{abstract}
 
-% Main body with filler text
 \section{Introduction}
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+% High precision astronomy is limited by systematics introduced by the optics and detector
+At the bleeding edge of space-based astronomical imaging, photometry, and spectroscopy, the precision is limited by systematics introduced by aberrations in the optics and noise processes in the detector.
+While problems are ubiquitous in astronomy, we are motivated by several core examples.
+In exoplanet direct imaging, we want to achieve high resolution and high contrast simultaneously with instruments like JWST Coronagraphy \cite{Girard2022}, Aperture Masking \cite{Sivaramakrishnan2022}, or Kernel Phase \cite{Kammerer2022} modes.
+The astrometric mission Toliman \cite{tuthill2018} aims to measure precise relative positions of the binary stars $\alpha$~Centauri~AB to reveal the gravitational influence of unseen planets; this will require micro-arcsecond, micro-pixel astrometric precision.
+We may also want to perform high precision photometry, whether with dedicated missions like \textit{Kepler} \cite{Borucki2010} and TESS \cite{Ricker2015}, or as ancillary science with Toliman.
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+In each case, we face serious limitations from an imperfect knowledge of the \ac{prf}, the map of the intra-pixel and inter-pixel variations in detector sensitivity; and the \ac{psf}, the diffraction-limited pattern by which light from a point source like a star is spatially spread out across a detector.
 
+% PSFs
+The \ac{psf} depends on the successive planes through which light passes from the entrance pupil of the telescope through to the detector.
+This can be calculated from physical optics: for a simple camera bringing light from the pupil to focus at a detector plane, the \ac{psf} is the Fourier transform of the input wavefront, the regime of Fraunhofer diffraction.
+In the more general Fresnel regime where the detector is slightly out of focus, there are additional quadratic phase factors before and after this Fourier transform.
+In more complicated instruments like coronagraphs, the light might pass through multiple re-imaged pupil and focal planes, being operated on in each.
+In either case, the trouble is that the \ac{psf} is distorted by unknown aberrations - distortions in the wavefront, which can be represented as a spatially-varying phase map across the optical plane.
+
+% Phase Retrieval - G-S, Fienup, review papers, kernel phase
+
+% Pixel Sensitivities - Photometry
+% K2
+% K2SFF
+% K2SC
+% EVEREST
+% OWL/HALO
+% CPM
+% LSST
+% Pixel Sensitivities - Imaging
+% Euclid
+% JWST AMI
+
+In this series of papers, we present a new software package, \dlux, for fitting high-dimensional parametrized physical models to astronomical data.
+By using the Python library \textsc{Jax}\cite{jax}, we obtain hardware acceleration on the \ac{gpu}, as well as automatic differentiation that enables high dimensional optimization and inference with gradient descent or \ac{hmc}.
+In the present paper, we focus on the particular problem of estimating the \ac{prf} and \ac{psf} simultaneously - i.e.
+joint flat field calibration and phase retrieval.
+
+\section{\dlux Differentiable Optical Models}
+\label{sec:dlux}
+
+% Physical Optics Forwards Models
+% poppy/WebbPSF
+% Autodiff
+% Review backprop/NNs
+% Jax, PyTorch, Julia
+% Autodiff Optics
+% WaveBlocks
+% Sitzmann etc
+% Liaudat
+% Morphine
+% Overview of Wong et al 2021
+% dLux
+% Equinox
+% Features
+% Example of how to set up a model from layers
+
+% Figure 1: (Hand-drawn?) Illustration of the Optical System & ∂Lux model layers
+
+\section{Phase Retrieval and Flat Field Estimation}
+\label{sec:phaseretrieval}
+% Describe simulation setup
+% Figure 2: OPD & Pupil in our test, PSF, Flat Field histogram
+
+% Describe gradient descent implementation & scheduling
+% Takes X time on Y machine with Z precision on params (numbers from ShowYourWork)
+
+% Figure showing results
+
+% Figure 3: Like final 2l figures of tutorial merged together: 
+% Top: loss function vs epoch
+% Middle: correlation plot of Zernikes (with residual below); PSF (true, recovered, residual)
+% Bottom: correlation plot of star fluxes (with residual below), correlation plot of sensitivities (with residual below), 2d residual map of pixels
+
+\section{Discussion}
+% Can this break down?
+% Nyquist?
+% Low photon SNR limit?
+% Dithering or N stars requirement
+
+\section{Conclusions \& Future Work}
+% Review outcomes
+% Useful for space photometry in crowded fields
+% Toliman Pleiades?
+% Useful for detector calibration & monitoring
+% JWST, Roman
+% Time series: regularize with GP?
+
+% \begin{figure}[ht!]
+%     \script{gen_optics.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/optics.pdf}
+%         \caption{
+%             The optics
+%         }
+%         \label{fig:optics}
+%     \end{centering}
+% \end{figure}
+
+\section{Acknowledgements}
+
+We acknowledge and pay respect to the Gadigal people of the Eora Nation, upon whose unceded, sovereign, ancestral lands the University of Sydney is built; and the traditional owners of the land on which the University of Queensland is situated, the Turrbal and Jagera people.
+We pay respects to their Ancestors and descendants, who continue cultural and spiritual connections to Country.
+
+This research made use of \textsc{Numpy} \cite{numpy}; SciPy \cite{scipy}; Matplotlib \cite{matplotlib}; Astropy \cite{astropy:2013,astropy:2018}; \textsc{Jax} \cite{jax}; and \texttt{equinox} \cite{kidger2021equinox}.
+
+\bibliographystyle{spiebib} % makes bibtex use spiebib.bst
 \bibliography{bib}

### 2022-11-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\subsection{Phase Retrieval}
 % PSFs
@@
 % Phase Retrieval - G-S, Fienup, review papers, kernel phase
+Phase retrieval is the problem of inferring these aberrations from data, which is in general ill-posed: because of the Hermitian symmetry of the Fourier transform, and because we measure \textit{intensity} and not electric field in optical astronomy, there is a large space of aberrations that would generate the same intensity \ac{psf}.
+Fortunately, this space can be restricted to physically-realistic solutions and readily solved by algorithms such as the Gerchberg-Saxton algorithm \cite{gerchberg1972}, using ideas from compressed sensing \cite{candes2011}, or by machine learning \cite{metzler2018,isil2019,nishizaki2020}.
+Phase retrieval was memorably performed to infer and correct the serious aberration on the \textit{Hubble Space Telescope} mirror at launch \cite{hubble_phase_ret}.
+Earlier work in the vein of this paper has shown that the phase retrieval problem can be efficiently solved by taking a forwards model of physical optics, obtaining partial derivatives with automatic differentiation, and optimizing its parameters by gradient descent \cite{jurling_fienup,phase_ret_and_design}.
 
+\subsection{Detector Calibration}
 % Pixel Sensitivities - Photometry

### 2022-11-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+In this series of papers, we present a new software package, \dlux, for fitting high-dimensional parametrized physical optics models to astronomical data.
+By using the Python library \textsc{Jax}\cite{jax}, we obtain \ac{gpu} hardware acceleration, as well as automatic differentiation\cite{Margossian2018} features that enable high dimensional optimization and inference with gradient descent or \ac{hmc}.
+In the present paper, we focus on the particular problem of estimating the \ac{prf} and \ac{psf} simultaneously - i.e.
+joint flat field calibration and phase retrieval.
+In Paper II, we will show how automatic differentiation enables improvements to hardware \textit{design}, directly calculating and optimizing the Fisher information and therefore fundamental figures of merit of an optical system end-to-end.
+
 \subsection{Phase Retrieval}
@@
 In the more general Fresnel regime where the detector is slightly out of focus, there are additional quadratic phase factors before and after this Fourier transform.
-In more complicated instruments like coronagraphs, the light might pass through multiple re-imaged pupil and focal planes, being operated on in each.
+In more complicated instruments like coronagraphs \cite{Bowler2016}, the light might pass through multiple re-imaged pupil and focal planes, being operated on in each.
 In either case, the trouble is that the \ac{psf} is distorted by unknown aberrations - distortions in the wavefront, which can be represented as a spatially-varying phase map across the optical plane.
@@
 % Phase Retrieval - G-S, Fienup, review papers, kernel phase
-Phase retrieval is the problem of inferring these aberrations from data, which is in general ill-posed: because of the Hermitian symmetry of the Fourier transform, and because we measure \textit{intensity} and not electric field in optical astronomy, there is a large space of aberrations that would generate the same intensity \ac{psf}.
+Phase retrieval is the problem of inferring these aberrations from data\cite{schechtman2014}, which is in general ill-posed \cite{barnett2020}: because of the Hermitian symmetry of the Fourier transform, and because we measure \textit{intensity} and not electric field in optical astronomy, there is a large space of aberrations that would generate the same intensity \ac{psf}.
 Fortunately, this space can be restricted to physically-realistic solutions and readily solved by algorithms such as the Gerchberg-Saxton algorithm \cite{gerchberg1972}, using ideas from compressed sensing \cite{candes2011}, or by machine learning \cite{metzler2018,isil2019,nishizaki2020}.
@@
 % JWST AMI
-
-In this series of papers, we present a new software package, \dlux, for fitting high-dimensional parametrized physical models to astronomical data.
-By using the Python library \textsc{Jax}\cite{jax}, we obtain hardware acceleration on the \ac{gpu}, as well as automatic differentiation that enables high dimensional optimization and inference with gradient descent or \ac{hmc}.
-In the present paper, we focus on the particular problem of estimating the \ac{prf} and \ac{psf} simultaneously - i.e.
-joint flat field calibration and phase retrieval.
 

### 2022-11-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[]{spie}
+\documentclass[twocolumn]{spie}
+% \documentclass[twocolumn]{aastex631}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{graphicx}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-
-\usepackage{relsize}
+% \usepackage[colorlinks=true, allcolors=blue]{hyperref}
 
@@
 In Paper II, we will show how automatic differentiation enables improvements to hardware \textit{design}, directly calculating and optimizing the Fisher information and therefore fundamental figures of merit of an optical system end-to-end.
+We make \dlux available as open-source software on GitHub\footnote{\href{https://github.com/LouisDesdoigts/dLux}{github.com/LouisDesdoigts/dLux}}, and encourage interested readers to use and contribute to this package as a community resource.
+This paper has been compiled using ShowYourWork \cite{Luger2021}, so that all figures link to the code used to produce them.
 
@@
 \label{sec:dlux}
-
+\textcolor{red}{Louis - this is your bit to write!}
 % Physical Optics Forwards Models

### 2022-11-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{graphicx}
-% \usepackage[colorlinks=true, allcolors=blue]{hyperref}
-
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage{relsize}
 \usepackage{acro}
@@
 
+\affil[a]{Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW~2006, Australia}
+\affil[b]{School of Mathematics and Physics, University of Queensland, St Lucia, QLD~4072, Australia}
+\affil[c]{Centre for Astrophysics, University of Southern Queensland, West Street, Toowoomba, QLD~4350, Australia}
+
 \author[a]{Louis C. Desdoigts}
@@
 
-\affil[a]{Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW~2006, Australia}
-\affil[b]{School of Mathematics and Physics, University of Queensland, St Lucia, QLD~4072, Australia}
-\affil[c]{Centre for Astrophysics, University of Southern Queensland, West Street, Toowoomba, QLD~4350, Australia}
-
 % Begin!
-\authorinfo{Further author information: (Send correspondence to L. C. D.)\\L. C. D.: E-mail: louis.desdoigts@sydney.edu.au}
+\authorinfo{Send correspondence to L. C. D.: \\E-mail: \href{mailto:louis.desdoigts@sydney.edu.au}{louis.desdoigts@sydney.edu.au}}
 
@@
 \pagestyle{empty} % change to \pagestyle{plain} for page numbers   
-\setcounter{page}{301} % Set start page numbering at e.g. 301
+% \setcounter{page}{301} % Set start page numbering at e.g. 301
 
@@
 \subsection{Detector Calibration}
+
+High-precision, high-cadence time series of space photometry from \textit{Kepler} \cite{Borucki2010}, K2 \cite{K2}, TESS \cite{TESS} and CHEOPS \cite{CHEOPS} space telescopes have been revolutionary for exoplanetary science and stellar asr.
+This increased precision over that last 20 years has also revolutionized asteroseismology [cite multiple papers here], the study of stellar interiors.
+
+% Photometry is hard because of flat field & optical aberrations
+In many practical cases, photometric precision is limited by systematic errors due to variations in sensitivity within and between pixels.
+Changes in telescope pointing or PSF couple to these inter- and intra-pixel variations in the flat field.
+This was particularly severe for the K2 mission where the breakdown of two reaction wheels in the original Kepler mission necessitated the telescope pointing be maintained with periodic thruster firings.
+This pointing instability introduced a saw-tooth systematic on 6 hour timescales, comparable to the timescales of exoplanet transits and asteroseismic signals of interest.
+Similarly, throughout one Earth orbit telescopes experience a changing thermal gradient across the optical train as one side faces the Sun, and the other deep space.
+Depending on the orbital height, this introduces periodic optical aberrations that distort the PSF over timescales of interest.
+The systematic coupling of these effects has given rise to many methods used to either de-correlate these signals or calibrate the systematics.
+
+%  flat field calibration
+Much work has been put towards developing data-driven self-calibration of the flat field in K2, which are of general interest and applicability for other space photometry missions. 
+% 		- self-flat fielding (Andrew Vanderburg) 
+\citet{self_flat-fielding} developed the `self flat-fielding' approach that harnesses the correlation between causally-disconnected sources in order to isolate the noise processes of interest.
+First the thruster firings and pointing drift is calibrated by calculating the correlation between all stellar sources on the detector.
+Time-series photomtetric signals that correlate with source position are then used to isolate the signal introduced by pixel sensitivity variation and correct for it.
+%       - dun wang causal pixel model 
+Taking a similar approach \citet{causal_pixel} constrains instrumental effects by measuring correlations in the light curves of causally uncorrelated pixels.
+% 		- Christina Hedges (pixel cal papers) ??
+% 		- Luger papers (pixel-level decorrelation)
+In \citet{pixel_decorr} ... 
+% 		- halo photometry (white, pope)
+Other non-linear instrumental effects such as saturation can further complicate the recovery of precise light curves.
+Halo photometry \citep{halo, halo2} sidesteps the saturation problem by extracting light curves from the `halo' of scattered light around the saturated pixels.
+Instrumental and astrophysical effects are separated by assigning individual pixel weights and biases, which are optimised in order to maximise correlation the between pixel light-curves within the same halo, and minimise correlations between pixels light curves across different halos.
+
 % Pixel Sensitivities - Photometry

### 2022-11-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{relsize}
+% \usepackage{relsize}
 \usepackage{acro}
@@
 
-High-precision, high-cadence time series of space photometry from \textit{Kepler} \cite{Borucki2010}, K2 \cite{K2}, TESS \cite{TESS} and CHEOPS \cite{CHEOPS} space telescopes have been revolutionary for exoplanetary science and stellar asr.
-This increased precision over that last 20 years has also revolutionized asteroseismology [cite multiple papers here], the study of stellar interiors.
+High-precision, high-cadence time series of space photometry from \textit{Kepler} \cite{Borucki2010}, K2 \cite{K2}, TESS \cite{TESS} and CHEOPS \cite{CHEOPS} space telescopes have been revolutionary for exoplanetary science and stellar astrophysics.
 

### 2022-11-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
-% \usepackage{relsize}
+\usepackage{hyperref}
+\usepackage{relsize}
 \usepackage{acro}
@@
 
-\newcommand\jax{\textsc{Jax}\xspace}
+% \newcommand\jax{\textsc{Jax}\xspace}
 \newcommand\dlux{$\partial$Lux\xspace}
 
-\title{Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase with Automatic Differentiation}
+% \title{Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase with Automatic Differentiation}
+\title{Differentiable Optics with dLux: I - Deep Calibration of Flat Field and Phase with Automatic Differentiation}
 
@@
 
-High-precision, high-cadence time series of space photometry from \textit{Kepler} \cite{Borucki2010}, K2 \cite{K2}, TESS \cite{TESS} and CHEOPS \cite{CHEOPS} space telescopes have been revolutionary for exoplanetary science and stellar astrophysics.
+High-precision, high-cadence time series of space photometry from \textit{Kepler} \cite{Borucki2010}, K2 \cite{K2}, TESS \cite{TESS} and CHEOPS \cite{CHEOPS} space telescopes have been revolutionary for exoplanetary science \cite{Zhu2021} and stellar astrophysics \cite{Aerts21,Jackiewicz2021}.
 
@@
 % 		- self-flat fielding (Andrew Vanderburg) 
-\citet{self_flat-fielding} developed the `self flat-fielding' approach that harnesses the correlation between causally-disconnected sources in order to isolate the noise processes of interest.
-First the thruster firings and pointing drift is calibrated by calculating the correlation between all stellar sources on the detector.
-Time-series photomtetric signals that correlate with source position are then used to isolate the signal introduced by pixel sensitivity variation and correct for it.
+In  `self flat-fielding' (K2SFF\cite{self_flat-fielding}), the light curve is decorrelated against principal components of all light curves on the detector.
+Time-series photometric signals that correlate with source position are then used to isolate the signal introduced by pixel sensitivity variation and correct for it.
 %       - dun wang causal pixel model 

### 2022-11-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{graphicx}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+% \usepackage[colorlinks=true, allcolors=blue]{hyperref}
 \usepackage{hyperref}
@@
 % 		- self-flat fielding (Andrew Vanderburg) 
-In  `self flat-fielding' (K2SFF\cite{self_flat-fielding}), the light curve is decorrelated against principal components of all light curves on the detector.
-Time-series photometric signals that correlate with source position are then used to isolate the signal introduced by pixel sensitivity variation and correct for it.
+In  `self flat-fielding' (K2SFF\cite{self_flat-fielding}), the systematics are modelled with a spline regression of raw flux versus detector position, together with a linear combination of principal components of all light curves on the detector.
 %       - dun wang causal pixel model 
-Taking a similar approach \citet{causal_pixel} constrains instrumental effects by measuring correlations in the light curves of causally uncorrelated pixels.
+The `Causal Pixel Model` \cite{causal_pixel} constrains instrumental effects by measuring correlations in the light curves of causally uncorrelated pixels.
 % 		- Christina Hedges (pixel cal papers) ??
 % 		- Luger papers (pixel-level decorrelation)
-In \citet{pixel_decorr} ... 
+In \cite{pixel_decorr} ... 
 % 		- halo photometry (white, pope)
 Other non-linear instrumental effects such as saturation can further complicate the recovery of precise light curves.
-Halo photometry \citep{halo, halo2} sidesteps the saturation problem by extracting light curves from the `halo' of scattered light around the saturated pixels.
+Halo photometry \cite{halo, halo2} sidesteps the saturation problem by extracting light curves from the `halo' of scattered light around the saturated pixels.
 Instrumental and astrophysical effects are separated by assigning individual pixel weights and biases, which are optimised in order to maximise correlation the between pixel light-curves within the same halo, and minimise correlations between pixels light curves across different halos.

### 2022-11-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\affil[a]{Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW~2006, Australia}
-\affil[b]{School of Mathematics and Physics, University of Queensland, St Lucia, QLD~4072, Australia}
-\affil[c]{Centre for Astrophysics, University of Southern Queensland, West Street, Toowoomba, QLD~4350, Australia}
+% \affil[a]{Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW~2006, Australia}
+% \affil[b]{School of Mathematics and Physics, University of Queensland, St Lucia, QLD~4072, Australia}
+% \affil[c]{Centre for Astrophysics, University of Southern Queensland, West Street, Toowoomba, QLD~4350, Australia}
 

### 2022-11-17 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2022-11-10 src/tex/ms.tex
@@
 \documentclass[twocolumn]{spie}
+% \documentclass[]{spie}
 % \documentclass[twocolumn]{aastex631}
@@
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{hyperref}
+% \usepackage{hyperref}
 \usepackage{relsize}
@@
   long = Gaussian Process,
-  cite = {rasmussen:williams:2006}
+  cite = {Rasmussen2005,Aigrain2022}
 }
@@
 
-% \newcommand\jax{\textsc{Jax}\xspace}
+\DeclareAcronym{pld}{
+  short = PLD,
+  long = Pixel Level Decorrelation,
+  cite = {everest}
+}
+
+\DeclareAcronym{k2sff}{
+  short = SFF,
+  long = Self Flat-Fielding,
+  cite = {self_flat-fielding}
+}
+
+\DeclareAcronym{cpm}{
+  short = CPM,
+  long = Causal Pixel Model,
+  cite = {causal_pixel}
+}
+
+\DeclareAcronym{api}{
+  short = API,
+  long = Application Programming Interface
+}
+
+\DeclareAcronym{jit}{
+  short = jit,
+  long = Just-In-Time
+}
+
+\DeclareAcronym{opd}{
+  short = OPD,
+  long = Optical Path Difference
+}
+
+\newcommand\jax{\textsc{Jax}\xspace}
 \newcommand\dlux{$\partial$Lux\xspace}
+\newcommand\equinox{\texttt{equinox}\xspace}
+\newcommand\optax{\texttt{optax}\xspace}
+\newcommand\numpyro{\texttt{numpyro}\xspace}
 
-% \title{Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase with Automatic Differentiation}
-\title{Differentiable Optics with dLux: I - Deep Calibration of Flat Field and Phase with Automatic Differentiation}
+\newcommand\kepler{\textit{Kepler}\xspace}
+
+\title{Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase with autodiff}
+% \title{Differentiable Optics with dLux: I - Deep Calibration of Flat Field and Phase with autodiff}
 
@@
 % Option to view page numbers
-\pagestyle{empty} % change to \pagestyle{plain} for page numbers   
+\pagestyle{plain} % change to \pagestyle{empty} for no page numbers   
 % \setcounter{page}{301} % Set start page numbering at e.g. 301
@@
 \begin{abstract}
-Stuff about \dlux
+
+$\partial$Lux is a newly developed optical modelling framework deigned to harness the powerful mathematical tools underpinning machine learning and directly apply them to the field of optics. 
+Both neural networks and optical systems map input vectors to output vectors employing a series of intermediary linear transformations and non-linear matrix operations. 
+This isomorphism allows for optical models to be directly constructed within existing automatic differentiation (autodiff) libraries. 
+$\partial$Lux exploits this relationship harnessing autodiff libraries to create a naively end-to-end fully differentiable optical modelling framework. 
+This may precipitate a paradigm shift in the power and utility of these optical models, opening the possibility to entirely novel algorithms and approaches. 
+Astronomical optics is one field where notoriously challenging or intractable problems can be re-approached with new and simple techniques. 
+This manuscript explores how efficient high-dimensional gradient calculations enabled by autodiff can be harnessed using differentiable optical models in order to simultaneously recover uncalibrated astronomical, optical and instrumental parameters.
 
@@
 While problems are ubiquitous in astronomy, we are motivated by several core examples.
-In exoplanet direct imaging, we want to achieve high resolution and high contrast simultaneously with instruments like JWST Coronagraphy \cite{Girard2022}, Aperture Masking \cite{Sivaramakrishnan2022}, or Kernel Phase \cite{Kammerer2022} modes.
+In exoplanet direct imaging, we want to achieve high resolution and high contrast simultaneously with instruments like JWST Coronagraphy \cite{Girard2022}, Aperture Masking \cite{Sivaramakrishnan2022}, or Kernel Phase \cite{Martinache2010,Kammerer2022} modes.
 The astrometric mission Toliman \cite{tuthill2018} aims to measure precise relative positions of the binary stars $\alpha$~Centauri~AB to reveal the gravitational influence of unseen planets; this will require micro-arcsecond, micro-pixel astrometric precision.
@@
 In this series of papers, we present a new software package, \dlux, for fitting high-dimensional parametrized physical optics models to astronomical data.
-By using the Python library \textsc{Jax}\cite{jax}, we obtain \ac{gpu} hardware acceleration, as well as automatic differentiation\cite{Margossian2018} features that enable high dimensional optimization and inference with gradient descent or \ac{hmc}.
+By using the Python library \textsc{Jax}\cite{jax}, we obtain \ac{gpu} hardware acceleration, as well as autodiff\cite{Margossian2018} features that enable high-dimensional optimization and inference with gradient descent or \ac{hmc}.
 In the present paper, we focus on the particular problem of estimating the \ac{prf} and \ac{psf} simultaneously - i.e.
 joint flat field calibration and phase retrieval.
-In Paper II, we will show how automatic differentiation enables improvements to hardware \textit{design}, directly calculating and optimizing the Fisher information and therefore fundamental figures of merit of an optical system end-to-end.
+In Paper II, we will show how autodiff enables improvements to hardware \textit{design}, directly calculating and optimizing the Fisher information and therefore fundamental figures of merit of an optical system end-to-end.
 We make \dlux available as open-source software on GitHub\footnote{\href{https://github.com/LouisDesdoigts/dLux}{github.com/LouisDesdoigts/dLux}}, and encourage interested readers to use and contribute to this package as a community resource.
@@
 Phase retrieval was memorably performed to infer and correct the serious aberration on the \textit{Hubble Space Telescope} mirror at launch \cite{hubble_phase_ret}.
-Earlier work in the vein of this paper has shown that the phase retrieval problem can be efficiently solved by taking a forwards model of physical optics, obtaining partial derivatives with automatic differentiation, and optimizing its parameters by gradient descent \cite{jurling_fienup,phase_ret_and_design}.
+Earlier work in the vein of this paper has shown that the phase retrieval problem can be efficiently solved by taking a forwards model of physical optics, obtaining partial derivatives with autodiff, and optimizing its parameters by gradient descent \cite{jurling_fienup,phase_ret_and_design}.
 
@@
 In many practical cases, photometric precision is limited by systematic errors due to variations in sensitivity within and between pixels.
-Changes in telescope pointing or PSF couple to these inter- and intra-pixel variations in the flat field.
-This was particularly severe for the K2 mission where the breakdown of two reaction wheels in the original Kepler mission necessitated the telescope pointing be maintained with periodic thruster firings.
-This pointing instability introduced a saw-tooth systematic on 6 hour timescales, comparable to the timescales of exoplanet transits and asteroseismic signals of interest.
-Similarly, throughout one Earth orbit telescopes experience a changing thermal gradient across the optical train as one side faces the Sun, and the other deep space.
-Depending on the orbital height, this introduces periodic optical aberrations that distort the PSF over timescales of interest.
-The systematic coupling of these effects has given rise to many methods used to either de-correlate these signals or calibrate the systematics.
+Changes in telescope pointing or variations in the \ac{psf} with focus couple to these inter- and intra-pixel variations in the flat field to produce changes in overall measured flux.
+This was particularly severe for the K2 mission where periodic thruster firings introduced a saw-tooth systematic on 6 hour timescales, comparable to the timescales of exoplanet transits and asteroseismic signals of interest.
 
@@
 % 		- self-flat fielding (Andrew Vanderburg) 
-In  `self flat-fielding' (K2SFF\cite{self_flat-fielding}), the light curve is decorrelated against principal components of all light curves on the detector.
-Time-series photometric signals that correlate with source position are then used to isolate the signal introduced by pixel sensitivity variation and correct for it.
+In \ac{k2sff}, the systematics are modelled with a spline regression of raw flux versus detector position, together with a linear combination of principal components of all light curves on the detector, and \textsc{k2sc} \cite{Aigrain2016} uses a \ac{gp} similarly.
 %       - dun wang causal pixel model 
-Taking a similar approach \citet{causal_pixel} constrains instrumental effects by measuring correlations in the light curves of causally uncorrelated pixels.
+The \ac{cpm} constrains instrumental effects by measuring correlations in the light curves of spatially distant, causally-disconnected pixels.
 % 		- Christina Hedges (pixel cal papers) ??
 % 		- Luger papers (pixel-level decorrelation)
-In \citet{pixel_decorr} ... 
+In \ac{pld} the light curve is detrended by a linear combination of regressors formed by the ensemble of normalized pixel time series contributing to the light curve, together with their higher-order products.
 % 		- halo photometry (white, pope)
-Other non-linear instrumental effects such as saturation can further complicate the recovery of precise light curves.
-Halo photometry \citep{halo, halo2} sidesteps the saturation problem by extracting light curves from the `halo' of scattered light around the saturated pixels.
-Instrumental and astrophysical effects are separated by assigning individual pixel weights and biases, which are optimised in order to maximise correlation the between pixel light-curves within the same halo, and minimise correlations between pixels light curves across different halos.
+Halo photometry \cite{halo, halo2} extracts light curves from a weighted sum of pixels where the weights are learned by minimizing the total variation of the resulting light curve, a convex optimization which is tractable because of autodiff.
 
-% Pixel Sensitivities - Photometry
-% K2
-% K2SFF
-% K2SC
-% EVEREST
-% OWL/HALO
-% CPM
-% LSST
 % Pixel Sensitivities - Imaging
-% Euclid
-% JWST AMI
+Accurate measurement of background noise is also a serious issue for imaging science, where for example bad pixels, background noise, and uncertainties in the detector systematics limit sensitivity of aperture masking interferometry and kernel phase \cite{Kammerer2019}, and for low-surface-brightness science such as with \textit{Euclid}\cite{EuclidCollaboration2022}.
+In each case it is normal for detectors to be rigorously calibrated on the ground before flight, and with dedicated observations in space; but at the levels of contrast and resolution required for aperture masking on JWST (ideally $10^{-4}$ at a few $\lambda/D$) even small residual miscalibrations can seriously affect performance \cite{Sivaramakrishnan2022} and we require some method of self-calibration from science data themselves.
 
@@
 \label{sec:dlux}
-\textcolor{red}{Louis - this is your bit to write!}
+
 % Physical Optics Forwards Models
+Rather than working with reduced data products like visibilities and light curves, an alternative is to directly model the pixel-level images with a parametrized model of the optics and detector. 
 % poppy/WebbPSF
+For example, the popular package \textsc{poppy} \cite{poppy} can accurately simulate Fresnel and Fraunhofer propagation, with a WebbPSF extension for preset JWST instrument models \cite{Perrin2014}. 
 % Autodiff
+In principle, it is possible to fit such a model to data with \ac{mcmc}, fitting the positions of stars and a modal basis representation of aberrations; but common samplers can be very slow, or fail to converge in high dimensions \cite{Huijser2022}.
+This restraint is lifted by using \ac{hmc}, which requires computation of the gradient of the log-likelihood with respect to parameters.
+
 % Review backprop/NNs
-% Jax, PyTorch, Julia
+The core technology that enables this for physics also underpins the last decade's revolution in deep learning \cite{lecun15}: autodiff.
+By application of the chain rule, it is possible to calculate the partial derivatives of almost arbitrary numerical functions with respect to their floating-point arguments, so long as they are executed in an appropriate software framework.
+% Theano, TensorFlow, PyTorch, Julia, Jax
+Considerable private-sector and open-source investment has gone into building autodiff frameworks to enable machine learning, such as Theano \cite{theano}, TensorFlow \cite{tensorflow2015}, PyTorch \cite{pytorch}, native support in the Julia language \cite{julia}, and the framework we apply in this paper, \textsc{Jax} \cite{jax}.
+\jax has an \ac{api} mirroring the common Python array library \textsc{NumPy} \cite{numpy}, while supporting \ac{gpu} acceleration, \ac{jit} compilation, and autodiff.
+
+In this paper we introduce a new optical simulation package, \dlux, available under an open-source BSD 3 license.
+It significantly extends the capabilities of \texttt{morphine} \cite{pope2021,phase_ret_and_design}, a previous \jax optics package from our team based on \texttt{poppy}, by taking full advantage of more-recent \jax features and libraries.
+It is built in an object-oriented \jax framework \equinox \cite{kidger2021equinox}, with an optical system consisting of an \equinox module wrapping a stack of `layers' which are themselves \equinox layers applying phase or amplitude screens, or performing Fourier or Fresnel transforms on a wavefront.
+We use the optimization library \optax \cite{optax2020github} for gradient-descent optimization, and the probabilistic programming library \numpyro \cite{Phan2019} for \ac{hmc}.
+
 % Autodiff Optics
+There are several other frameworks for automatically-differentiable optics simulation:
 % WaveBlocks
+the most similar is WaveBlocks \cite{page2020}, a PyTorch package for object-oriented modelling of fluorescence microscopy. 
 % Sitzmann etc
+The DeepOptics package \cite{sitzmann2018} has been developed in TensorFlow for camera design and computational imaging, and
 % Liaudat
-% Morphine
-% Overview of Wong et al 2021
-% dLux
-% Equinox
-% Features
-% Example of how to set up a model from layers
+WaveDiff in TensorFlow has been used for detector calibration \cite{liaudat2021} and \ac{psf} modelling \cite{Liaudat2022}. 
+We believe that \dlux fills a unique role in this ecosystem as the only fully-generalised modelling framework for physical optics.
+This has driven a design focus on flexibility and ease of open-source contribution and development.
+The goal is to provide a general object-oriented framework that enables all those working within optics to harness the advantages of autodiff and hardware acceleration provided by \jax with its user-friendly Numpy-like API.
 
-% Figure 1: (Hand-drawn?) Illustration of the Optical System & ∂Lux model layers
+% We believe that integration with \jax, a language of rapidly emerging popularity in physics, makes \dlux a credible alternative to any of these in the context of astronomical imaging.
 
-\section{Phase Retrieval and Flat Field Estimation}
+\section{Optical and Instrumental Modelling}
+\label{sec:simulation}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figs/optics.pdf}
+    \caption{Summary of the optical configuration.
+Left panel: The top plot show the \ac{opd} of the pupil at the aperture of the telescope.
+The binary values create a half-wave step at the mean observation wavelength.
+The bottom plot shows the resulting large single-star \ac{psf} without aberrations applied.
+Middle panel: The top plot shows the total \ac{opd} of the pupil with the optical aberrations applied using low-order Zernike polynomials that is used to generate the data.
+The bottom plot shows the resulting single-star \ac{psf} with the aberrations applied.
+Clearly these aberrations have a large effect on the \ac{psf} and would make recovering information very difficult without appropriate calibration.
+Right panel: The top plot shows the histogram of the PRF that is applied in the focal plane.
+These values have a large spread and would greatly affect any results without calibration.
+The bottom plot shows the full PRF across the whole detector, with a small zoomed region used to show the fine detail that can not be seen when examining the full detector.}
+    \label{fig:optics}
+\end{figure*}
+
+% Describe simulation setup
+% Figure 1: OPD & Pupil in our test, PSF, Flat Field histogram
+Within the $\partial$Lux framework we explicitly define each operation applied to the wavefront as a unique layer.
+Here we employ a simple optical model with only a pupil and focal plane, defined by 6 sequential layers.
+
+\begin{enumerate}
+    \item Wavefront initialisation
+    \item Application of the pupil-support
+    \item Normalisation of the wavefront
+    \item Application of the pupil phase mask
+    \item Application of the optical-aberrations
+    \item Propagation to the focal plane
+\end{enumerate}
+
+The first three layers simply modify the amplitude of the wavefront, representing its response to the opaque components of the optics.
+
+Next the wavefront encounters the binary phase plate diffractive pupil from Toliman \cite{tuthill2018}, with a half-wave total \ac{opd} between the binary regions.
+We use this as an information-preserving way to spread the \ac{psf} across a large number of pixels which encodes the PRF information for a large number of pixels simultaneously.
+Other conventional methods such as defocusing or using a diffuser \cite{Stefansson2017} can be employed but result in a larger degradation in the information carried by the \ac{psf} and therefore result in all round worse performance.
+An important consideration which is alleviated by the Toliman pupil is that focal-plane wavefront sensing suffers a sign ambiguity for all phase modes that are inversion-symmetric about the origin, and for unambiguous phase retrieval we therefore need an asymmetric or odd-mode symmetric pupil \cite{Martinache2013}.
+This pupil and corresponding aberrated \ac{psf} are shown in the left panel of Figure \ref{fig:optics}.
+
+The wavefront then encounters low-order phase errors parameterised by coefficients on Zernike polynomials.
+We include the first 4 radial terms, ignoring global piston, tip and tilt for a total of 7 terms.
+Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20nm, for a total of \textcolor{red}{showyourwork variance} nm RMS total \ac{opd}.
+These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}.
+As can be seen from the \ac{psf} this is a large amount of optical distortion.
+
+This is then propagated to the 1024x1024 pixel focal plane by a two-sided matrix Fourier transform \cite{Soummer2007,Martinache2020}.
+The detector PRF is modelled by multiplying the resulting \ac{psf} by a pixel sensitivity map which is drawn from a normal distribution with mean 0 and standard deviation 0.05.
+The histogram of pixel sensitivities and total pixel sensitivity map is shown in the right panel of Figure~\ref{fig:optics}.
+
+This operation is then performed over 20 stars with a uniform 100nm bandpass spectra sampled at three wavelengths.
+This set of stars is then imaged four times with telescope pointing dithers applied between each image.
+This is done to break the natural degeneracy of stellar fluxes and pixel sensitivities.
+This final model complexity requires the calculations of 240 individual \ac{psf}s over a 1024x1024 detector, which takes a total of \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 CPU.
+The data set is generated by taking this set of images and applying Poissonian noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise.
+Using this set of images we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities.
+This gives us 40 stellar positional parameters (20 RA and DEC each), 20 stellar flux parameters, 7 optical aberration parameters, and 1'048'576 detector parameters.
+The dimensionality is vastly dominated by the \ac{prf}.
+
+\section{Results: Phase Retrieval and Flat Field Estimation}
 \label{sec:phaseretrieval}
-% Describe simulation setup
-% Figure 2: OPD & Pupil in our test, PSF, Flat Field histogram
+
+% Figure 2: Like final 2l figures of tutorial merged together: 
+% Top: loss function vs epoch
+% Middle: correlation plot of Zernikes (with residual below); PSF (true, recovered, residual)
+
+% Bottom: correlation plot of star fluxes (with residual below), correlation plot of sensitivities (with residual below), 2d residual map of pixels
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figs/data_resid.pdf}
+    \caption{The left panel shows a single image of the full four that is used to create the full data set.
+This image has had both photon and detecotor noise applied.
+By eye it is clearly difficult to detangle the astrophysical information.
+The large number of overlapping \ac{psf}s is chosen in order to spread light across the majority of the detector so that we encode the PRF information for as much of the detector as possible.
+The middle panel shows the residual of this single image from the data and the initial uncalibrated model.
+Clearly these residuals are large, showing that there is a large amount of calibration required.
+The right panel shows this same residual after the model has been optimised.
+The residual values are much smaller and at the noise floor.
+A small zoomed region is shown so that the individual pixel-level residuals can be seen.}
+    \label{fig:data_residual}
+\end{figure*}
+
+To recover these parameters we initialise a naive unaberrated optical model with a uniform \ac{prf}.
+Stellar positions and fluxes are similarly perturbed from their true values.
+We will use this model to recover the true values.
+Figure~\ref{fig:data_residual} shows the resultant a sample residual of this model with the data.
 
@@
 % Takes X time on Y machine with Z precision on params (numbers from ShowYourWork)
+In order to recover these input aberrations, we perform gradient descent using Adam \cite{Kingma2014} as implemented in \optax \cite{optax2020github}, minimizing the Poisson log-likelihood with respect to the position and flux of all stars, the response in all pixels, and the Zernike coefficients.
+We also use a staged optimisation strategy, initially optimising the positions, fluxes and optical aberrations in order to recover the correct \ac{psf} model before learning the \ac{prf}.
+This is necessary to circumvent the high degree of covariance that each individual pixel experiences with the parameters of the \ac{psf}.
+To achieve this the learning rate of the \ac{prf} is set to zero for the first 100 epochs of optimisation.
 
-% Figure showing results
+\textcolor{green}{Only by employing gradient evaluations using autodiff makes this problem computationally tractable.
+In comparison to the standard finite-differencing gradient evaluation algorithms with a complexity that scales linearly with the number of parameters, a problem of this dimensionality would take of order a million times longer to compute, which would remain impossible regardless of the hardware, let alone laptop CPU.
+The optimisation is performed for a total of 200 epochs, with each gradient calculation taking \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 Max CPU, leading to total optimisation time of \textcolor{red}{showyourwork variable} minutes.}
 
-% Figure 3: Like final 2l figures of tutorial merged together: 
-% Top: loss function vs epoch
-% Middle: correlation plot of Zernikes (with residual below); PSF (true, recovered, residual)
-% Bottom: correlation plot of star fluxes (with residual below), correlation plot of sensitivities (with residual below), 2d residual map of pixels
+\textcolor{green}{The optical propagation algorithm employs matrix Fourier transforms which uses a matrix multiplication to model the propagation between conjugate pupil-focal planes, and so can be greatly improved if deployed onto GPU hardware.
+Similarly as a test case, we initialise from a very naive state far from the truth, requiring a larger number of optimisation epochs than would be required in a real-world example.
+Furthermore as a calibration algorithm across a whole detector as an example the complexity of the model is greatly increased, applying this to real data would likely be focused on small detector regions resulting in less computationally intensive models.}
+
+\textcolor{green}{Should this green text be moved to discussion?}
+
+This optimisation strategy performs well, with all of the parameters being recovered down to the noise level.
+This can be seen in Figure~\ref{fig:data_residual} with the right panel showing a sample residual of the data and model after optimisation, which is clearly only noise.
+Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}, and as expected from the residuals all of these parameters are well recovered and tightly constrained around their true values.
+Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of \textcolor{red}{showyourwork variable} nm RMS.
+Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram.
+The points of the correlation plot are colored-coded with the total counts for each pixel, showing as expected that pixels with higher photon counts having a better recovery of the PRF since these pixels have a greater \ac{snr}.
+Overall the parameters are very well recovered down to the noise level.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
+    \caption{The left panel shows the recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds.
+The bottom section shows the individual residuals for each.
+Clearly these parameters have been well recovered through the optimisation.
+The right panel shows the recovery of the flux parameters in units of photons.
+The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals.
+\textcolor{red}{Should we address the slight overestimation of fluxes due to the PRF values being a relative measure so a small global covaraince is present?}}
+    \label{fig:astro_params}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figs/aberrations.pdf}
+    \caption{This figure shows the recovery of the optical aberrations after optimisation.
+The left panel shows the amplitudes of the Zernike polynomial coefficients, with the top section showing the correlation plot of the true and recovered values, and the bottom showing the individual residuals.
+The middle and right panels show the the true total \ac{opd} and the residuals generated by these values respectively.}
+    \label{fig:aberrations}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.8\textwidth]{figs/ff.pdf}
+    \caption{This figure shows the PRF values after optimisation.
+The left panel shows the correlation between the pixel-level true and recovered values.
+Each ploint is color-coded with the total flux incident to each pixel, ie its \ac{snr}.
+The pixels with a greater signal are recovered better as expected.
+The right panel shows the histogram of these residuals, with the majority of values being very well recovered and small symmetric about a residual of zero.}
+    \label{fig:flat_field}
+\end{figure*}
 
 \section{Discussion}
-% Can this break down?
+
+This algorithm demonstrates promising performance applied to complex high-dimensional optimisation problems as shown in Section~\ref{sec:phaseretrieval}.
+However, this method does have its limitations.
+The use of a forwards model clearly requires a good match between the true and model \ac{psf}.
+The benefit of using differentiable optical models however is that it provides the perfect platform upon which to directly learn this model from the data, with the ability to scale the complexity as required for each problem.
+Similarly since we are working with direct image data as opposed to reduced data-products is that miss-calibrations of the system are simple to diagnose through direct examination of the residuals of the data and model.
+
+There are also clear limits in which these methods will begin to fail, such as the low \ac{snr} regime.
+Without the ability to recover the \ac{psf} model accurately, pixel level information will be dominated by the noise processes and the PRF will not be sufficiently encoded.
+This can be directly seen in Figure~\ref{fig:flat_field}, with the PRF residual having greater variance in for low-flux pixels.
+This however has a simple solution of observing brighter targets for calibration purposes or observing chosen targets for longer, until a sufficient amount of information is encoded into the data.
+
+It might also seem limited without a dense region of stars to observe as we can not learn about the pixel-level effects in regions with no signal, however we only care about the regions of the detector that have photons landing in them as these are the only ones that will be affected by miss-calibrations in detector level effects.
+
 % Nyquist?
@@
 % Review outcomes
+In this first paper of a planned series, we have demonstrated photometry, phase retrieval, and detector calibration using autodiff to make tractable gradient descent in very high dimensions.
+Our software \dlux integrates with a rich and user-friendly ecosystem of \jax packages for optimization and sampling, which point to a number of natural extensions.
+
+% Time series: regularize with GP?
+The most obvious is to produce calibrated photometric time series, making use of correlations in time rather than the `bag of images' approach taken in this work.
+Instead of independently optimizing each frame, they could be treated hierarchically, with regularization imposed on each star's light curve.
+\ac{gp} models are widely used in astronomy, and \jax packages such as \texttt{tinygp} \cite{Aigrain2022} and GPJax \cite{Pinder2022} permit differentiable optimization and sampling that integrate efficiently with \dlux.
+
 % Useful for space photometry in crowded fields
 % Toliman Pleiades?
+Photometry of crowded fields is often of poor quality even from space telescopes, and there has been considerable investment in algorithms for deblending light curves in the K2 galactic plane campaigns \cite{Zhu2017} and in TESS with its large pixels \cite{Nardiello2019,Hedges2021,Higgins2022}.
+High dimensional optimization with autodiff in principle renders deblending tractable even close to the diffraction limit with unknown \ac{psf}s.
+Prior knowledge of a structured \ac{psf}, such as with the diffractive pupil architecture of Toliman \cite{Guyon2012,Guyon2013}, may further improve this deblending and retrieval.
+
+% JWST, Roman
+High contrast imaging with the recently-launched JWST \cite{Gardner2006} and the upcoming Roman Space Telescope coronagraph \cite{Zellem2022} will in each case demand a new generation of data-driven \ac{psf} and detector calibrations beyond those used fruitfully on ground-based data \cite{Cantalloube2021}.
+While we have not attempted this in the present work, \dlux can flexibly model aberrations in the multiple intermediate planes of these coronagraphs, and in future work we intend to assess the practical and fundamental limits of multiplane phase retrieval and calibration to improve coronagraphic modelling.
 % Useful for detector calibration & monitoring
-% JWST, Roman
-% Time series: regularize with GP?
+Uncertainties in the flat field are also the leading contributor to noise in the exoplanet transmission spectroscopy with JWST \cite{Rustamkulov2022}, and a promising future direction with \dlux will be to build models not just of imaging data, but of spectrophotometry.
 
-% \begin{figure}[ht!]
-%     \script{gen_optics.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/optics.pdf}
-%         \caption{
-%             The optics
-%         }
-%         \label{fig:optics}
-%     \end{centering}
-% \end{figure}
+In Paper II in this series, we will tackle the ismorphic problem of hardware design.
+While we have tackled phase design in previous work \cite{phase_ret_and_design}, autodiff permits calculation and optimization of the Fisher information directly \cite{Coe2009}, which means we can optimize towards the fundamental limits of parameter sensitivity.
+In subsequent papers, we intend to explore parametric and nonparametric deconvolution of images at high angular resolution and contrast.
 
@@
 
-This research made use of \textsc{Numpy} \cite{numpy}; SciPy \cite{scipy}; Matplotlib \cite{matplotlib}; Astropy \cite{astropy:2013,astropy:2018}; \textsc{Jax} \cite{jax}; and \texttt{equinox} \cite{kidger2021equinox}.
+This research made use of \textsc{NumPy} \cite{numpy}; Matplotlib \cite{matplotlib}; \textsc{Jax} \cite{jax}; \texttt{equinox} \cite{kidger2021equinox}; and \optax \cite{optax2020github}.
 

### 2022-11-17 src/scripts/gen_FF.py
Adding updated figure scripts

@@ -1,0 +1,171 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+# import pickle as p
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+# Load model
+tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
+losses = np.load(paths.data / 'losses.npy')
+data = np.load(paths.data / "data.npy")
+
+positions = 'MultiPointSource.position'
+fluxes = 'MultiPointSource.flux'
+zernikes = 'ApplyBasisOPD.coefficients'
+flatfield = 'ApplyPixelResponse.pixel_response'
+parameters = [positions, fluxes, zernikes, flatfield]
+
+
+
+# Get parameters
+nepochs = len(models_out)
+psfs_out = models_out[-1].observe()
+
+positions_found  = np.array([model.get(positions) for model in models_out])
+fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
+zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
+flatfields_found = np.array([model.get(flatfield) for model in models_out])
+
+
+# Get the residuals
+coeff_residuals = tel.get(zernikes) - zernikes_found
+flux_residuals = tel.get(fluxes) - fluxes_found
+
+scaler = 1e3
+positions_residuals = tel.get(positions) - positions_found
+r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
+r_residuals = r2a(r_residuals_rads)
+
+
[…]

### 2022-11-17 src/scripts/gen_aberrations.py

@@ -1,0 +1,145 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+# import pickle as p
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+# Load model
+tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
+losses = np.load(paths.data / 'losses.npy')
+data = np.load(paths.data / "data.npy")
+
+positions = 'MultiPointSource.position'
+fluxes = 'MultiPointSource.flux'
+zernikes = 'ApplyBasisOPD.coefficients'
+flatfield = 'ApplyPixelResponse.pixel_response'
+parameters = [positions, fluxes, zernikes, flatfield]
+
+
+
+# Get parameters
+nepochs = len(models_out)
+psfs_out = models_out[-1].observe()
+
+positions_found  = np.array([model.get(positions) for model in models_out])
+fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
+zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
+flatfields_found = np.array([model.get(flatfield) for model in models_out])
+
+
+# Get the residuals
+coeff_residuals = tel.get(zernikes) - zernikes_found
+flux_residuals = tel.get(fluxes) - fluxes_found
+
+scaler = 1e3
+positions_residuals = tel.get(positions) - positions_found
+r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
+r_residuals = r2a(r_residuals_rads)
+
+
+
+
[…]

### 2022-11-17 src/scripts/gen_astro_params.py

@@ -1,0 +1,127 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+# import pickle as p
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+# Load model
+tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
+losses = np.load(paths.data / 'losses.npy')
+data = np.load(paths.data / "data.npy")
+
+positions = 'MultiPointSource.position'
+fluxes = 'MultiPointSource.flux'
+zernikes = 'ApplyBasisOPD.coefficients'
+flatfield = 'ApplyPixelResponse.pixel_response'
+parameters = [positions, fluxes, zernikes, flatfield]
+
+
+
+# Get parameters
+nepochs = len(models_out)
+psfs_out = models_out[-1].observe()
+
+positions_found  = np.array([model.get(positions) for model in models_out])
+fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
+zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
+flatfields_found = np.array([model.get(flatfield) for model in models_out])
+
+
+# Get the residuals
+coeff_residuals = tel.get(zernikes) - zernikes_found
+flux_residuals = tel.get(fluxes) - fluxes_found
+
+scaler = 1e3
+positions_residuals = tel.get(positions) - positions_found
+r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
+r_residuals = r2a(r_residuals_rads)
+
+
+# Positions
+arcsec = r2a(1)
+true_positions = tel.get(positions).flatten() * arcsec
[…]

### 2022-11-17 src/scripts/gen_data.py

@@ -1,440 +1,0 @@
-# Core jax
-import jax
-import jax.numpy as np
-import jax.random as jr
-
-# Optimisation
-import equinox as eqx
-import optax
-
-# Optics
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
-from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
-import paths
-
-# Pickle
-import pickle as p
-
-# Plotting/visualisation
-import matplotlib.pyplot as plt
-
-plt.rcParams['image.cmap'] = 'inferno'
-plt.rcParams["font.family"] = "serif"
-plt.rcParams["image.origin"] = 'lower'
-plt.rcParams['figure.dpi'] = 120
-
-
-tel = p.load(open(paths.data / "instrument.p"))
-
-# Observe!
-psfs = tel.observe()
-
-# Observe!
-psfs = tel.observe()
-
-'''Figure 1'''
-from matplotlib.cm import get_cmap
-opd = tel.ApplyBasisOPD.get_total_opd()
-psf = tel.model(source=dl.PointSource(wavelengths=wavels))
-throughput = tel.CompoundAperture.get_aperture(npixels=wf_npix)
-pupil = opd.at[np.where(throughput==0.)].set(np.nan)
-
-FF = tel.ApplyPixelResponse.pixel_response
-
-cmap = get_cmap("inferno")
-cmap.set_bad('k',1.)
-
-plt.figure(figsize=(15, 4))
-plt.subplot(1, 3, 1)
-plt.title("Pupil OPD")
-plt.imshow(pupil * 1e9, cmap=cmap)
-plt.xticks([])
-plt.yticks([])
-cbar = plt.colorbar()
-cbar.set_label("OPD (nm)")
-
-plt.subplot(1, 3, 2)
-plt.title("PSF")
-plt.imshow(psf)
-plt.xlabel("Pixels")
-plt.ylabel("Pixels")
-cbar = plt.colorbar()
-cbar.set_label("Probability")
-
-plt.subplot(1, 3, 3)
-plt.title("Pixel Response")
-plt.hist(FF.flatten(), bins=25)
-plt.ylabel("Counts")
-plt.xlabel("Relative Sensitivity")
-
-plt.tight_layout()
-plt.savefig(paths.figures / "optics.pdf", dpi=300)
-plt.show()
-
-
-
-
-
-'''Figure 2'''
-# Apply some noise to the PSF Background noise
[…]

### 2022-11-17 src/scripts/gen_data_resid.py

@@ -1,0 +1,117 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+# import pickle as p
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+# Load model
+tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+source = p.load(open(paths.data / 'source.p', 'rb'))
+
+# Observe!
+psfs = tel.observe()
+
+# Define Params
+positions = 'MultiPointSource.position'
+fluxes = 'MultiPointSource.flux'
+zernikes = 'ApplyBasisOPD.coefficients'
+flatfield = 'ApplyPixelResponse.pixel_response'
+parameters = [positions, fluxes, zernikes, flatfield]
+
+det_pixsize = tel.get("AngularMFT.pixel_scale_out")
+det_npix = tel.get("AngularMFT.npixels_out")
+nzern = len(tel.get(zernikes))
+Nstars = len(source.get('flux'))
+
+# Add small random values to the positions
+model = tel.add(positions, 2.5*det_pixsize*jr.normal(jr.PRNGKey(0),  (Nstars, 2)))
+
+# Multiply the fluxes by small random values
+model = model.multiply(fluxes, 1 + 0.1*jr.normal(jr.PRNGKey(0), (Nstars,)))
+
+# Set the zernike coefficients to zero
+model = model.set(zernikes, np.zeros(nzern))
+
+# Set the flat fiel to uniform
+model = model.set(flatfield, np.ones((det_npix, det_npix)))
+
+
+# Save optimisation model
+p.dump(model, open(paths.data / 'model.p', 'wb'))
+
+
+# Apply some noise to the PSF Background noise
[…]

### 2022-11-17 src/scripts/gen_model.py

@@ -1,0 +1,100 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+# import pickle as p
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+
+
+'''Create Optics, Detector, Source & Instrument'''
+# Define wavelengths
+wavels = 1e-9 * np.linspace(545, 645, 3)
+np.save(paths.data / 'wavelengths.npy', wavels)
+
+# Basic Optical Parameters
+aperture = 0.5
+wf_npix = 512
+
+# Detector Parameters
+det_npix = 1024
+sampling_rate = 20
+det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
+
+# Load mask
+raw_mask = dl.utils.phase_to_opd(np.load("mask.npy"), wavels.mean())
+mask = dl.utils.scale_array(raw_mask, wf_npix, 0)
+
+# Zernike Basis
+zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
+coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), [len(zern_basis)])
+
+# Define Optical Configuration
+optical_layers = [
+    dl.CreateWavefront    (wf_npix, aperture),
+    dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
+    dl.ApplyBasisOPD      (zern_basis, coeffs),
+    dl.AddOPD             (mask),
+    dl.NormaliseWavefront (),
+    dl.AngularMFT         (det_npix, det_pixsize)]
+
+# Create Optics object
+optics = dl.Optics(optical_layers)
+
+# Pixel response
+pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(0), [det_npix, det_npix])
+
+# Create Detector object
[…]

### 2022-11-17 src/scripts/gen_optics.py

@@ -16,108 +16,54 @@ from dLux.utils import radians_to_arcseconds as r2a
 import paths
 
 # Pickle
-import pickle as p
+# import pickle as p
+import dill as p
 
 # Plotting/visualisation
 import matplotlib.pyplot as plt
+from matplotlib.cm import get_cmap
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
+# Load model
+tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+wavels = np.load(paths.data / 'wavelengths.npy')
 
+# Pupil
+throughput = tel.CompoundAperture.get_aperture(npixels=tel.get('CreateWavefront.npixels'))
+mask = tel.AddOPD.opd
+pupil = mask.at[np.where(throughput==0.)].set(np.nan)
 
-
-'''Create Optics, Detector, Source & Instrument'''
-# Define wavelengths
-wavels = 1e-9 * np.linspace(545, 645, 3)
-
-# Basic Optical Parameters
-aperture = 0.5
-wf_npix = 512
-
-# Detector Parameters
-det_npix = 1024
-sampling_rate = 20
-det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
-
-# Load mask
-raw_mask = dl.utils.phase_to_opd(np.load("mask.npy"), wavels.mean())
-mask = dl.utils.scale_array(raw_mask, wf_npix, 0)
-
-# Zernike Basis
-zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
-coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), [len(zern_basis)])
-
-# Define Optical Configuration
-optical_layers = [
-    dl.CreateWavefront    (wf_npix, aperture),
-    dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
-    dl.ApplyBasisOPD      (zern_basis, coeffs),
-    dl.AddOPD             (mask),
-    dl.NormaliseWavefront (),
-    dl.AngularMFT         (det_npix, det_pixsize)]
-
-# Create Optics object
-optics = dl.Optics(optical_layers)
-
-# Pixel response
[…]

### 2022-11-17 src/scripts/optimise.py

@@ -1,0 +1,90 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+# import pickle as p
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+# Load model
+model = p.load(open(paths.data / 'model.p', 'rb'))
+# source = p.load(open(paths.data / 'source.p', 'rb'))
+data = np.load(paths.data / "data.npy")
+
+positions = 'MultiPointSource.position'
+fluxes = 'MultiPointSource.flux'
+zernikes = 'ApplyBasisOPD.coefficients'
+flatfield = 'ApplyPixelResponse.pixel_response'
+parameters = [positions, fluxes, zernikes, flatfield]
+
+# So first we simply set the simple parameters to use an adam optimiser 
+# algorithm, with individual learning rates
+pos_optimiser   = optax.adam(2e-8)
+flux_optimiser  = optax.adam(1e6)
+coeff_optimiser = optax.adam(2e-9)
+
+# Now the flat-field, becuase it is highly covariant with the mean flux level 
+# we don't start learning its parameters until the 100th epoch.
+FF_sched = optax.piecewise_constant_schedule(init_value=1e-2*1e-8, 
+                             boundaries_and_scales={100 : int(1e8)})
+FF_optimiser = optax.adam(FF_sched)
+
+# Combine the optimisers into a list
+optimisers = [pos_optimiser, flux_optimiser, coeff_optimiser, FF_optimiser]
+
+# Generate out optax optimiser, and also get our args
+optim, opt_state, args = model.get_optimiser(parameters, optimisers, get_args=True)
+
+
+@eqx.filter_jit
+@eqx.filter_value_and_grad(arg=args)
+def loss_fn(model, data):
+    out = model.observe()
+    return -np.sum(jax.scipy.stats.poisson.logpmf(data, out))
+
+
+# %%time
[…]

### 2022-11-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[twocolumn]{spie}
+% \documentclass[twocolumn]{spie}
+\documentclass[]{spieman}
 % \documentclass[]{spie}
@@
 \usepackage{graphicx}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-% \usepackage{hyperref}
 \usepackage{relsize}
 \usepackage{acro}
+% \usepackage{authblk}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+
 \acsetup{
@@
 
+\author[a,*]{Louis Desdoigts}
+\author[b]{Benjamin J. S. Pope}
+\author[a]{Peter G. Tuthill}
+
 \affil[a]{Sydney Institute for Astronomy, School of Physics, University of Sydney, NSW~2006, Australia}
@@
 
-\author[a]{Louis C. Desdoigts}
-\author[b]{Benjamin J. S. Pope}
-\author[a]{Peter G. Tuthill}
-
 % Begin!
-\authorinfo{Send correspondence to L. C. D.: \\E-mail: \href{mailto:louis.desdoigts@sydney.edu.au}{louis.desdoigts@sydney.edu.au}}
+% \authorinfo{Send correspondence to L. C. D.: \\E-mail: \href{mailto:louis.desdoigts@sydney.edu.au}{louis.desdoigts@sydney.edu.au}}
 
@@
 \maketitle
+
+% Include email contact information for corresponding author
+{\noindent \footnotesize\textbf{Send correspondence to L. Desdoigts: \\E-mail: \linkable{louis.desdoigts@sydney.edu.au }}}
 
@@
 It significantly extends the capabilities of \texttt{morphine} \cite{pope2021,phase_ret_and_design}, a previous \jax optics package from our team based on \texttt{poppy}, by taking full advantage of more-recent \jax features and libraries.
-It is built in an object-oriented \jax framework \equinox \cite{kidger2021equinox}, with an optical system consisting of an \equinox module wrapping a stack of `layers' which are themselves \equinox layers applying phase or amplitude screens, or performing Fourier or Fresnel transforms on a wavefront.
+It is built in an object-oriented \jax framework \equinox \cite{kidger2021equinox}, with an optical system consisting of an \equinox module wrapping a stack of `layers' which are themselves \equinox modules applying phase or amplitude screens, or performing Fourier or Fresnel transforms on a wavefront.
 We use the optimization library \optax \cite{optax2020github} for gradient-descent optimization, and the probabilistic programming library \numpyro \cite{Phan2019} for \ac{hmc}.
@@
     \item Application of the pupil phase mask
-    \item Application of the optical-aberrations
+    \item Application of the optical aberrations
     \item Propagation to the focal plane
@@
 We use this as an information-preserving way to spread the \ac{psf} across a large number of pixels which encodes the PRF information for a large number of pixels simultaneously.
-Other conventional methods such as defocusing or using a diffuser \cite{Stefansson2017} can be employed but result in a larger degradation in the information carried by the \ac{psf} and therefore result in all round worse performance.
+Alternative methods such as defocusing or using a diffuser \cite{Stefansson2017} come at the cost of eliminating Fourier information which constrain the wavefront and positions.
 An important consideration which is alleviated by the Toliman pupil is that focal-plane wavefront sensing suffers a sign ambiguity for all phase modes that are inversion-symmetric about the origin, and for unambiguous phase retrieval we therefore need an asymmetric or odd-mode symmetric pupil \cite{Martinache2013}.
@@
 We include the first 4 radial terms, ignoring global piston, tip and tilt for a total of 7 terms.
-Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20nm, for a total of \textcolor{red}{showyourwork variance} nm RMS total \ac{opd}.
-These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}.
-As can be seen from the \ac{psf} this is a large amount of optical distortion.
+Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of \textcolor{red}{showyourwork variance}\,nm RMS total \ac{opd}.
+These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}, where we see this induces a large amount of distortion in the \ac{psf}.
 
-This is then propagated to the 1024x1024 pixel focal plane by a two-sided matrix Fourier transform \cite{Soummer2007,Martinache2020}.
+This is then propagated to the $1024\times1024$ pixel focal plane by a two-sided matrix Fourier transform \cite{Soummer2007,Martinache2020}.
 The detector PRF is modelled by multiplying the resulting \ac{psf} by a pixel sensitivity map which is drawn from a normal distribution with mean 0 and standard deviation 0.05.
@@
 
-This operation is then performed over 20 stars with a uniform 100nm bandpass spectra sampled at three wavelengths.
+This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths.
 This set of stars is then imaged four times with telescope pointing dithers applied between each image.
 This is done to break the natural degeneracy of stellar fluxes and pixel sensitivities.
-This final model complexity requires the calculations of 240 individual \ac{psf}s over a 1024x1024 detector, which takes a total of \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 CPU.
-The data set is generated by taking this set of images and applying Poissonian noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise.
+This total model requires the calculation of 240 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 CPU.
+The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise.
 Using this set of images we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities.
-This gives us 40 stellar positional parameters (20 RA and DEC each), 20 stellar flux parameters, 7 optical aberration parameters, and 1'048'576 detector parameters.
-The dimensionality is vastly dominated by the \ac{prf}.
+This gives us 40 stellar positional parameters (20 RA and DEC each), 20 stellar flux parameters, 7 optical aberration parameters, and 1,048,576 detector parameters: the dimensionality of the problem is vastly dominated by the \ac{prf} parameters, which is why this problem is made tractable only by using autodiff.
 
@@
     \includegraphics[width=\textwidth]{figs/data_resid.pdf}
-    \caption{The left panel shows a single image of the full four that is used to create the full data set.
-This image has had both photon and detecotor noise applied.
-By eye it is clearly difficult to detangle the astrophysical information.
+    \caption{Left: single image of the full four that is used to create the full data set.
+This image has had both photon and detector noise applied.
+By eye it is clearly difficult to disentangle the astrophysical information.
 The large number of overlapping \ac{psf}s is chosen in order to spread light across the majority of the detector so that we encode the PRF information for as much of the detector as possible.
-The middle panel shows the residual of this single image from the data and the initial uncalibrated model.
+Middle: the residual of this single image from the data and the initial uncalibrated model.
 Clearly these residuals are large, showing that there is a large amount of calibration required.
-The right panel shows this same residual after the model has been optimised.
+Right: this same residual after the model has been optimised.
 The residual values are much smaller and at the noise floor.
@@
 We will use this model to recover the true values.
-Figure~\ref{fig:data_residual} shows the resultant a sample residual of this model with the data.
+Figure~\ref{fig:data_residual} shows maps of the initial guess and initial and final residuals.
 
@@
 
-\textcolor{green}{Only by employing gradient evaluations using autodiff makes this problem computationally tractable.
-In comparison to the standard finite-differencing gradient evaluation algorithms with a complexity that scales linearly with the number of parameters, a problem of this dimensionality would take of order a million times longer to compute, which would remain impossible regardless of the hardware, let alone laptop CPU.
-The optimisation is performed for a total of 200 epochs, with each gradient calculation taking \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 Max CPU, leading to total optimisation time of \textcolor{red}{showyourwork variable} minutes.}
-
-\textcolor{green}{The optical propagation algorithm employs matrix Fourier transforms which uses a matrix multiplication to model the propagation between conjugate pupil-focal planes, and so can be greatly improved if deployed onto GPU hardware.
-Similarly as a test case, we initialise from a very naive state far from the truth, requiring a larger number of optimisation epochs than would be required in a real-world example.
-Furthermore as a calibration algorithm across a whole detector as an example the complexity of the model is greatly increased, applying this to real data would likely be focused on small detector regions resulting in less computationally intensive models.}
-
-\textcolor{green}{Should this green text be moved to discussion?}
+The optimisation is performed for a total of 200 epochs, with each gradient calculation taking \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 Max CPU, leading to a total optimisation time of \textcolor{red}{showyourwork variable} minutes.
 
@@
 Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram.
-The points of the correlation plot are colored-coded with the total counts for each pixel, showing as expected that pixels with higher photon counts having a better recovery of the PRF since these pixels have a greater \ac{snr}.
+The points of the correlation plot are colour-coded with the total counts for each pixel, showing as expected that pixels with higher photon counts having a better recovery of the PRF since these pixels have a greater \ac{snr}.
 Overall the parameters are very well recovered down to the noise level.
+
+The optical propagation algorithm employs two-sided matrix Fourier transforms \cite{Soummer2007}, which can be greatly improved if deployed onto GPU hardware.
+In our a test case, we initialise from a very naive state far from the truth, requiring a larger number of optimisation epochs than would be required in a real-world example.
+Furthermore as a calibration algorithm across a whole detector as an example the complexity of the model is greatly increased, applying this to real data would likely be focused on small detector regions resulting in less computationally intensive models.
 
@@
     \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
-    \caption{The left panel shows the recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds.
+    \caption{Left: recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds.
 The bottom section shows the individual residuals for each.
 Clearly these parameters have been well recovered through the optimisation.
-The right panel shows the recovery of the flux parameters in units of photons.
+Right: the recovery of the flux parameters in units of photons.
 The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals.
@@
     \includegraphics[width=\textwidth]{figs/aberrations.pdf}
-    \caption{This figure shows the recovery of the optical aberrations after optimisation.
-The left panel shows the amplitudes of the Zernike polynomial coefficients, with the top section showing the correlation plot of the true and recovered values, and the bottom showing the individual residuals.
-The middle and right panels show the the true total \ac{opd} and the residuals generated by these values respectively.}
+    \caption{Recovery of the optical aberrations after optimisation.
+Left: the amplitudes of the Zernike polynomial coefficients, with the top section showing the correlation plot of the true and recovered values, and the bottom showing the individual residuals.
+Middle and right: the true total \ac{opd} and the residuals generated by these values respectively.}
     \label{fig:aberrations}

### 2022-11-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{acro}
-% \usepackage{authblk}
+\usepackage{listings}
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
@@
 \newcommand\numpyro{\texttt{numpyro}\xspace}
-
 \newcommand\kepler{\textit{Kepler}\xspace}
@@
 Here we employ a simple optical model with only a pupil and focal plane, defined by 6 sequential layers.
+Here we show how to create this optical system within the $\partial$Lux framework:
 
-\begin{enumerate}
-    \item Wavefront initialisation
-    \item Application of the pupil-support
-    \item Normalisation of the wavefront
-    \item Application of the pupil phase mask
-    \item Application of the optical aberrations
-    \item Propagation to the focal plane
-\end{enumerate}
+\begin{lstlisting}[language=Python]
+import jax.numpy as np
+import jax.random as jr
+import dLux as dl
+
+# Define wavelengths
+wavels = 1e-9 * np.linspace(545, 645, 3)
+
+# Basic Optical Parameters
+aperture = 0.5
+wf_npix = 512
+
+# Detector Parameters
+det_npix = 1024
+sampling_rate = 20
+det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
+
+# Zernike Basis (ignore piston, tip, tilt)
+zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
+coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), (len(zern_basis),))
+
+# Define Optical Configuration
+optical_layers = [
+    dl.CreateWavefront    (wf_npix, aperture),
+    dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
+    dl.ApplyBasisOPD      (zern_basis, coeffs),
+    dl.AddOPD             (np.load("mask.npy")),
+    dl.NormaliseWavefront (),
+    dl.AngularMFT         (det_npix, det_pixsize)]
+
+# Create Optics object
+optics = dl.Optics(optical_layers)
+\end{lstlisting}
 
@@
 We include the first 4 radial terms, ignoring global piston, tip and tilt for a total of 7 terms.
-Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of \textcolor{red}{showyourwork variance}\,nm RMS total \ac{opd}.
+Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of \variable{output/opd_in.txt}\,nm RMS total \ac{opd}.
 These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}, where we see this induces a large amount of distortion in the \ac{psf}.
@@
 The histogram of pixel sensitivities and total pixel sensitivity map is shown in the right panel of Figure~\ref{fig:optics}.
+Here we show how to create this within the $\partial$Lux framework and create an instrument object to model this:
 
-This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths.
+\begin{lstlisting}[language=Python]
+# Pixel response
+pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(0), [det_npix, det_npix])
+
+# Create Detector object
+detector = dl.Detector([dl.ApplyPixelResponse(pix_response)])
+
+# Combine into instrument
+tel = dl.Instrument(optics=optics, detector=detector)
+\end{lstlisting}
+
+This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm.
 This set of stars is then imaged four times with telescope pointing dithers applied between each image.
@@
 This total model requires the calculation of 240 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 CPU.
-The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise.
+The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise with mean 10 and variance 2.5.
+This gives us an average of \textcolor{red}{showyourwork variable} photons per pixel.
 Using this set of images we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities.
@@
 Right: the recovery of the flux parameters in units of photons.
-The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals.
-\textcolor{red}{Should we address the slight overestimation of fluxes due to the PRF values being a relative measure so a small global covaraince is present?}}
+The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals.}
     \label{fig:astro_params}

### 2022-11-21 src/scripts/gen_FF.py
updating to 0.9.2

@@ -68,7 +68,8 @@ pix_response = tel.get(flatfield)
 # Plot
 # calculate the mask where there was enough flux to infer the flat field
 # thresh = 2500
-thresh = 1000
+# thresh = 1000
+thresh = 500
 fmask = data.mean(0) >= thresh
 
 out_mask = np.where(data.mean(0) < thresh)
@@ -124,8 +125,7 @@ pr_found_flat = found_pr_masked.flatten()
 pr_true_sort = pr_true_flat[ind]
 pr_found_sort = pr_found_flat[ind]
 
-plt.subplot(1, 2, 1)
-plt.plot(np.linspace(0.8, 1.2), np.linspace(0.8, 1.2), c='k', alpha=0.75)
+ax = plt.subplot(1, 2, 1)
 plt.scatter(pr_true_sort, pr_found_sort, c=colours*1e-3, alpha=0.5, rasterized=True)
 cbar = plt.colorbar()
 cbar.set_label("Counts (Photons $x10^3$)")
@@ -133,15 +133,51 @@ plt.title("PRF Correlation")
 plt.ylabel("Recovered")
 plt.xlabel("True")
 
+xlims = ax.get_xlim()
+ylims = ax.get_ylim()
+ax.set_xlim(xlims)
+ax.set_ylim(ylims)
+plt.plot(np.linspace(0.7, 1.3), np.linspace(0.7, 1.3), c='k', alpha=0.5)
 
-plt.subplot(1, 2, 2)
+thresh_indx = np.where(fmask)
+res = (pix_response - flatfields_found[-1])[thresh_indx].flatten()
+
+ax2 = plt.subplot(1, 2, 2)
 # plt.plot(np.linspace(0.8, 1.2), np.linspace(0.8, 1.2), c='k', alpha=0.75)
 # plt.hist((pr_true_sort - pr_found_sort).flatten(), bins=25)
-plt.hist((pr_true_sort - pr_found_sort).flatten(), bins=51)
-# plt.colorbar()
+# plt.hist((pr_true_sort - pr_found_sort).flatten(), bins=51)
+# plt.hist((pr_true_sort - pr_found_sort).flatten(), bins=101)
+# plt.hist((pr_true_sort - pr_found_sort).flatten(), bins=201)
+plt.hist(res, bins=51)
 plt.title("PRF Residual Histogram")
 plt.ylabel("Counts")
 plt.xlabel("Residual")
+
+xlim = np.abs(np.array(ax2.get_xlim())).max()
[…]

### 2022-11-21 src/scripts/gen_aberrations.py

@@ -96,17 +96,29 @@ fig1.add_axes(ax2)
 
 plt.title("Zernike Coefficients (nm)")
 ax2.scatter(true_coeff*1e9, found_coeff*1e9)
-vals = np.linspace(true_coeff.min(), true_coeff.max(), 2)
-ax2.plot(vals*1e9, vals*1e9, c='k', alpha=0.5)
+
 ax2.set_xticks([])
 ax2.set_ylabel("Recovered")
 # ax2.yaxis.set_tick_params(labelbottom=False)
+
+xlims = ax2.get_xlim()
+ylims = ax2.get_ylim()
+
+vals = 1e9*np.linspace(1.2*true_coeff.min(), 1.2*true_coeff.max(), 2)
+ax2.plot(vals, vals, c='k', alpha=0.5)
+
+# vals = np.linspace(0.8*true_fluxes.min(), 1.2*true_fluxes.max(), 2)
+# ax4.plot(vals, vals, c='k', alpha=0.5)
+ax2.set_xlim(xlims)
+ax2.set_ylim(ylims)
 
 ax.scatter(true_coeff*1e9, (true_coeff - found_coeff)*1e9)
 # ax.scatter(true_coeff, (true_coeff - found_coeff)/true_coeff)
 ax.axhline(0, c='k')
 ax.set_ylabel("Residuals")
 ax.set_xlabel("True")
+
+ax.set_ylim(1.1 * np.array(ax.get_ylim()))
 
 
 
@@ -116,26 +128,38 @@ true_opd = true_opd.at[np.where(throughput==0.)].set(np.nan)
 found_opd = found_opd.at[np.where(throughput==0.)].set(np.nan)
 
 true_vales = true_opd[np.where(throughput==1.)]
-found_vales = (true_opd-found_opd)[np.where(throughput==1.)]
+found_vales = found_opd[np.where(throughput==1.)]
+residual_vales = (true_opd-found_opd)[np.where(throughput==1.)]
 
 true_rms = (true_vales**2).mean()**0.5
 found_rms = (found_vales**2).mean()**0.5
+residual_rms = (residual_vales**2).mean()**0.5
+
+# Save values
+with open(paths.output / "rms_opd_in.txt", 'w') as f:
+    f.write("{:.3}".format(true_rms*1e9))
+
+with open(paths.output / "rms_opd_resid.txt", 'w') as f:
[…]

### 2022-11-21 src/scripts/gen_astro_params.py

@@ -31,9 +31,13 @@ plt.rcParams['figure.dpi'] = 120
 
 # Load model
 tel = p.load(open(paths.data / 'instrument.p', 'rb'))
-models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
-losses = np.load(paths.data / 'losses.npy')
+# models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
+# losses = np.load(paths.data / 'losses.npy')
+# data = np.load(paths.data / "data.npy")
+models_out = p.load(open(paths.data / 'models_out_.p', 'rb'))
+losses = np.load(paths.data / 'losses_.npy')
 data = np.load(paths.data / "data.npy")
+
 
 positions = 'MultiPointSource.position'
 fluxes = 'MultiPointSource.flux'
@@ -69,11 +73,6 @@ true_positions = tel.get(positions).flatten() * arcsec
 initial_positions = positions_found[0].flatten() * arcsec
 final_positions = positions_found[-1].flatten() * arcsec
 
-# Fluxes
-true_fluxes = tel.get(fluxes) * 1e-6
-final_fluxes = fluxes_found[-1] * 1e-6
-initial_fluxes = fluxes_found[0] * 1e-6
-
 
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 fig = plt.figure(figsize=(10, 4))
@@ -86,18 +85,38 @@ ax2 = divider.new_vertical(size="150%", pad=0.00)
 fig.add_axes(ax2)
 ax2.set_title("Positions (arcseconds)")
 
-vals = np.linspace(true_positions.min(), true_positions.max(), 2)
-ax2.plot(vals, vals, c='k', alpha=0.5)
 ax2.scatter(true_positions, final_positions, label='final')
 # ax2.scatter(true_positions, initial_positions, c='tab:orange', marker='x', label="intitial")
+
+xlims = ax2.get_xlim()
+ylims = ax2.get_ylim()
+
+vals = np.linspace(1.2*true_positions.min(), 1.2*true_positions.max(), 2)
+ax2.plot(vals, vals, c='k', alpha=0.5)
 ax2.set_ylabel('Recovered')
 ax2.set_xticks([])
+ax2.set_xlim(xlims)
+ax2.set_ylim(ylims)
 
[…]

### 2022-11-21 src/scripts/gen_data_resid.py

@@ -21,6 +21,7 @@ import dill as p
 
 # Plotting/visualisation
 import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import make_axes_locatable
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
@@ -31,9 +32,6 @@ plt.rcParams['figure.dpi'] = 120
 # Load model
 tel = p.load(open(paths.data / 'instrument.p', 'rb'))
 source = p.load(open(paths.data / 'source.p', 'rb'))
-
-# Observe!
-psfs = tel.observe()
 
 # Define Params
 positions = 'MultiPointSource.position'
@@ -63,54 +61,71 @@ model = model.set(flatfield, np.ones((det_npix, det_npix)))
 # Save optimisation model
 p.dump(model, open(paths.data / 'model.p', 'wb'))
 
+# Observe!
+bg_val = tel.AddConstant.value
+psfs = tel.observe() - bg_val
+print(psfs.sum(0).mean())
+
 
 # Apply some noise to the PSF Background noise
-BG_noise = np.abs(5*jr.normal(jr.PRNGKey(0), psfs.shape))
+BG_noise = 2.5*jr.normal(jr.PRNGKey(0), psfs.shape) + bg_val
 data = jr.poisson(jr.PRNGKey(0), psfs) + BG_noise
 np.save(paths.data / "data", data)
 
 plt.figure(figsize=(15, 4.5))
 plt.suptitle("Data and Residuals", size=15)
 
-plt.subplot(1, 3, 1)
+ax1 = plt.subplot(1, 3, 1)
 plt.title("Sample Data")
 plt.imshow(data[0]*1e-3)
 plt.xlabel("Pixels")
 plt.ylabel("Pixels")
-cbar = plt.colorbar()
+divider = make_axes_locatable(ax1)
+cax = divider.append_axes("right", size="5%", pad=0.1)
+cbar = plt.colorbar(cax=cax)
 cbar.set_label("Counts $x10^3$")
 
 
 initital_psfs = model.observe()
-plt.subplot(1, 3, 2)
+ax2 = plt.subplot(1, 3, 2)
 plt.title("Initial Residual")
 plt.imshow((initital_psfs[0] - data[0])*1e-3)
+# plt.imshow(data[0]/initital_psfs[0])
 plt.xlabel("Pixels")
 plt.ylabel("Pixels")
[…]

### 2022-11-21 src/scripts/gen_model.py

@@ -67,8 +67,15 @@ optics = dl.Optics(optical_layers)
 # Pixel response
 pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(0), [det_npix, det_npix])
 
+# Detector Noise
+bg_mean = 10
+
+# Create detector layers
+detector_layers =[dl.ApplyPixelResponse(pix_response),
+                  dl.AddConstant(bg_mean)]
+
 # Create Detector object
-detector = dl.Detector([dl.ApplyPixelResponse(pix_response)])
+detector = dl.Detector(detector_layers)
 
 # Observation stratergy, define dithers
 dithers = 2**-.5 * det_pixsize * np.array([[+1, +1],

### 2022-11-21 src/scripts/optimise.py

@@ -86,5 +86,7 @@ with tqdm(range(200),desc='Gradient Descent') as t:
         models_out.append(model)
         t.set_description("Log Loss: {:.3f}".format(np.log10(loss))) # update the progress bar
 
-np.save(paths.data /'losses', np.array(losses))
-p.dump(models_out, open(paths.data / "models_out.p", 'wb'))
+# np.save(paths.data /'losses', np.array(losses))
+# p.dump(models_out, open(paths.data / "models_out.p", 'wb'))
+np.save(paths.data /'losses_', np.array(losses))
+p.dump(models_out, open(paths.data / "models_out_.p", 'wb'))

### 2022-11-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We include the first 4 radial terms, ignoring global piston, tip and tilt for a total of 7 terms.
-Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of \variable{output/opd_in.txt}\,nm RMS total \ac{opd}.
+Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of \variable{output/rms_opd_in.txt}\,nm RMS total \ac{opd}.
 These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}, where we see this induces a large amount of distortion in the \ac{psf}.

### 2022-11-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}, and as expected from the residuals all of these parameters are well recovered and tightly constrained around their true values.
-Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of \textcolor{red}{showyourwork variable} nm RMS.
+Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of \variable{output/rms_opd_resid.txt} nm RMS.
 Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram.

### 2022-11-21 src/scripts/gen_aberrations.py
Fixing dill import

@@ -143,7 +143,7 @@ with open(paths.output / "rms_opd_resid.txt", 'w') as f:
     f.write("{:.3}".format(residual_rms*1e9))
 
 from matplotlib.cm import get_cmap
-cmap = get_cmap("inferno")
+cmap = get_cmap("inferno").copy()
 cmap.set_bad('k',1.)
 
 plt.subplot(1, 3, 2)

### 2022-11-21 src/scripts/gen_optics.py

@@ -49,7 +49,7 @@ zernikes = 'ApplyBasisOPD.coefficients'
 psf = tel.set(zernikes, np.zeros(len(tel.get(zernikes)))).model(source=dl.PointSource(wavelengths=wavels))
 aberrated_psf = tel.model(source=dl.PointSource(wavelengths=wavels))
 
-cmap = get_cmap("inferno")
+cmap = get_cmap("inferno").copy()
 cmap.set_bad('k',1.)
 
 plt.figure(figsize=(15, 8))

### 2022-11-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \centering
-    \includegraphics[width=\textwidth]{figs/optics.pdf}
+    \includegraphics[width=\textwidth]{figures/optics.pdf}
     \caption{Summary of the optical configuration.
@@
     \label{fig:optics}
+    \script{gen_optics.py}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=\textwidth]{figs/data_resid.pdf}
+    \includegraphics[width=\textwidth]{figures/data_resid.pdf}
     \caption{Left: single image of the full four that is used to create the full data set.
@@
     \label{fig:data_residual}
+    \script{gen_data_resid.py}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
+    \includegraphics[width=0.8\textwidth]{figures/astro_params.pdf}
     \caption{Left: recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds.
@@
     \label{fig:astro_params}
+    \script{gen_astro_params.py}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=\textwidth]{figs/aberrations.pdf}
+    \includegraphics[width=\textwidth]{figures/aberrations.pdf}
     \caption{Recovery of the optical aberrations after optimisation.
@@
     \label{fig:aberrations}
+    \script{gen_aberrations.py}
 \end{figure*}
@@
     \label{fig:flat_field}
+    % \script{gen_FF.py}
 \end{figure*}
@@
 
+\subsection{Limitations}
 This algorithm demonstrates promising performance applied to complex high-dimensional optimisation problems as shown in Section~\ref{sec:phaseretrieval}.
 However, this method does have its limitations.
+
+% poor model
 The use of a forwards model clearly requires a good match between the true and model \ac{psf}.
 The benefit of using differentiable optical models however is that it provides the perfect platform upon which to directly learn this model from the data, with the ability to scale the complexity as required for each problem.
-Similarly since we are working with direct image data as opposed to reduced data-products is that miss-calibrations of the system are simple to diagnose through direct examination of the residuals of the data and model.
+Similarly since we are working with direct image data as opposed to reduced data-products, miss-calibrations of the system are simple to diagnose through direct examination of the residuals of the data and model.
+Figure~\ref{fig:data_residual} demonstrates that in this example we can be sure of a good fit because the final residual is purely noise
 
-There are also clear limits in which these methods will begin to fail, such as the low \ac{snr} regime.
+% Single noise realisation
+We further examine the limits in which this method begins to break down.
+When only examining a single image each pixel only has a single realisation of both photon and detector noise, resulting in an over-fitting of the PRF fit to this noise.
+Furthermore, in the low \ac{snr} regime, the model can fail to converge due to divergent flux values as the model tries to correct for the noise across the whole image.
+We test this limitation by fitting a single-star using a single image with increases levels of flux from $10^4$\,photons to $10^7$\,photons, using the same optical prescription as described in Section~\ref{sec:simulation}.
+Figure~\ref{fig:diverge} show the correlation of the recovered fluxes and the fractional flux errors.
+The low \ac{snr} models diverge from their true values
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.8\textwidth]{figs/divergence.png}
+    \caption{\textcolor{red}{TODO: Caption}}
+    \label{fig:diverge}
+    % \script{gen_FF.py}
+\end{figure*}
+
+In the high \ac{snr} regime and hence the per-pixel freedom recovered With near complete per-pixel freedom and only a single realisation of noise ie when only using a single image, each pixel
+Taking the low-flux single-star single image
+There are also clear limits in which these methods will begin to fail, such as the low \ac{snr} regime. One
+
 Without the ability to recover the \ac{psf} model accurately, pixel level information will be dominated by the noise processes and the PRF will not be sufficiently encoded.
@@
 
+\subsection{Other?}
 It might also seem limited without a dense region of stars to observe as we can not learn about the pixel-level effects in regions with no signal, however we only care about the regions of the detector that have photons landing in them as these are the only ones that will be affected by miss-calibrations in detector level effects.

### 2022-12-12 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2022-11-21 src/tex/ms.tex
@@
 
+\definecolor{mygreen}{rgb}{0,0.6,0}
+\definecolor{mygray}{rgb}{0.5,0.5,0.5}
+\definecolor{mymauve}{rgb}{0.58,0,0.82}
+
+\lstset{ %
+  backgroundcolor=\color{white},   % choose the background color
+  % basicstyle=\footnotesize,        % size of fonts used for the code
+  breaklines=true,                 % automatic line breaking only at whitespace
+  captionpos=b,                    % sets the caption-position to bottom
+  commentstyle=\color{mygreen},    % comment style
+  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
+  keywordstyle=\color{blue},       % keyword style
+  stringstyle=\color{mymauve},     % string literal style
+}
+
 \acsetup{
@@
   cite/group/cmd = \cite,
-  cite/group/pre = {},
+  cite/group/pre = {\xspace},
 }
@@
 
-\title{Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase with autodiff}
+\title{Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase Retrieval with Automatic Differentiation}
 % \title{Differentiable Optics with dLux: I - Deep Calibration of Flat Field and Phase with autodiff}
@@
 \author[a,*]{Louis Desdoigts}
-\author[b]{Benjamin J. S. Pope}
+\author[b,c]{Benjamin J. S. Pope}
 \author[a]{Peter G. Tuthill}
@@
 
-% Abstract with filler text
 \begin{abstract}
-
-$\partial$Lux is a newly developed optical modelling framework deigned to harness the powerful mathematical tools underpinning machine learning and directly apply them to the field of optics. 
-Both neural networks and optical systems map input vectors to output vectors employing a series of intermediary linear transformations and non-linear matrix operations. 
-This isomorphism allows for optical models to be directly constructed within existing automatic differentiation (autodiff) libraries. 
-$\partial$Lux exploits this relationship harnessing autodiff libraries to create a naively end-to-end fully differentiable optical modelling framework. 
-This may precipitate a paradigm shift in the power and utility of these optical models, opening the possibility to entirely novel algorithms and approaches. 
-Astronomical optics is one field where notoriously challenging or intractable problems can be re-approached with new and simple techniques. 
-This manuscript explores how efficient high-dimensional gradient calculations enabled by autodiff can be harnessed using differentiable optical models in order to simultaneously recover uncalibrated astronomical, optical and instrumental parameters.
-
+% phase retrieval and detector calibration are important limiting factors in many space missions
+The sensitivity limits of space telescopes are imposed by uncalibrated errors in the point spread function, background light, and detector sensitivity. 
+%these are hard to calibrate
+These are typically calibrated with specialized wavefront sensor hardware and with flat fields obtained on the ground or with calibration sources, but these leave vulnerabilities to residual time-varying or non-common path aberrations and variations in the detector conditions.
+It is therefore desirable to infer these from science data alone, facing the prohibitively high dimensional problems of phase retrieval and pixel-level calibration.
+% we introduce a new package dLux for differentiable, gpu-accelerated physical optics simulation
+We introduce a new Python package for physical optics simulation, \dlux, which uses the machine learning framework \jax to achieve GPU acceleration and automatic differentiation, and apply this to simulating astronomical imaging.
+% we show we can do flat field and phase retrieval simultaneously
+In this first of a series of papers, we show that gradient descent enabled by autodiff can be used to simultaneously perform phase retrieval and calibration of detector sensitivity, scaling efficiently to inferring millions of parameters.
+% this is the first of a series of applications
+This new framework enables high dimensional optimization and inference in data analysis and hardware design in astronomy and beyond, which we explore in subsequent papers in this series.
 \end{abstract}
@@
 % High precision astronomy is limited by systematics introduced by the optics and detector
-At the bleeding edge of space-based astronomical imaging, photometry, and spectroscopy, the precision is limited by systematics introduced by aberrations in the optics and noise processes in the detector.
+At the vanguard of space-based astronomical imaging, photometry, and spectroscopy, imaging precision is limited by systematics introduced by aberrations in the optics and noise processes in the detector.
 While problems are ubiquitous in astronomy, we are motivated by several core examples.
@@
 
-In each case, we face serious limitations from an imperfect knowledge of the \ac{prf}, the map of the intra-pixel and inter-pixel variations in detector sensitivity; and the \ac{psf}, the diffraction-limited pattern by which light from a point source like a star is spatially spread out across a detector.
+In each case, we face serious limitations from an imperfect knowledge of the \ac{prf} -- the map of the intra-pixel and inter-pixel variations in detector sensitivity; and also of the \ac{psf} -- the diffraction-limited pattern by which light from a point source like a starspreads across a detector.
 
-In this series of papers, we present a new software package, \dlux, for fitting high-dimensional parametrized physical optics models to astronomical data.
+In this series of papers, we present a new software package, \dlux, designed for fitting high-dimensional parametrized physical optics models to astronomical data.
 By using the Python library \textsc{Jax}\cite{jax}, we obtain \ac{gpu} hardware acceleration, as well as autodiff\cite{Margossian2018} features that enable high-dimensional optimization and inference with gradient descent or \ac{hmc}.
@@
 It significantly extends the capabilities of \texttt{morphine} \cite{pope2021,phase_ret_and_design}, a previous \jax optics package from our team based on \texttt{poppy}, by taking full advantage of more-recent \jax features and libraries.
-It is built in an object-oriented \jax framework \equinox \cite{kidger2021equinox}, with an optical system consisting of an \equinox module wrapping a stack of `layers' which are themselves \equinox modules applying phase or amplitude screens, or performing Fourier or Fresnel transforms on a wavefront.
+It is built in an object-oriented \jax framework called \equinox \cite{kidger2021equinox}, with an optical system consisting of an \equinox module wrapping a stack of `layers' which are themselves \equinox modules applying phase or amplitude screens, or performing Fourier or Fresnel transforms on a wavefront.
 We use the optimization library \optax \cite{optax2020github} for gradient-descent optimization, and the probabilistic programming library \numpyro \cite{Phan2019} for \ac{hmc}.
@@
     \centering
+    % \includegraphics[width=\textwidth]{figures/optics.pdf}
     \includegraphics[width=\textwidth]{figs/optics.pdf}
@@
     \label{fig:optics}
+    % \script{gen_optics.py}
 \end{figure*}
@@
 
-\begin{lstlisting}[language=Python]
+\newpage
+
+\begin{lstlisting}[language=Python,frame=single]
 import jax.numpy as np
@@
 zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
-coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), (len(zern_basis),))
+coeffs = 2e-8 * jr.normal(jr.PRNGKey(1), (len(zern_basis),))
 
@@
 Here we show how to create this within the $\partial$Lux framework and create an instrument object to model this:
-
-\begin{lstlisting}[language=Python]
+\newpage
+\begin{lstlisting}[language=Python,frame=single]
 # Pixel response
-pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(0), [det_npix, det_npix])
+pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(2), [det_npix, det_npix])
 
@@
 
-This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm.
-This set of stars is then imaged four times with telescope pointing dithers applied between each image.
-This is done to break the natural degeneracy of stellar fluxes and pixel sensitivities.
-This total model requires the calculation of 240 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 CPU.
-The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise with mean 10 and variance 2.5.
-This gives us an average of \textcolor{red}{showyourwork variable} photons per pixel.
-Using this set of images we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities.
-This gives us 40 stellar positional parameters (20 RA and DEC each), 20 stellar flux parameters, 7 optical aberration parameters, and 1,048,576 detector parameters: the dimensionality of the problem is vastly dominated by the \ac{prf} parameters, which is why this problem is made tractable only by using autodiff.
+This operation is then performed over 15 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm..
+This total model requires the calculation of 30 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 CPU.
+The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise with mean 10 and deviation 2.5 Using this image we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities.
+This gives us 30 stellar positional parameters (15 RA and DEC each), 15 stellar flux parameters, 7 optical aberration parameters, and 1,048,576 detector parameters: the dimensionality of the problem is vastly dominated by the \ac{prf} parameters, which is why this problem is made tractable only by using autodiff.
 
-\section{Results: Phase Retrieval and Flat Field Estimation}
+\section{Optimisation and Results: Phase Retrieval and Flat Field Estimation}
 \label{sec:phaseretrieval}
@@
     \label{fig:data_residual}
+    % \script{gen_data_resid.py}
 \end{figure*}
@@
 To recover these parameters we initialise a naive unaberrated optical model with a uniform \ac{prf}.
-Stellar positions and fluxes are similarly perturbed from their true values.
+Stellar positions and fluxes are randomly perturbed from their true values.
 We will use this model to recover the true values.
-Figure~\ref{fig:data_residual} shows maps of the initial guess and initial and final residuals.
+Figure~\ref{fig:data_residual} shows the data and the initial and final residuals.
 
@@
 % Takes X time on Y machine with Z precision on params (numbers from ShowYourWork)
-In order to recover these input aberrations, we perform gradient descent using Adam \cite{Kingma2014} as implemented in \optax \cite{optax2020github}, minimizing the Poisson log-likelihood with respect to the position and flux of all stars, the response in all pixels, and the Zernike coefficients.
-We also use a staged optimisation strategy, initially optimising the positions, fluxes and optical aberrations in order to recover the correct \ac{psf} model before learning the \ac{prf}.
+In order to recover these input aberrations, we perform gradient descent using Adam \cite{Kingma2014} as implemented in \optax \cite{optax2020github}, minimizing the posterior of a per-pixel Poisson log-likelihood and a $\chi^2$ log prior on the PRF values, matching the true distribution of mean 1 deviation 0.05.
+We take the gradient of this loss function with respect to the position and flux of all stars, the response in all pixels, and the Zernike coefficients.
+We also use a staged optimisation strategy, initially optimising the positions, fluxes and optical aberrations in order to recover a good \ac{psf} model before learning the \ac{prf}.
 This is necessary to circumvent the high degree of covariance that each individual pixel experiences with the parameters of the \ac{psf}.
-To achieve this the learning rate of the \ac{prf} is set to zero for the first 100 epochs of optimisation.
+To achieve this the learning rate of the \ac{prf} is set to zero for the first 50 epochs of optimisation.
 
-The optimisation is performed for a total of 200 epochs, with each gradient calculation taking \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 Max CPU, leading to a total optimisation time of \textcolor{red}{showyourwork variable} minutes.
+The optimisation is performed for a total of 100 epochs, with each gradient calculation taking \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 Max CPU, leading to a total optimisation time of \textcolor{red}{showyourwork variable} minutes.
 
@@
 This can be seen in Figure~\ref{fig:data_residual} with the right panel showing a sample residual of the data and model after optimisation, which is clearly only noise.
-Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}, and as expected from the residuals all of these parameters are well recovered and tightly constrained around their true values.
+Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}.
+The 1-$sigma$ error bars indicate a poor recovery, however these are only marginalised over positions, fluxes and aberration, not the \ac{prf} which is the dominant noise source.
+These parameters are recovered to the expected distribution when performed without modelling the \ac{prf}.
 Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of \variable{output/rms_opd_resid.txt} nm RMS.
@@
 In our a test case, we initialise from a very naive state far from the truth, requiring a larger number of optimisation epochs than would be required in a real-world example.
-Furthermore as a calibration algorithm across a whole detector as an example the complexity of the model is greatly increased, applying this to real data would likely be focused on small detector regions resulting in less computationally intensive models.
+Furthermore as a calibration algorithm across a whole detector the complexity of the model is greatly increased, applying this to real data would likely be focused on small detector regions resulting in less computationally intensive models.
 
@@
     \centering
+    % \includegraphics[width=0.8\textwidth]{figures/astro_params.pdf}
     \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
@@
 Right: the recovery of the flux parameters in units of photons.
-The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals.}
+The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals.
+The error bars are the 1-$\sigma$ deviations calculated from the covariance Hessian matrix marginalised over positions, fluxes and aberrations.}
     \label{fig:astro_params}
+    % \script{gen_astro_params.py}
 \end{figure*}
@@
 Left: the amplitudes of the Zernike polynomial coefficients, with the top section showing the correlation plot of the true and recovered values, and the bottom showing the individual residuals.
-Middle and right: the true total \ac{opd} and the residuals generated by these values respectively.}
+Middle and right: the true total \ac{opd} and the residuals generated by these values respectively.
+The error bars are the 1-$\sigma$ deviations calculated from the covariance Hessian matrix marginalised over positions, fluxes and aberrations.}
     \label{fig:aberrations}
+    % \script{gen_aberrations.py}
 \end{figure*}
@@
     \caption{This figure shows the PRF values after optimisation.
-The left panel shows the correlation between the pixel-level true and recovered values.
-Each ploint is color-coded with the total flux incident to each pixel, ie its \ac{snr}.
+Left: Correlation between the pixel-level true and recovered values.
+Each point is color-coded with the total flux incident to each pixel, ie its \ac{snr}.
 The pixels with a greater signal are recovered better as expected.
-The right panel shows the histogram of these residuals, with the majority of values being very well recovered and small symmetric about a residual of zero.}
+Right: Histogram of these residuals, with the majority of values being very well recovered and small symmetric about a residual of zero.}
     \label{fig:flat_field}
+    % \script{gen_FF.py}
 \end{figure*}
@@
 
+% poor model
+The use of a forwards model clearly requires a good match between the true and model \ac{psf}.
+The benefit of using differentiable optical models however is that it provides the perfect platform upon which to directly learn this model from the data, with the ability to scale the complexity as required for each problem.
+Similarly since we are working with direct image data as opposed to reduced data-products, miss-calibrations of the system are simple to diagnose through direct examination of the residuals of the data and model.
+Figure~\ref{fig:data_residual} demonstrates that in this example we can be sure of a good fit because the final residual is pure noise.
+
+% \subsection{Limitations}
 This algorithm demonstrates promising performance applied to complex high-dimensional optimisation problems as shown in Section~\ref{sec:phaseretrieval}.
 However, this method does have its limitations.
-The use of a forwards model clearly requires a good match between the true and model \ac{psf}.
-The benefit of using differentiable optical models however is that it provides the perfect platform upon which to directly learn this model from the data, with the ability to scale the complexity as required for each problem.
-Similarly since we are working with direct image data as opposed to reduced data-products is that miss-calibrations of the system are simple to diagnose through direct examination of the residuals of the data and model.
+To test these limits we fit a single star across a wide range of fluxes and look at the recovered positions and fluxes, following the same optimisation strategy as described in Section~\ref{sec:phaseretrieval}.
+Figure~\ref{fig:diverge} shows the recovered positional residuals and fractional flux errors.
+The error bars on the fluxes are 1-$\sigma$ calculated as the square-root of the total flux.
+Clearly using this optical architecture requires more than $10^4$ photons in order to have well recovered parameters.
 
-There are also clear limits in which these methods will begin to fail, such as the low \ac{snr} regime.
-Without the ability to recover the \ac{psf} model accurately, pixel level information will be dominated by the noise processes and the PRF will not be sufficiently encoded.
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.8\textwidth]{figs/divergence.pdf}
+    \caption{Parameter recovery of a single star over varying \ac{snr}. Left: Absolute positional residuals vs total flux. Right: Fractional flux error vs total flux.}
+    \label{fig:diverge}
+    % \script{gen_FF.py}
+\end{figure*}
+
+% Without the ability to recover the \ac{psf} model accurately, pixel level information will be dominated by the noise processes and the PRF will not be sufficiently encoded.
 This can be directly seen in Figure~\ref{fig:flat_field}, with the PRF residual having greater variance in for low-flux pixels.
 This however has a simple solution of observing brighter targets for calibration purposes or observing chosen targets for longer, until a sufficient amount of information is encoded into the data.
-
-It might also seem limited without a dense region of stars to observe as we can not learn about the pixel-level effects in regions with no signal, however we only care about the regions of the detector that have photons landing in them as these are the only ones that will be affected by miss-calibrations in detector level effects.
 
@@
 
-In Paper II in this series, we will tackle the ismorphic problem of hardware design.
+In Paper II in this series, we will tackle the isomorphic problem of hardware design.
 While we have tackled phase design in previous work \cite{phase_ret_and_design}, autodiff permits calculation and optimization of the Fisher information directly \cite{Coe2009}, which means we can optimize towards the fundamental limits of parameter sensitivity.
@@
 \section{Acknowledgements}
+
+We thank Laurent Pueyo and Tim White for their helpful discussions.
 

### 2022-12-12 src/scripts/analyse.py
Adding and updaing scripts

@@ -1,0 +1,8 @@
+import os
+os.system('python plot_progress.py')
+os.system('python plot_astro_params.py')
+os.system('python plot_aberrations.py')
+os.system('python plot_data_resid.py')
+os.system('python plot_FF.py')
+os.system('python plot_optics.py')
+os.system('python plot_divergence.py')

### 2022-12-12 src/scripts/calc_errors.py

@@ -1,0 +1,81 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+# Load model
+tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
+losses = np.load(paths.data / 'losses.npy')
+data = np.load(paths.data / "data.npy")
+
+positions = 'MultiPointSource.position'
+fluxes = 'MultiPointSource.flux'
+zernikes = 'ApplyBasisOPD.coefficients'
+flatfield = 'ApplyPixelResponse.pixel_response'
+parameters = [positions, fluxes, zernikes, flatfield]
+
+def perturb(X, model):
+    """
+    Perturbs the values of the model
+    """
+    model = model.add(positions, X[:2*Nstars].reshape((Nstars, 2)))
+    model = model.add(fluxes, X[2*Nstars:3*Nstars])
+    model = model.add(zernikes, X[3*Nstars:])
+    return model
+
+Nstars = len(tel.get(positions))
+Nzern  = len(tel.get(zernikes))
+X = np.zeros(3*Nstars + Nzern)
+
+from jax import jvp, grad, jit, linearize
+
+def hvp(f, primals, tangents):
+    return jvp(grad(f), primals, tangents)[1]
+
+def hessian(f, x):
+    _, hvp = linearize(jax.grad(f), x)
+    hvp = jit(hvp)  # seems like a substantial speedup to do this
+    basis = np.eye(np.prod(np.array(x.shape))).reshape(-1, *x.shape)
+    return np.stack([hvp(e) for e in basis]).reshape(x.shape + x.shape)
+
+model = tel.set(['observation'], [None])
[…]

### 2022-12-12 src/scripts/data.py

@@ -1,0 +1,96 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+# import pickle as p
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+# Load model
+tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+model = p.load(open(paths.data / 'model.p', 'rb'))
+source = p.load(open(paths.data / 'source.p', 'rb'))
+data = np.load(paths.data / "data.npy")
+
+plt.figure(figsize=(15, 4.5))
+plt.suptitle("Data and Residuals", size=15)
+
+ax1 = plt.subplot(1, 3, 1)
+plt.title("Sample Data")
+plt.imshow(data[0]*1e-3)
+plt.xlabel("Pixels")
+plt.ylabel("Pixels")
+divider = make_axes_locatable(ax1)
+cax = divider.append_axes("right", size="5%", pad=0.1)
+cbar = plt.colorbar(cax=cax)
+cbar.set_label("Counts $x10^3$")
+
+
+initital_psfs = model.observe()
+ax2 = plt.subplot(1, 3, 2)
+plt.title("Initial Residual")
+plt.imshow((initital_psfs[0] - data[0])*1e-3)
+# plt.imshow(data[0]/initital_psfs[0])
+plt.xlabel("Pixels")
+plt.ylabel("Pixels")
+divider = make_axes_locatable(ax2)
+cax = divider.append_axes("right", size="5%", pad=0.1)
+cbar = plt.colorbar(cax=cax)
+cbar.set_label("Counts $x10^3$")
+
+
+final_model = p.load(open(paths.data / 'models_out.p', 'rb'))[-1]
+final_psfs = final_model.observe()
+
+# ax3 = plt.subplot(1, 3, 3)
[…]

### 2022-12-12 src/scripts/divergence.py

@@ -1,0 +1,219 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+# import pickle as p
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+# from tqdm.notebook import tqdm
+from tqdm import tqdm
+
+# %matplotlib inline
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+
+# Define wavelengths
+wavels = 1e-9 * np.linspace(545, 645, 3)
+
+# Basic Optical Parameters
+aperture = 0.5
+wf_npix = 512
+
+# Detector Parameters
+det_npix = 100
+sampling_rate = 3
+det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
+
+# Load mask
+raw_mask = dl.utils.phase_to_opd(np.load("mask.npy"), wavels.mean())
+mask = dl.utils.scale_array(raw_mask, wf_npix, 0)
+
+# Zernike Basis
+zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
+coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), [len(zern_basis)])
+
+# Define Optical Configuration
+optical_layers = [
+    dl.CreateWavefront    (wf_npix, aperture),
+    dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
+    dl.ApplyBasisOPD      (zern_basis, coeffs),
+    dl.AddOPD             (mask),
+    dl.NormaliseWavefront (),
+    dl.AngularMFT         (det_npix, det_pixsize)]
+
+# Create Optics object
+optics = dl.Optics(optical_layers)
+
+# Pixel response
+pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(0), [det_npix, det_npix])
+
+# Create Detector object
[…]

### 2022-12-12 src/scripts/make_model_and_data.py

@@ -28,8 +28,6 @@ plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
 
-
-
 '''Create Optics, Detector, Source & Instrument'''
 # Define wavelengths
 wavels = 1e-9 * np.linspace(545, 645, 3)
@@ -42,6 +40,7 @@ wf_npix = 512
 # Detector Parameters
 det_npix = 1024
 sampling_rate = 20
+
 det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
 
 # Load mask
@@ -65,7 +64,9 @@ optical_layers = [
 optics = dl.Optics(optical_layers)
 
 # Pixel response
-pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(0), [det_npix, det_npix])
+pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(1), [det_npix, det_npix])
+# pix_response = np.ones([det_npix, det_npix])
+
 
 # Detector Noise
 bg_mean = 10
@@ -77,22 +78,30 @@ detector_layers =[dl.ApplyPixelResponse(pix_response),
 # Create Detector object
 detector = dl.Detector(detector_layers)
 
-# Observation stratergy, define dithers
-dithers = 2**-.5 * det_pixsize * np.array([[+1, +1],
-                                           [+1, -1],
-                                           [-1, +1],
-                                           [-1, -1]])
-
-def observe_fn(model, dithers):
-    return model.dither_and_model(dithers)
+# Observation stratergy, nims
+nims = 1
+def observe_fn(model, nims):
+    psf = model.apply('MultiPointSource.flux', lambda x: x/nims).model()
+    return np.tile(psf, (nims, 1, 1))
 
 # Observation dictionary
-observation = {'fn': observe_fn, 'args': dithers}
+observation = {'fn': observe_fn, 'args': nims}
 
 # Multiple sources to observe
-Nstars = 20
-true_positions = a2r(jr.uniform(jr.PRNGKey(0), (Nstars, 2), minval=-5, maxval=5))
-true_fluxes = 1e8 + 1e7*jr.normal(jr.PRNGKey(0), (Nstars,))
+Nstars = 25
+flux = 1e8
+true_fluxes = flux + (flux/10)*jr.normal(jr.PRNGKey(3), (Nstars,))
+
+# Random
+r_max = 4
[…]

### 2022-12-12 src/scripts/optimise.py

@@ -31,7 +31,6 @@ plt.rcParams['figure.dpi'] = 120
 
 # Load model
 model = p.load(open(paths.data / 'model.p', 'rb'))
-# source = p.load(open(paths.data / 'source.p', 'rb'))
 data = np.load(paths.data / "data.npy")
 
 positions = 'MultiPointSource.position'
@@ -40,17 +39,40 @@ zernikes = 'ApplyBasisOPD.coefficients'
 flatfield = 'ApplyPixelResponse.pixel_response'
 parameters = [positions, fluxes, zernikes, flatfield]
 
-# So first we simply set the simple parameters to use an adam optimiser 
-# algorithm, with individual learning rates
-pos_optimiser   = optax.adam(2e-8)
-flux_optimiser  = optax.adam(1e6)
-coeff_optimiser = optax.adam(2e-9)
+# Optimisation hyper parameters
+b1 = .7 # Momentum -> Higer = more momentum
+# b2 = 0.5 # Acceleration -> Higer = more momentum
 
-# Now the flat-field, becuase it is highly covariant with the mean flux level 
-# we don't start learning its parameters until the 100th epoch.
-FF_sched = optax.piecewise_constant_schedule(init_value=1e-2*1e-8, 
-                             boundaries_and_scales={100 : int(1e8)})
-FF_optimiser = optax.adam(FF_sched)
+# Position
+pos_lr, pos_start, pos_stop, pos_restart = 1e-8, 0, 50, 75
+pos_sched = optax.piecewise_constant_schedule(init_value=pos_lr*1e-8,
+                             boundaries_and_scales={pos_start   : int(1e8),
+                                                    pos_stop    : int(1e-8),
+                                                    pos_restart : int(1e8)})
+pos_optimiser   = optax.adam(pos_sched, b1=b1)
+
+# Flux
+flux_lr, flux_start, flux_stop, flux_restart = 2.5e6, 0, 50, 75
+flux_sched = optax.piecewise_constant_schedule(init_value=flux_lr*1e-8,
+                             boundaries_and_scales={flux_start   : int(1e8),
+                                                    flux_stop    : int(1e-8),
+                                                    flux_restart : int(1e8)})
+flux_optimiser = optax.adam(flux_sched, b1=b1)
+
+# Zernikes
+coeff_lr, coeff_start, coeff_stop, coeff_restart = 5e-9, 0, 50, 75
[…]

### 2022-12-12 src/scripts/plot_FF.py

@@ -68,14 +68,64 @@ pix_response = tel.get(flatfield)
 # Plot
 # calculate the mask where there was enough flux to infer the flat field
 # thresh = 2500
-# thresh = 1000
-thresh = 500
+thresh = 1000
+# thresh = 500
+
+data_sort = np.sort(data[0].flatten())
+nhists = 4
+size = len(data_sort)//nhists
+
+threshes = [0]
+for i in range(nhists):
+    threshes.append(data_sort[size*i])
+threshes.append(data_sort[-1] + 1)
+
+print(threshes)
+
+# threshes = np.array([0, 2500, 5000, 7500, 10000, 1e6])
+
+data_flat = data[0].flatten()
+indexes = []
+for i in range(len(threshes)-1):
+    low = np.where(data_flat >= threshes[i])[0]
+    high = np.where(data_flat < threshes[i+1])[0]
+    indexes.append(np.intersect1d(low, high))
+    # intersects.append(len(np.intersect1d(low, high)))
+    # print(len(np.intersect1d(low, high)))
+
+# intersects = np.array(intersects)
+# print()
+# print(intersects.sum())
+# print(len(data_flat))
+
+    # print(threshes[i])
+    # print(np.where(data_flat >= threshes[i]))
+    # print(np.where((data_flat >= threshes[i]) and (data_flat < threshes[i+1])))
+    # print()
+
+prf_flat = (pix_response - flatfields_found[-1]).flatten()
+
+# prf_flat = pix_response.flatten()
+for i in range(len(indexes)):
+    # print(len(indexes[i]))
+    lab = "{} - {}".format(int(threshes[i]), int(threshes[i+1]))
+    plt.hist(prf_flat[indexes[i]], bins=50, alpha=0.85, label=lab)
+plt.legend()
+plt.xlim(-0.2, 0.2)
+plt.savefig(paths.figures / "hists.pdf", dpi=300)
+print("Done hists")
+
+
 fmask = data.mean(0) >= thresh
+# fmask = data >= thresh
 
 out_mask = np.where(data.mean(0) < thresh)
+# out_mask = np.where(data < thresh)
 in_mask = np.where(data.mean(0) >= thresh)
+# in_mask = np.where(data >= thresh)
 
[…]

### 2022-12-12 src/scripts/plot_aberrations.py

@@ -81,6 +81,11 @@ true_coeff = tel.get(zernikes)
 found_coeff = models_out[-1].get(zernikes)
 index = np.arange(len(true_coeff))+4
 
+Nzern = len(true_coeff)
+errs = np.diag(np.load('cov_mat.npy'))**0.5
+zerr = errs[-Nzern:]
+# print(zerr)
+
 plt.figure(figsize=(15, 4))
 plt.suptitle("Optical Aberrations", size=15)
 
@@ -95,7 +100,8 @@ fig1.add_axes(ax2)
 
 
 plt.title("Zernike Coefficients (nm)")
-ax2.scatter(true_coeff*1e9, found_coeff*1e9)
+# ax2.scatter(true_coeff*1e9, found_coeff*1e9)
+ax2.errorbar(true_coeff*1e9, found_coeff*1e9, yerr=zerr*1e9, fmt='o', capsize=5)
 
 ax2.set_xticks([])
 ax2.set_ylabel("Recovered")
@@ -112,7 +118,8 @@ ax2.plot(vals, vals, c='k', alpha=0.5)
 ax2.set_xlim(xlims)
 ax2.set_ylim(ylims)
 
-ax.scatter(true_coeff*1e9, (true_coeff - found_coeff)*1e9)
+# ax.scatter(true_coeff*1e9, (true_coeff - found_coeff)*1e9)
+ax.errorbar(true_coeff*1e9, (true_coeff - found_coeff)*1e9, yerr=zerr*1e9, fmt='o', capsize=5)
 # ax.scatter(true_coeff, (true_coeff - found_coeff)/true_coeff)
 ax.axhline(0, c='k')
 ax.set_ylabel("Residuals")
@@ -167,3 +174,4 @@ cbar.set_label("OPD (nm)")
 
 plt.tight_layout()
 plt.savefig(paths.figures / "aberrations.pdf", dpi=300)
+

### 2022-12-12 src/scripts/plot_astro_params.py

@@ -31,13 +31,9 @@ plt.rcParams['figure.dpi'] = 120
 
 # Load model
 tel = p.load(open(paths.data / 'instrument.p', 'rb'))
-# models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
-# losses = np.load(paths.data / 'losses.npy')
-# data = np.load(paths.data / "data.npy")
-models_out = p.load(open(paths.data / 'models_out_.p', 'rb'))
-losses = np.load(paths.data / 'losses_.npy')
+models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
+losses = np.load(paths.data / 'losses.npy')
 data = np.load(paths.data / "data.npy")
-
 
 positions = 'MultiPointSource.position'
 fluxes = 'MultiPointSource.flux'
@@ -66,13 +62,16 @@ positions_residuals = tel.get(positions) - positions_found
 r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
 r_residuals = r2a(r_residuals_rads)
 
+# Errors
+Nstars = len(tel.get(positions))
+errs = np.diag(np.load('cov_mat.npy'))**0.5
 
 # Positions
 arcsec = r2a(1)
 true_positions = tel.get(positions).flatten() * arcsec
 initial_positions = positions_found[0].flatten() * arcsec
 final_positions = positions_found[-1].flatten() * arcsec
-
+pos_err = errs[:2*Nstars] * arcsec
 
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 fig = plt.figure(figsize=(10, 4))
@@ -85,7 +84,9 @@ ax2 = divider.new_vertical(size="150%", pad=0.00)
 fig.add_axes(ax2)
 ax2.set_title("Positions (arcseconds)")
 
-ax2.scatter(true_positions, final_positions, label='final')
+ax2.scatter(true_positions, final_positions)
+# ax2.errorbar(true_positions, final_positions, yerr=pos_err, fmt='o', capsize=5)
+
 # ax2.scatter(true_positions, initial_positions, c='tab:orange', marker='x', label="intitial")
 
 xlims = ax2.get_xlim()
@@ -98,7 +99,9 @@ ax2.set_xticks([])
[…]

### 2022-12-12 src/scripts/plot_data_resid.py

@@ -31,46 +31,9 @@ plt.rcParams['figure.dpi'] = 120
 
 # Load model
 tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+model = p.load(open(paths.data / 'model.p', 'rb'))
 source = p.load(open(paths.data / 'source.p', 'rb'))
-
-# Define Params
-positions = 'MultiPointSource.position'
-fluxes = 'MultiPointSource.flux'
-zernikes = 'ApplyBasisOPD.coefficients'
-flatfield = 'ApplyPixelResponse.pixel_response'
-parameters = [positions, fluxes, zernikes, flatfield]
-
-det_pixsize = tel.get("AngularMFT.pixel_scale_out")
-det_npix = tel.get("AngularMFT.npixels_out")
-nzern = len(tel.get(zernikes))
-Nstars = len(source.get('flux'))
-
-# Add small random values to the positions
-model = tel.add(positions, 2.5*det_pixsize*jr.normal(jr.PRNGKey(0),  (Nstars, 2)))
-
-# Multiply the fluxes by small random values
-model = model.multiply(fluxes, 1 + 0.1*jr.normal(jr.PRNGKey(0), (Nstars,)))
-
-# Set the zernike coefficients to zero
-model = model.set(zernikes, np.zeros(nzern))
-
-# Set the flat fiel to uniform
-model = model.set(flatfield, np.ones((det_npix, det_npix)))
-
-
-# Save optimisation model
-p.dump(model, open(paths.data / 'model.p', 'wb'))
-
-# Observe!
-bg_val = tel.AddConstant.value
-psfs = tel.observe() - bg_val
-print(psfs.sum(0).mean())
-
-
-# Apply some noise to the PSF Background noise
-BG_noise = 2.5*jr.normal(jr.PRNGKey(0), psfs.shape) + bg_val
-data = jr.poisson(jr.PRNGKey(0), psfs) + BG_noise
-np.save(paths.data / "data", data)
+data = np.load(paths.data / "data.npy")
 
 plt.figure(figsize=(15, 4.5))
 plt.suptitle("Data and Residuals", size=15)
@@ -78,6 +41,7 @@ plt.suptitle("Data and Residuals", size=15)
 ax1 = plt.subplot(1, 3, 1)
 plt.title("Sample Data")
[…]

### 2022-12-12 src/scripts/plot_divergence.py

@@ -1,0 +1,83 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+# %matplotlib inline
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+positions = 'PointSource.position'
+fluxes = 'PointSource.flux'
+zernikes = 'ApplyBasisOPD.coefficients'
+flatfield = 'ApplyPixelResponse.pixel_response'
+
+fluxes_in = np.load(paths.data / "divergence_fluxes_in.npy")
+models_out = p.load(open(paths.data / 'divergence_models_out.p', 'rb'))
+
+# Positions
+positions_found = np.array([model.get(positions) for model in models_out])
+x_res, y_res = (np.zeros(2)- positions_found[-1]).T
+r_res = np.hypot(x_res, y_res)
+
+# Fluxes
+final_fluxes = models_out[-1].get(fluxes)
+res = np.abs(fluxes_in - final_fluxes)
+l1 = res/fluxes_in
+xs = fluxes_in
+diffs = (models_out[-1].get(fluxes) - models_out[-2].get(fluxes))/fluxes_in
+
+sigma = 1
+
+plt.figure(figsize=(10, 4))
+
+# plt.subplot(1, 2, 1)
+# plt.title("Final Flux Error")
+# plt.scatter(fluxes_in, res)
+# plt.errorbar(fluxes_in, res, yerr=sigma*fluxes_in**0.5, fmt='o')
+# plt.semilogx()
+# plt.axhline(0, c='k')
+# plt.xlabel("True")
+# plt.ylabel("Recovered")
+plt.subplot(1, 2, 1)
+plt.title("Positional Residual")
+plt.scatter(fluxes_in, r2a(r_res))
+plt.semilogx()
+plt.axhline(0, c='k')
[…]

### 2022-12-12 src/scripts/plot_optics.py

@@ -44,10 +44,12 @@ aberrated_pupil = pupil + opd
 # FF
 FF = tel.ApplyPixelResponse.pixel_response
 
+psf_tel = tel.set(['detector'], [None])
+
 # Get psfs
 zernikes = 'ApplyBasisOPD.coefficients'
-psf = tel.set(zernikes, np.zeros(len(tel.get(zernikes)))).model(source=dl.PointSource(wavelengths=wavels))
-aberrated_psf = tel.model(source=dl.PointSource(wavelengths=wavels))
+psf = psf_tel.set(zernikes, np.zeros(len(tel.get(zernikes)))).model(source=dl.PointSource(wavelengths=wavels))
+aberrated_psf = psf_tel.model(source=dl.PointSource(wavelengths=wavels))
 
 cmap = get_cmap("inferno").copy()
 cmap.set_bad('k',1.)

### 2022-12-12 src/scripts/plot_progress.py

@@ -1,0 +1,129 @@
+# Core jax
+import jax
+import jax.numpy as np
+import jax.random as jr
+
+# Optimisation
+import equinox as eqx
+import optax
+
+# Optics
+import dLux as dl
+from dLux.utils import arcseconds_to_radians as a2r
+from dLux.utils import radians_to_arcseconds as r2a
+
+# Paths
+import paths
+
+# Pickle
+# import pickle as p
+import dill as p
+
+# Plotting/visualisation
+import matplotlib.pyplot as plt
+from tqdm import tqdm
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+
+# Load model
+tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
+losses = np.load(paths.data / 'losses.npy')
+data = np.load(paths.data / "data.npy")
+
+positions = 'MultiPointSource.position'
+fluxes = 'MultiPointSource.flux'
+zernikes = 'ApplyBasisOPD.coefficients'
+flatfield = 'ApplyPixelResponse.pixel_response'
+parameters = [positions, fluxes, zernikes, flatfield]
+
+
+
+# Get parameters
+nepochs = len(models_out)
+psfs_out = models_out[-1].observe()
+
+positions_found  = np.array([model.get(positions) for model in models_out])
+fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
+zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
+flatfields_found = np.array([model.get(flatfield) for model in models_out])
+
+
+# Get the residuals
+coeff_residuals     = tel.get(zernikes)  - zernikes_found
+flux_residuals      = tel.get(fluxes)    - fluxes_found
+flatfield_residuals = tel.get(flatfield) - flatfields_found
+
+scaler = 1e3
+positions_residuals = tel.get(positions) - positions_found
+r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
[…]

### 2022-12-12 src/scripts/run.py

@@ -1,0 +1,5 @@
+import os
+os.system('python make_model_and_data.py')
+os.system('python optimise.py')
+os.system('python calc_error.py')
+os.system('python divergence.py')

### 2022-12-12 src/scripts/analyse.py
Updating scripts

@@ -1,8 +1,37 @@
 import os
+import time as t
+
+t0 = t.time()
+print("\nRunning plot_progress.py")
 os.system('python plot_progress.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))
+
+t0 = t.time()
+print("\nRunning plot_astro_params.py")
 os.system('python plot_astro_params.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))
+
+t0 = t.time()
+print("\nRunning plot_aberrations.py")
 os.system('python plot_aberrations.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))
+
+t0 = t.time()
+print("\nRunning plot_data_resid.py")
 os.system('python plot_data_resid.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))
+
+t0 = t.time()
+print("\nRunning plot_FF.py")
 os.system('python plot_FF.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))
+
+t0 = t.time()
+print("\nRunning plot_optics.py")
 os.system('python plot_optics.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))
+
+t0 = t.time()
+print("\nRunning plot_divergence.py")
 os.system('python plot_divergence.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))

### 2022-12-12 src/scripts/calc_errors.py

@@ -1,32 +1,10 @@
-# Core jax
-import jax
 import jax.numpy as np
 import jax.random as jr
-
-# Optimisation
-import equinox as eqx
-import optax
-
-# Optics
+from jax import jvp, grad, jit, linearize
 import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
-from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
 import paths
-
-# Pickle
 import dill as p
-
-# Plotting/visualisation
-import matplotlib.pyplot as plt
 from tqdm import tqdm
-
-plt.rcParams['image.cmap'] = 'inferno'
-plt.rcParams["font.family"] = "serif"
-plt.rcParams["image.origin"] = 'lower'
-plt.rcParams['figure.dpi'] = 120
-
 
 # Load model
 tel = p.load(open(paths.data / 'instrument.p', 'rb'))
@@ -53,24 +31,24 @@ Nstars = len(tel.get(positions))
 Nzern  = len(tel.get(zernikes))
 X = np.zeros(3*Nstars + Nzern)
 
-from jax import jvp, grad, jit, linearize
 
 def hvp(f, primals, tangents):
     return jvp(grad(f), primals, tangents)[1]
 
 def hessian(f, x):
-    _, hvp = linearize(jax.grad(f), x)
+    _, hvp = linearize(grad(f), x)
     hvp = jit(hvp)  # seems like a substantial speedup to do this
     basis = np.eye(np.prod(np.array(x.shape))).reshape(-1, *x.shape)
     return np.stack([hvp(e) for e in basis]).reshape(x.shape + x.shape)
 
-model = tel.set(['observation'], [None])
-fn = lambda X: dl.utils.poisson_log_likelihood(X, data[0], model, perturb, 'model')
+# model = tel.set(['observation'], [None])
+model = tel
+fn = lambda X: dl.utils.poisson_log_likelihood(X, data, model, perturb, 'model')
 
 H = hessian(fn, X)
 cov_mat = -np.linalg.inv(H)
 
-np.save('cov_mat', cov_mat)
+np.save(paths.data / 'cov_mat', cov_mat)
 
 
 

### 2022-12-12 src/scripts/divergence.py

@@ -1,36 +1,12 @@
-# Core jax
-import jax
 import jax.numpy as np
 import jax.random as jr
-
-# Optimisation
+from jax.scipy.stats import poisson
 import equinox as eqx
 import optax
-
-# Optics
 import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
-from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
 import paths
-
-# Pickle
-# import pickle as p
 import dill as p
-
-# Plotting/visualisation
-import matplotlib.pyplot as plt
-# from tqdm.notebook import tqdm
 from tqdm import tqdm
-
-# %matplotlib inline
-plt.rcParams['image.cmap'] = 'inferno'
-plt.rcParams["font.family"] = "serif"
-plt.rcParams["image.origin"] = 'lower'
-plt.rcParams['figure.dpi'] = 120
-
-
 
 # Define wavelengths
 wavels = 1e-9 * np.linspace(545, 645, 3)
@@ -113,12 +89,11 @@ def make_images(model, key):
 
 @eqx.filter_vmap
 def initialise_models(model):
-    # model = model.add(positions, 1.*det_pixsize*jr.normal(jr.PRNGKey(0),  (2,)))
-    model = model.add(positions, 1.*det_pixsize)
+    model = model.add(positions, 1.*det_pixsize*jr.normal(jr.PRNGKey(0),  (2,)))
+    # model = model.add(positions, 1.*det_pixsize)
 
     # Multiply the fluxes by small random values
-    # model = model.multiply(fluxes, 1 + 0.1*jr.normal(jr.PRNGKey(1), (1,)))
-    model = model.multiply(fluxes, 1.1)
+    model = model.multiply(fluxes, 1 + 0.1*jr.normal(jr.PRNGKey(1), (1,)))
 
     # Set the zernike coefficients to zero
     model = model.set(zernikes, np.zeros(len(zern_basis)))
@@ -134,13 +109,12 @@ flux_ratios = fluxes_in/fluxes_in[0]
 tels = make_instruments(tel, fluxes_in)
 psfs = evaluate_ensemble(tels)
 data = make_images(tels, np.ones(len(fluxes_in), dtype=int))
-
-
-
 models = initialise_models(tels)
 
+
+
 # Optimisation hyper parameters
-b1 = .7 # Momentum -> Higer = more momentum
[…]

### 2022-12-12 src/scripts/make_model_and_data.py

@@ -1,32 +1,11 @@
 # Core jax
-import jax
 import jax.numpy as np
 import jax.random as jr
-
-# Optimisation
 import equinox as eqx
-import optax
-
-# Optics
 import dLux as dl
 from dLux.utils import arcseconds_to_radians as a2r
-from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
-import paths
-
-# Pickle
-# import pickle as p
 import dill as p
-
-# Plotting/visualisation
-import matplotlib.pyplot as plt
-
-plt.rcParams['image.cmap'] = 'inferno'
-plt.rcParams["font.family"] = "serif"
-plt.rcParams["image.origin"] = 'lower'
-plt.rcParams['figure.dpi'] = 120
-
+import paths
 
 '''Create Optics, Detector, Source & Instrument'''
 # Define wavelengths
@@ -40,7 +19,6 @@ wf_npix = 512
 # Detector Parameters
 det_npix = 1024
 sampling_rate = 20
-
 det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
 
 # Load mask
@@ -67,7 +45,6 @@ optics = dl.Optics(optical_layers)
 pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(1), [det_npix, det_npix])
 # pix_response = np.ones([det_npix, det_npix])
 
-
 # Detector Noise
 bg_mean = 10
 
@@ -78,23 +55,14 @@ detector_layers =[dl.ApplyPixelResponse(pix_response),
 # Create Detector object
 detector = dl.Detector(detector_layers)
 
-# Observation stratergy, nims
-nims = 1
-def observe_fn(model, nims):
-    psf = model.apply('MultiPointSource.flux', lambda x: x/nims).model()
-    return np.tile(psf, (nims, 1, 1))
-
-# Observation dictionary
-observation = {'fn': observe_fn, 'args': nims}
-
 # Multiple sources to observe
-Nstars = 25
+Nstars = 20
 flux = 1e8
-true_fluxes = flux + (flux/10)*jr.normal(jr.PRNGKey(3), (Nstars,))
+true_fluxes = flux + (flux/10)*jr.normal(jr.PRNGKey(2), (Nstars,))
 
 # Random
-r_max = 4
-true_positions = a2r(jr.uniform(jr.PRNGKey(2), (Nstars, 2), minval=-r_max, maxval=r_max))
+r_max = 4.5
[…]

### 2022-12-12 src/scripts/optimise.py

@@ -1,33 +1,11 @@
-# Core jax
-import jax
 import jax.numpy as np
 import jax.random as jr
-
-# Optimisation
+from jax.scipy.stats import poisson
 import equinox as eqx
 import optax
-
-# Optics
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
-from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
 import paths
-
-# Pickle
-# import pickle as p
 import dill as p
-
-# Plotting/visualisation
-import matplotlib.pyplot as plt
 from tqdm import tqdm
-
-plt.rcParams['image.cmap'] = 'inferno'
-plt.rcParams["font.family"] = "serif"
-plt.rcParams["image.origin"] = 'lower'
-plt.rcParams['figure.dpi'] = 120
-
 
 # Load model
 model = p.load(open(paths.data / 'model.p', 'rb'))
@@ -40,7 +18,7 @@ flatfield = 'ApplyPixelResponse.pixel_response'
 parameters = [positions, fluxes, zernikes, flatfield]
 
 # Optimisation hyper parameters
-b1 = .7 # Momentum -> Higer = more momentum
+b1 = .75 # Momentum -> Higer = more momentum
 # b2 = 0.5 # Acceleration -> Higer = more momentum
 
 # Position
@@ -81,8 +59,8 @@ optimisers = [pos_optimiser, flux_optimiser, coeff_optimiser, FF_optimiser]
 optim, opt_state, args = model.get_optimiser(parameters, optimisers, get_args=True)
 
 def log_like(model, data):
-    psfs = np.maximum(model.observe(), 1e-8)
-    return -np.sum(jax.scipy.stats.poisson.logpmf(data, psfs))
+    psfs = np.maximum(model.model(), 1e-8)
+    return -np.sum(poisson.logpmf(data, psfs))
 
 def log_prior(model, ff_mean=1., ff_std=0.05):
     return 0.5*(np.square((ff_mean - model.get(flatfield))/ff_std)).sum()
@@ -92,13 +70,10 @@ def log_prior(model, ff_mean=1., ff_std=0.05):
 def loss_fn(model, data):
     return log_prior(model) + log_like(model, data)
 
-# %%time
-loss, grads = loss_fn(model, data) # Compile
-print(loss)
-print("Initial Loss: {}".format(int(loss)))
-print("Initial Loss: {}".format(np.log10(loss)))
+# Compile
[…]

### 2022-12-12 src/scripts/plot_FF.py

@@ -1,27 +1,8 @@
-# Core jax
-import jax
 import jax.numpy as np
-import jax.random as jr
-
-# Optimisation
-import equinox as eqx
-import optax
-
-# Optics
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
 from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
 import paths
-
-# Pickle
-# import pickle as p
 import dill as p
-
-# Plotting/visualisation
 import matplotlib.pyplot as plt
-from tqdm import tqdm
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
@@ -34,6 +15,7 @@ tel = p.load(open(paths.data / 'instrument.p', 'rb'))
 models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
 losses = np.load(paths.data / 'losses.npy')
 data = np.load(paths.data / "data.npy")
+psfs_out = np.load(paths.data / "final_psfs.npy")
 
 positions = 'MultiPointSource.position'
 fluxes = 'MultiPointSource.flux'
@@ -43,9 +25,10 @@ parameters = [positions, fluxes, zernikes, flatfield]
 
 
 
-# Get parameters
-nepochs = len(models_out)
-psfs_out = models_out[-1].observe()
+# # Get parameters
+# nepochs = len(models_out)
+# # psfs_out = models_out[-1].observe()
+# psfs_out = models_out[-1].model()
 
 positions_found  = np.array([model.get(positions) for model in models_out])
 fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
@@ -71,7 +54,8 @@ pix_response = tel.get(flatfield)
 thresh = 1000
 # thresh = 500
 
-data_sort = np.sort(data[0].flatten())
+# data_sort = np.sort(data[0].flatten())
+data_sort = np.sort(data.flatten())
 nhists = 4
 size = len(data_sort)//nhists
 
@@ -80,11 +64,12 @@ for i in range(nhists):
     threshes.append(data_sort[size*i])
 threshes.append(data_sort[-1] + 1)
 
-print(threshes)
+# print(threshes)
 
 # threshes = np.array([0, 2500, 5000, 7500, 10000, 1e6])
 
-data_flat = data[0].flatten()
+# data_flat = data[0].flatten()
[…]

### 2022-12-12 src/scripts/plot_aberrations.py

@@ -1,27 +1,8 @@
-# Core jax
-import jax
 import jax.numpy as np
-import jax.random as jr
-
-# Optimisation
-import equinox as eqx
-import optax
-
-# Optics
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
 from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
 import paths
-
-# Pickle
-# import pickle as p
 import dill as p
-
-# Plotting/visualisation
 import matplotlib.pyplot as plt
-from tqdm import tqdm
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
@@ -34,6 +15,7 @@ tel = p.load(open(paths.data / 'instrument.p', 'rb'))
 models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
 losses = np.load(paths.data / 'losses.npy')
 data = np.load(paths.data / "data.npy")
+psfs_out = np.load(paths.data / "final_psfs.npy")
 
 positions = 'MultiPointSource.position'
 fluxes = 'MultiPointSource.flux'
@@ -43,9 +25,10 @@ parameters = [positions, fluxes, zernikes, flatfield]
 
 
 
-# Get parameters
-nepochs = len(models_out)
-psfs_out = models_out[-1].observe()
+# # Get parameters
+# nepochs = len(models_out)
+# # psfs_out = models_out[-1].observe()
+# psfs_out = models_out[-1].model()
 
 positions_found  = np.array([model.get(positions) for model in models_out])
 fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
@@ -82,7 +65,7 @@ found_coeff = models_out[-1].get(zernikes)
 index = np.arange(len(true_coeff))+4
 
 Nzern = len(true_coeff)
-errs = np.diag(np.load('cov_mat.npy'))**0.5
+errs = np.diag(np.load(paths.data / 'cov_mat.npy'))**0.5
 zerr = errs[-Nzern:]
 # print(zerr)
 

### 2022-12-12 src/scripts/plot_astro_params.py

@@ -1,27 +1,8 @@
-# Core jax
-import jax
 import jax.numpy as np
-import jax.random as jr
-
-# Optimisation
-import equinox as eqx
-import optax
-
-# Optics
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
 from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
 import paths
-
-# Pickle
-# import pickle as p
 import dill as p
-
-# Plotting/visualisation
 import matplotlib.pyplot as plt
-from tqdm import tqdm
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
@@ -34,6 +15,7 @@ tel = p.load(open(paths.data / 'instrument.p', 'rb'))
 models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
 losses = np.load(paths.data / 'losses.npy')
 data = np.load(paths.data / "data.npy")
+psfs_out = np.load(paths.data / "final_psfs.npy")
 
 positions = 'MultiPointSource.position'
 fluxes = 'MultiPointSource.flux'
@@ -44,8 +26,9 @@ parameters = [positions, fluxes, zernikes, flatfield]
 
 
 # Get parameters
-nepochs = len(models_out)
-psfs_out = models_out[-1].observe()
+# nepochs = len(models_out)
+# psfs_out = models_out[-1].observe()
+# psfs_out = models_out[-1].model()
 
 positions_found  = np.array([model.get(positions) for model in models_out])
 fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
@@ -64,7 +47,7 @@ r_residuals = r2a(r_residuals_rads)
 
 # Errors
 Nstars = len(tel.get(positions))
-errs = np.diag(np.load('cov_mat.npy'))**0.5
+errs = np.diag(np.load(paths.data / 'cov_mat.npy'))**0.5
 
 # Positions
 arcsec = r2a(1)

### 2022-12-12 src/scripts/plot_data_resid.py

@@ -1,25 +1,6 @@
-# Core jax
-import jax
 import jax.numpy as np
-import jax.random as jr
-
-# Optimisation
-import equinox as eqx
-import optax
-
-# Optics
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
-from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
 import paths
-
-# Pickle
-# import pickle as p
 import dill as p
-
-# Plotting/visualisation
 import matplotlib.pyplot as plt
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
@@ -34,13 +15,16 @@ tel = p.load(open(paths.data / 'instrument.p', 'rb'))
 model = p.load(open(paths.data / 'model.p', 'rb'))
 source = p.load(open(paths.data / 'source.p', 'rb'))
 data = np.load(paths.data / "data.npy")
+final_psfs = np.load(paths.data / "final_psfs.npy")
+initital_psfs = np.load(paths.data / "initial_psfs.npy")
 
 plt.figure(figsize=(15, 4.5))
 plt.suptitle("Data and Residuals", size=15)
 
 ax1 = plt.subplot(1, 3, 1)
 plt.title("Sample Data")
-plt.imshow(data[0]*1e-3)
+# plt.imshow(data[0]*1e-3)
+plt.imshow(data*1e-3)
 # plt.imshow(data*1e-3)
 plt.xlabel("Pixels")
 plt.ylabel("Pixels")
@@ -50,10 +34,12 @@ cbar = plt.colorbar(cax=cax)
 cbar.set_label("Counts $x10^3$")
 
 
-initital_psfs = model.observe()
+# initital_psfs = model.observe()
+# initital_psfs = model.model()
 ax2 = plt.subplot(1, 3, 2)
 plt.title("Initial Residual")
-plt.imshow((initital_psfs[0] - data[0])*1e-3)
+# plt.imshow((initital_psfs[0] - data[0])*1e-3)
+plt.imshow((initital_psfs - data)*1e-3)
 # plt.imshow((initital_psfs[0] - data)*1e-3)
 # plt.imshow(data[0]/initital_psfs[0])
 plt.xlabel("Pixels")
@@ -64,13 +50,15 @@ cbar = plt.colorbar(cax=cax)
 cbar.set_label("Counts $x10^3$")
 
 
[…]

### 2022-12-12 src/scripts/plot_divergence.py

@@ -1,28 +1,9 @@
-# Core jax
-import jax
 import jax.numpy as np
-import jax.random as jr
-
-# Optimisation
-import equinox as eqx
-import optax
-
-# Optics
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
 from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
 import paths
-
-# Pickle
 import dill as p
-
-# Plotting/visualisation
 import matplotlib.pyplot as plt
-from tqdm import tqdm
 
-# %matplotlib inline
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'

### 2022-12-12 src/scripts/plot_optics.py

@@ -1,25 +1,7 @@
-# Core jax
-import jax
 import jax.numpy as np
-import jax.random as jr
-
-# Optimisation
-import equinox as eqx
-import optax
-
-# Optics
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
-from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
 import paths
-
-# Pickle
-# import pickle as p
 import dill as p
-
-# Plotting/visualisation
+import dLux as dl
 import matplotlib.pyplot as plt
 from matplotlib.cm import get_cmap
 

### 2022-12-12 src/scripts/plot_progress.py

@@ -1,27 +1,8 @@
-# Core jax
-import jax
 import jax.numpy as np
-import jax.random as jr
-
-# Optimisation
-import equinox as eqx
-import optax
-
-# Optics
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
 from dLux.utils import radians_to_arcseconds as r2a
-
-# Paths
 import paths
-
-# Pickle
-# import pickle as p
 import dill as p
-
-# Plotting/visualisation
 import matplotlib.pyplot as plt
-from tqdm import tqdm
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
@@ -34,6 +15,7 @@ tel = p.load(open(paths.data / 'instrument.p', 'rb'))
 models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
 losses = np.load(paths.data / 'losses.npy')
 data = np.load(paths.data / "data.npy")
+psfs_out = np.load(paths.data / "final_psfs.npy")
 
 positions = 'MultiPointSource.position'
 fluxes = 'MultiPointSource.flux'
@@ -41,17 +23,16 @@ zernikes = 'ApplyBasisOPD.coefficients'
 flatfield = 'ApplyPixelResponse.pixel_response'
 parameters = [positions, fluxes, zernikes, flatfield]
 
-
-
 # Get parameters
-nepochs = len(models_out)
-psfs_out = models_out[-1].observe()
+# nepochs = len(models_out)
+# psfs_out = models_out[-1].observe()
+# psfs_out = models_out[-1].model()
+
 
 positions_found  = np.array([model.get(positions) for model in models_out])
 fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
 zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
 flatfields_found = np.array([model.get(flatfield) for model in models_out])
-
 
 # Get the residuals
 coeff_residuals     = tel.get(zernikes)  - zernikes_found
@@ -107,7 +88,8 @@ plt.plot(np.square(flatfield_residuals).sum((-1, -2)))
 plt.axhline(0, c='k', alpha=0.5)
 
 # FF Scatter Plot
-colours = data.sum(0).flatten()
+# colours = data.sum(0).flatten()
+colours = data.flatten()
 # colours = data.flatten()
 ind = np.argsort(colours)
 colours = colours[ind]

### 2022-12-12 src/scripts/run.py

@@ -1,5 +1,22 @@
 import os
+import time as t
+
+t0 = t.time()
+print("\nRunning make_model_and_data.py")
 os.system('python make_model_and_data.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))
+
+t0 = t.time()
+print("\nRunning optimise.py")
 os.system('python optimise.py')
-os.system('python calc_error.py')
-os.system('python divergence.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))
+
+# t0 = t.time()
+# print("\nRunning calc_errors.py")
+# os.system('python calc_errors.py')
+# print("Done in {:.2f} seconds".format(t.time() - t0))
+
+# t0 = t.time()
+# print("\nRunning divergence.py")
+# os.system('python divergence.py')
+# print("Done in {:.2f} seconds".format(t.time() - t0))

### 2022-12-12 src/scripts/divergence.py
Updating scripts

@@ -127,7 +127,7 @@ pos_sched = optax.piecewise_constant_schedule(init_value=pos_lr*1e-8,
 pos_optimiser   = optax.adam(pos_sched, b1=b1)
 
 # Flux
-flux_lr = 0.5e3*flux_ratios * np.ones(models.get(fluxes).shape)
+flux_lr = 0.5e3*flux_ratios[:, None] * np.ones(models.get(fluxes).shape)
 flux_start, flux_stop, flux_restart = 0, 50, 75
 flux_sched = optax.piecewise_constant_schedule(init_value=flux_lr*1e-8,
                              boundaries_and_scales={flux_start   : int(1e8),
@@ -168,6 +168,7 @@ def loss_fn(model, data):
 # Compile
 losses, grads = loss_fn(models, data) 
 
+# Optimise
 losses_out, models_out = [], []
 with tqdm(range(100),desc='Gradient Descent') as t:
     for i in t:

### 2022-12-12 src/scripts/plot_FF.py

@@ -9,7 +9,6 @@ plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
-
 # Load model
 tel = p.load(open(paths.data / 'instrument.p', 'rb'))
 models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
@@ -23,18 +22,10 @@ zernikes = 'ApplyBasisOPD.coefficients'
 flatfield = 'ApplyPixelResponse.pixel_response'
 parameters = [positions, fluxes, zernikes, flatfield]
 
-
-
-# # Get parameters
-# nepochs = len(models_out)
-# # psfs_out = models_out[-1].observe()
-# psfs_out = models_out[-1].model()
-
 positions_found  = np.array([model.get(positions) for model in models_out])
 fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
 zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
 flatfields_found = np.array([model.get(flatfield) for model in models_out])
-
 
 # Get the residuals
 coeff_residuals = tel.get(zernikes) - zernikes_found
@@ -44,17 +35,11 @@ scaler = 1e3
 positions_residuals = tel.get(positions) - positions_found
 r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
 r_residuals = r2a(r_residuals_rads)
-
-
 pix_response = tel.get(flatfield)
 
 # Plot
 # calculate the mask where there was enough flux to infer the flat field
-# thresh = 2500
 thresh = 1000
-# thresh = 500
-
-# data_sort = np.sort(data[0].flatten())
 data_sort = np.sort(data.flatten())
 nhists = 4
 size = len(data_sort)//nhists
@@ -64,57 +49,27 @@ for i in range(nhists):
     threshes.append(data_sort[size*i])
 threshes.append(data_sort[-1] + 1)
 
-# print(threshes)
-
-# threshes = np.array([0, 2500, 5000, 7500, 10000, 1e6])
-
-# data_flat = data[0].flatten()
 data_flat = data.flatten()
 indexes = []
 for i in range(len(threshes)-1):
     low = np.where(data_flat >= threshes[i])[0]
[…]

### 2022-12-12 src/scripts/plot_aberrations.py

@@ -3,6 +3,7 @@ from dLux.utils import radians_to_arcseconds as r2a
 import paths
 import dill as p
 import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import make_axes_locatable
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
@@ -23,18 +24,11 @@ zernikes = 'ApplyBasisOPD.coefficients'
 flatfield = 'ApplyPixelResponse.pixel_response'
 parameters = [positions, fluxes, zernikes, flatfield]
 
-
-
 # # Get parameters
-# nepochs = len(models_out)
-# # psfs_out = models_out[-1].observe()
-# psfs_out = models_out[-1].model()
-
 positions_found  = np.array([model.get(positions) for model in models_out])
 fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
 zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
 flatfields_found = np.array([model.get(flatfield) for model in models_out])
-
 
 # Get the residuals
 coeff_residuals = tel.get(zernikes) - zernikes_found
@@ -44,9 +38,6 @@ scaler = 1e3
 positions_residuals = tel.get(positions) - positions_found
 r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
 r_residuals = r2a(r_residuals_rads)
-
-
-
 
 # Plot
 # OPDs
@@ -67,13 +58,10 @@ index = np.arange(len(true_coeff))+4
 Nzern = len(true_coeff)
 errs = np.diag(np.load(paths.data / 'cov_mat.npy'))**0.5
 zerr = errs[-Nzern:]
-# print(zerr)
 
 plt.figure(figsize=(15, 4))
 plt.suptitle("Optical Aberrations", size=15)
 
-
-from mpl_toolkits.axes_grid1 import make_axes_locatable
 ax = plt.subplot(1, 3, 1)
 divider = make_axes_locatable(ax)
 
@@ -81,36 +69,22 @@ ax2 = divider.new_vertical(size="150%", pad=0.00)
[…]

### 2022-12-12 src/scripts/plot_astro_params.py

@@ -9,7 +9,6 @@ plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
-
 # Load model
 tel = p.load(open(paths.data / 'instrument.p', 'rb'))
 models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
@@ -23,18 +22,11 @@ zernikes = 'ApplyBasisOPD.coefficients'
 flatfield = 'ApplyPixelResponse.pixel_response'
 parameters = [positions, fluxes, zernikes, flatfield]
 
-
-
 # Get parameters
-# nepochs = len(models_out)
-# psfs_out = models_out[-1].observe()
-# psfs_out = models_out[-1].model()
-
 positions_found  = np.array([model.get(positions) for model in models_out])
 fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
 zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
 flatfields_found = np.array([model.get(flatfield) for model in models_out])
-
 
 # Get the residuals
 coeff_residuals = tel.get(zernikes) - zernikes_found
@@ -66,11 +58,7 @@ divider = make_axes_locatable(ax1)
 ax2 = divider.new_vertical(size="150%", pad=0.00)
 fig.add_axes(ax2)
 ax2.set_title("Positions (arcseconds)")
-
-ax2.scatter(true_positions, final_positions)
-# ax2.errorbar(true_positions, final_positions, yerr=pos_err, fmt='o', capsize=5)
-
-# ax2.scatter(true_positions, initial_positions, c='tab:orange', marker='x', label="intitial")
+ax2.errorbar(true_positions, final_positions, yerr=pos_err, fmt='o', capsize=5)
 
 xlims = ax2.get_xlim()
 ylims = ax2.get_ylim()
@@ -84,8 +72,6 @@ ax2.set_ylim(ylims)
 
 # ax1.scatter(true_positions, true_positions-final_positions)
 ax1.errorbar(true_positions, true_positions-final_positions, yerr=pos_err, fmt='o', capsize=5)
-
[…]

### 2022-12-12 src/scripts/plot_data_resid.py

@@ -9,7 +9,6 @@ plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
-
 # Load model
 tel = p.load(open(paths.data / 'instrument.p', 'rb'))
 model = p.load(open(paths.data / 'model.p', 'rb'))
@@ -23,9 +22,7 @@ plt.suptitle("Data and Residuals", size=15)
 
 ax1 = plt.subplot(1, 3, 1)
 plt.title("Sample Data")
-# plt.imshow(data[0]*1e-3)
 plt.imshow(data*1e-3)
-# plt.imshow(data*1e-3)
 plt.xlabel("Pixels")
 plt.ylabel("Pixels")
 divider = make_axes_locatable(ax1)
@@ -33,15 +30,9 @@ cax = divider.append_axes("right", size="5%", pad=0.1)
 cbar = plt.colorbar(cax=cax)
 cbar.set_label("Counts $x10^3$")
 
-
-# initital_psfs = model.observe()
-# initital_psfs = model.model()
 ax2 = plt.subplot(1, 3, 2)
 plt.title("Initial Residual")
-# plt.imshow((initital_psfs[0] - data[0])*1e-3)
 plt.imshow((initital_psfs - data)*1e-3)
-# plt.imshow((initital_psfs[0] - data)*1e-3)
-# plt.imshow(data[0]/initital_psfs[0])
 plt.xlabel("Pixels")
 plt.ylabel("Pixels")
 divider = make_axes_locatable(ax2)
@@ -49,18 +40,9 @@ cax = divider.append_axes("right", size="5%", pad=0.1)
 cbar = plt.colorbar(cax=cax)
 cbar.set_label("Counts $x10^3$")
 
-
-# final_model = p.load(open(paths.data / 'models_out.p', 'rb'))[-1]
-# final_psfs = final_model.observe()
-# final_psfs = final_model.model()
-
 ax3 = plt.subplot(1, 3, 3)
 plt.title("Final Residual")
-# plt.title("Final Fractional Residual")
-# plt.imshow((final_psfs[0] - data[0])*1e-3)
 plt.imshow((final_psfs - data)*1e-3)
-# plt.imshow((final_psfs[0] - data)*1e-3)
-# plt.imshow(data[0]/final_psfs[0])
 plt.xlabel("Pixels")
 plt.ylabel("Pixels")
[…]

### 2022-12-12 src/scripts/plot_divergence.py

@@ -9,7 +9,6 @@ plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
-
 positions = 'PointSource.position'
 fluxes = 'PointSource.flux'
 zernikes = 'ApplyBasisOPD.coefficients'
@@ -25,23 +24,13 @@ r_res = np.hypot(x_res, y_res)
 
 # Fluxes
 final_fluxes = models_out[-1].get(fluxes)
-res = np.abs(fluxes_in - final_fluxes)
+res = np.abs(fluxes_in - final_fluxes[:, 0])
 l1 = res/fluxes_in
 xs = fluxes_in
 diffs = (models_out[-1].get(fluxes) - models_out[-2].get(fluxes))/fluxes_in
-
 sigma = 1
 
 plt.figure(figsize=(10, 4))
-
-# plt.subplot(1, 2, 1)
-# plt.title("Final Flux Error")
-# plt.scatter(fluxes_in, res)
-# plt.errorbar(fluxes_in, res, yerr=sigma*fluxes_in**0.5, fmt='o')
-# plt.semilogx()
-# plt.axhline(0, c='k')
-# plt.xlabel("True")
-# plt.ylabel("Recovered")
 plt.subplot(1, 2, 1)
 plt.title("Positional Residual")
 plt.scatter(fluxes_in, r2a(r_res))
@@ -49,7 +38,6 @@ plt.semilogx()
 plt.axhline(0, c='k')
 plt.xlabel("Flux")
 plt.ylabel("Positional Residual (arcseconds)")
-
 
 plt.subplot(1, 2, 2)
 plt.title("Fractional Flux Error $|(\hat{f} - f)/\hat{f}|$")
@@ -60,5 +48,4 @@ plt.xlabel("Flux")
 plt.ylabel("Fractional Error")
 
 plt.tight_layout()
-# plt.savefig("divergence")
 plt.savefig(paths.figures / "divergence.pdf", dpi=300)

### 2022-12-12 src/scripts/plot_optics.py

@@ -25,7 +25,6 @@ aberrated_pupil = pupil + opd
 
 # FF
 FF = tel.ApplyPixelResponse.pixel_response
-
 psf_tel = tel.set(['detector'], [None])
 
 # Get psfs
@@ -73,7 +72,6 @@ cbar.set_label("Probability")
 
 plt.subplot(2, 3, 3)
 plt.title("Pixel Response Distribution")
-# plt.hist(FF.flatten(), bins=25)
 plt.hist(FF.flatten(), bins=50)
 plt.ylabel("Counts")
 plt.xlabel("Relative Sensitivity")
@@ -83,7 +81,6 @@ plt.title("Pixel Response")
 plt.imshow(FF)
 plt.xlabel("Pixels")
 plt.ylabel("Pixels")
-# fig.colorbar(im, ax)
 cbar = plt.colorbar()
 cbar.set_label("Sensitivity")
 

### 2022-12-12 src/scripts/plot_progress.py

@@ -24,11 +24,6 @@ flatfield = 'ApplyPixelResponse.pixel_response'
 parameters = [positions, fluxes, zernikes, flatfield]
 
 # Get parameters
-# nepochs = len(models_out)
-# psfs_out = models_out[-1].observe()
-# psfs_out = models_out[-1].model()
-
-
 positions_found  = np.array([model.get(positions) for model in models_out])
 fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
 zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
@@ -43,7 +38,6 @@ scaler = 1e3
 positions_residuals = tel.get(positions) - positions_found
 r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
 r_residuals = r2a(r_residuals_rads)
-
 
 # Positions
 arcsec = r2a(1)
@@ -88,9 +82,7 @@ plt.plot(np.square(flatfield_residuals).sum((-1, -2)))
 plt.axhline(0, c='k', alpha=0.5)
 
 # FF Scatter Plot
-# colours = data.sum(0).flatten()
 colours = data.flatten()
-# colours = data.flatten()
 ind = np.argsort(colours)
 colours = colours[ind]
 

### 2022-12-12 src/scripts/run.py

@@ -11,12 +11,12 @@ print("\nRunning optimise.py")
 os.system('python optimise.py')
 print("Done in {:.2f} seconds".format(t.time() - t0))
 
-# t0 = t.time()
-# print("\nRunning calc_errors.py")
-# os.system('python calc_errors.py')
-# print("Done in {:.2f} seconds".format(t.time() - t0))
+t0 = t.time()
+print("\nRunning calc_errors.py")
+os.system('python calc_errors.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))
 
-# t0 = t.time()
-# print("\nRunning divergence.py")
-# os.system('python divergence.py')
-# print("Done in {:.2f} seconds".format(t.time() - t0))
+t0 = t.time()
+print("\nRunning divergence.py")
+os.system('python divergence.py')
+print("Done in {:.2f} seconds".format(t.time() - t0))

### 2022-12-12 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-In each case, we face serious limitations from an imperfect knowledge of the \ac{prf} -- the map of the intra-pixel and inter-pixel variations in detector sensitivity; and also of the \ac{psf} -- the diffraction-limited pattern by which light from a point source like a starspreads across a detector.
+In each case, we face serious limitations from an imperfect knowledge of the \ac{prf} -- the map of the intra-pixel and inter-pixel variations in detector sensitivity; and also of the \ac{psf} -- the diffraction-limited pattern by which light from a point source like a star spreads across a detector.
 
@@
     \centering
-    % \includegraphics[width=\textwidth]{figures/optics.pdf}
-    \includegraphics[width=\textwidth]{figs/optics.pdf}
+    \includegraphics[width=\textwidth]{figures/optics.pdf}
+    % \includegraphics[width=\textwidth]{figs/optics.pdf}
     \caption{Summary of the optical configuration.
@@
     \label{fig:optics}
-    % \script{gen_optics.py}
+    \script{plot_optics.py}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=\textwidth]{figs/data_resid.pdf}
+    % \includegraphics[width=\textwidth]{figs/data_resid.pdf}
+    \includegraphics[width=\textwidth]{figures/data_resid.pdf}
     \caption{Left: single image of the full four that is used to create the full data set.
@@
     \label{fig:data_residual}
-    % \script{gen_data_resid.py}
+    \script{plot_data_resid.py}
 \end{figure*}
@@
     \centering
-    % \includegraphics[width=0.8\textwidth]{figures/astro_params.pdf}
-    \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
+    \includegraphics[width=0.8\textwidth]{figures/astro_params.pdf}
+    % \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
     \caption{Left: recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds.
@@
     \label{fig:astro_params}
-    % \script{gen_astro_params.py}
+    \script{plot_astro_params.py}
 \end{figure*}
@@
     \label{fig:aberrations}
-    % \script{gen_aberrations.py}
+    \script{plot_aberrations.py}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figs/divergence.pdf}
+    % \includegraphics[width=0.8\textwidth]{figs/divergence.pdf}
+    \includegraphics[width=0.8\textwidth]{figures/divergence.pdf}
     \caption{Parameter recovery of a single star over varying \ac{snr}. Left: Absolute positional residuals vs total flux. Right: Fractional flux error vs total flux.}
     \label{fig:diverge}
-    % \script{gen_FF.py}
+    \script{plot_divergence.py}
 \end{figure*}

### 2022-12-12 src/scripts/make_model_and_data.py
Pre-calcuating some psfs

@@ -109,6 +109,13 @@ model = model.set(flatfield, np.ones((det_npix, det_npix)))
 # Initial PSFs
 initial_psfs = model.model()
 
+# Get plain and aberrated psfs
+psf_tel = tel.set(['detector'], [None])
+psf = psf_tel.set(zernikes, np.zeros(len(tel.get(zernikes)))).model(source=dl.PointSource(wavelengths=wavels))
+aberrated_psf = psf_tel.model(source=dl.PointSource(wavelengths=wavels))
+np.save(paths.data / "plain_psf", psf)
+np.save(paths.data / "aberrated_psf", aberrated_psf)
+
 # Save models and data
 p.dump(tel, open(paths.data / "instrument.p", 'wb'))
 p.dump(source, open(paths.data / "source.p", 'wb'))

### 2022-12-12 src/scripts/plot_optics.py

@@ -25,12 +25,15 @@ aberrated_pupil = pupil + opd
 
 # FF
 FF = tel.ApplyPixelResponse.pixel_response
-psf_tel = tel.set(['detector'], [None])
+# psf_tel = tel.set(['detector'], [None])
 
-# Get psfs
-zernikes = 'ApplyBasisOPD.coefficients'
-psf = psf_tel.set(zernikes, np.zeros(len(tel.get(zernikes)))).model(source=dl.PointSource(wavelengths=wavels))
-aberrated_psf = psf_tel.model(source=dl.PointSource(wavelengths=wavels))
+# # Get psfs
+# zernikes = 'ApplyBasisOPD.coefficients'
+# psf = psf_tel.set(zernikes, np.zeros(len(tel.get(zernikes)))).model(source=dl.PointSource(wavelengths=wavels))
+# aberrated_psf = psf_tel.model(source=dl.PointSource(wavelengths=wavels))
+
+psf = np.load(paths.data / "plain_psf.npy")
+aberrated_psf = np.load(paths.data / "aberrated_psf.npy")
 
 cmap = get_cmap("inferno").copy()
 cmap.set_bad('k',1.)

### 2022-12-12 src/scripts/make_model_and_data.py
Script update

@@ -44,6 +44,9 @@ optics = dl.Optics(optical_layers)
 # Pixel response
 pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(1), [det_npix, det_npix])
 # pix_response = np.ones([det_npix, det_npix])
+counts, bins = np.histogram(pix_response.flatten(), bins=50)
+np.save(paths.data / "pixel_response_counts.npy", counts)
+np.save(paths.data / "pixel_response_bins.npy", bins)
 
 # Detector Noise
 bg_mean = 10

### 2022-12-12 src/scripts/optimise.py

@@ -91,3 +91,60 @@ p.dump(models_out, open(paths.data / "models_out.p", 'wb'))
 # Get final PSFs
 psfs_out = models_out[-1].model()
 np.save(paths.data / 'final_psfs', psfs_out)
+
+# Pre calc FF errors
+thresh = 1000
+fmask = data >= thresh
+out_mask = np.where(data < thresh)
+in_mask = np.where(data >= thresh)
+
+data_tile = np.tile(data, [len(models_out), 1, 1])
+in_mask_tiled = np.where(data_tile >= thresh)
+
+# calculate residuals
+pr_residuals = pix_response[in_mask] - flatfields_found[-1][in_mask]
+
+# for correlation plot
+true_pr_masked = pix_response.at[out_mask].set(1)
+found_pr_masked = flatfields_found[-1].at[out_mask].set(1)
+
+# FF Scatter Plot
+data_sum = data
+colours = data_sum.flatten()
+ind = np.argsort(colours)
+colours = colours[ind]
+
+pr_true_flat = true_pr_masked.flatten()
+pr_found_flat = found_pr_masked.flatten()
+
+pr_true_sort = pr_true_flat[ind]
+pr_found_sort = pr_found_flat[ind]
+
+# Errors
+pfound = flatfields_found[in_mask_tiled].reshape([len(models_out), len(in_mask[0])])
+ptrue = pix_response[in_mask]
+pr_res = ptrue - pfound
+masked_error = np.abs(pr_res).mean(-1)
+
+# FF Scatter Plot
+data_sum = data
+colours = data_sum.flatten()
+ind = np.argsort(colours)
+colours = colours[ind]
+
+pr_true_flat = true_pr_masked.flatten()
+pr_found_flat = found_pr_masked.flatten()
+
+pr_true_sort = pr_true_flat[ind]
+pr_found_sort = pr_found_flat[ind]
+
+np.save(paths.data / 'true_prf_sorted', pr_true_sort)
+np.save(paths.data / 'found_prf_sorted', pr_found_sort)
+np.save(paths.data / 'colours', colours)
+
+# Histogram
+thresh_indx = np.where(fmask)
+res = (pix_response - flatfields_found[-1])[thresh_indx].flatten()
+counts, bins = np.histogram(res.flatten(), bins=51)
+np.save(paths.data / "pixel_response_resid_counts.npy", counts)
+np.save(paths.data / "pixel_response_resid_bins.npy", bins)

### 2022-12-12 src/scripts/plot_FF.py

@@ -9,109 +9,115 @@ plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
-# Load model
-tel = p.load(open(paths.data / 'instrument.p', 'rb'))
-models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
-losses = np.load(paths.data / 'losses.npy')
-data = np.load(paths.data / "data.npy")
-psfs_out = np.load(paths.data / "final_psfs.npy")
+# # Load model
+# tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+# models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
+# losses = np.load(paths.data / 'losses.npy')
+# data = np.load(paths.data / "data.npy")
+# psfs_out = np.load(paths.data / "final_psfs.npy")
 
-positions = 'MultiPointSource.position'
-fluxes = 'MultiPointSource.flux'
-zernikes = 'ApplyBasisOPD.coefficients'
-flatfield = 'ApplyPixelResponse.pixel_response'
-parameters = [positions, fluxes, zernikes, flatfield]
+# positions = 'MultiPointSource.position'
+# fluxes = 'MultiPointSource.flux'
+# zernikes = 'ApplyBasisOPD.coefficients'
+# flatfield = 'ApplyPixelResponse.pixel_response'
+# parameters = [positions, fluxes, zernikes, flatfield]
 
-positions_found  = np.array([model.get(positions) for model in models_out])
-fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
-zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
-flatfields_found = np.array([model.get(flatfield) for model in models_out])
+# positions_found  = np.array([model.get(positions) for model in models_out])
+# fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
+# zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
+# flatfields_found = np.array([model.get(flatfield) for model in models_out])
 
-# Get the residuals
-coeff_residuals = tel.get(zernikes) - zernikes_found
-flux_residuals = tel.get(fluxes) - fluxes_found
[…]

### 2022-12-12 src/scripts/plot_optics.py

@@ -73,9 +73,11 @@ plt.ylabel("Pixels")
 cbar = plt.colorbar()
 cbar.set_label("Probability")
 
+counts = np.load(paths.data / "pixel_response_counts.npy")
+bins = np.load(paths.data / "pixel_response_bins.npy")
 plt.subplot(2, 3, 3)
 plt.title("Pixel Response Distribution")
-plt.hist(FF.flatten(), bins=50)
+plt.hist(bins[:-1], bins, weights=counts)
 plt.ylabel("Counts")
 plt.xlabel("Relative Sensitivity")
 

### 2022-12-12 src/scripts/plot_hist.py
Adding extra histogram plot

@@ -1,0 +1,66 @@
+import jax.numpy as np
+from dLux.utils import radians_to_arcseconds as r2a
+import paths
+import dill as p
+import matplotlib.pyplot as plt
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+# Load model
+tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
+losses = np.load(paths.data / 'losses.npy')
+data = np.load(paths.data / "data.npy")
+psfs_out = np.load(paths.data / "final_psfs.npy")
+
+positions = 'MultiPointSource.position'
+fluxes = 'MultiPointSource.flux'
+zernikes = 'ApplyBasisOPD.coefficients'
+flatfield = 'ApplyPixelResponse.pixel_response'
+parameters = [positions, fluxes, zernikes, flatfield]
+
+positions_found  = np.array([model.get(positions) for model in models_out])
+fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
+zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
+flatfields_found = np.array([model.get(flatfield) for model in models_out])
+
+# Get the residuals
+coeff_residuals = tel.get(zernikes) - zernikes_found
+flux_residuals = tel.get(fluxes) - fluxes_found
+
+scaler = 1e3
+positions_residuals = tel.get(positions) - positions_found
+r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
+r_residuals = r2a(r_residuals_rads)
+pix_response = tel.get(flatfield)
+
+# Plot
+# calculate the mask where there was enough flux to infer the flat field
+thresh = 1000
+data_sort = np.sort(data.flatten())
+nhists = 4
+size = len(data_sort)//nhists
+
+threshes = [0]
+for i in range(nhists):
+    threshes.append(data_sort[size*i])
+threshes.append(data_sort[-1] + 1)
+
+data_flat = data.flatten()
+indexes = []
+for i in range(len(threshes)-1):
+    low = np.where(data_flat >= threshes[i])[0]
+    high = np.where(data_flat < threshes[i+1])[0]
[…]

### 2022-12-12 src/scripts/optimise.py
Updaing FF plot for speed

@@ -102,6 +102,9 @@ data_tile = np.tile(data, [len(models_out), 1, 1])
 in_mask_tiled = np.where(data_tile >= thresh)
 
 # calculate residuals
+tel = p.load(open(paths.data / 'instrument.p', 'rb'))
+pix_response = tel.get(flatfield)
+flatfields_found = np.array([model.get(flatfield) for model in models_out])
 pr_residuals = pix_response[in_mask] - flatfields_found[-1][in_mask]
 
 # for correlation plot

### 2022-12-12 src/scripts/plot_FF.py

@@ -136,8 +136,8 @@ plt.plot(np.linspace(0.7, 1.3), np.linspace(0.7, 1.3), c='k', alpha=0.5)
 # thresh_indx = np.where(fmask)
 # res = (pix_response - flatfields_found[-1])[thresh_indx].flatten()
 
-counts = np.save(paths.data / "pixel_response_resid_counts.npy")
-bins = np.save(paths.data / "pixel_response_resid_bins.npy")
+counts = np.load(paths.data / "pixel_response_resid_counts.npy")
+bins = np.load(paths.data / "pixel_response_resid_bins.npy")
 ax2 = plt.subplot(1, 2, 2)
 # plt.hist(res, bins=51)
 plt.hist(bins[:-1], bins, weights=counts)

### 2022-12-13 src/scripts/plot_astro_params.py
Testing rule script name fix

@@ -16,7 +16,7 @@ models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
 # data = np.load(paths.data / "make_model_and_data/data.npy")
 # psfs_out = np.load(paths.data / "optimise/final_psfs.npy")
 
-# Adding null line to attempt zenodo rerun
+# Adding null line to attempt zenodo re-run
 
 positions = 'MultiPointSource.position'
 fluxes = 'MultiPointSource.flux'

### 2022-12-13 src/scripts/plot_progress.py

@@ -98,6 +98,6 @@ plt.scatter(pr_true_flat, pr_found_flat, c=colours, rasterized=True)
 plt.colorbar()
 
 plt.tight_layout()
-plt.savefig('test')
+# plt.savefig('test')
 
 plt.savefig(paths.figures / "progress.pdf", dpi=300)

### 2022-12-13 src/scripts/plot_FF.py
Updating script dependencies

@@ -9,107 +9,6 @@ plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
-# # Load model
-# tel = p.load(open(paths.data / 'instrument.p', 'rb'))
-# models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
-# losses = np.load(paths.data / 'losses.npy')
-# data = np.load(paths.data / "data.npy")
-# psfs_out = np.load(paths.data / "final_psfs.npy")
-
-# positions = 'MultiPointSource.position'
-# fluxes = 'MultiPointSource.flux'
-# zernikes = 'ApplyBasisOPD.coefficients'
-# flatfield = 'ApplyPixelResponse.pixel_response'
-# parameters = [positions, fluxes, zernikes, flatfield]
-
-# positions_found  = np.array([model.get(positions) for model in models_out])
-# fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
-# zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
-# flatfields_found = np.array([model.get(flatfield) for model in models_out])
-
-# # Get the residuals
-# coeff_residuals = tel.get(zernikes) - zernikes_found
-# flux_residuals = tel.get(fluxes) - fluxes_found
-
-# scaler = 1e3
-# positions_residuals = tel.get(positions) - positions_found
-# r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
-# r_residuals = r2a(r_residuals_rads)
-# pix_response = tel.get(flatfield)
-
-# # Plot
-# # calculate the mask where there was enough flux to infer the flat field
-# thresh = 1000
-# data_sort = np.sort(data.flatten())
-# nhists = 4
-# size = len(data_sort)//nhists
-
-# threshes = [0]
-# for i in range(nhists):
-#     threshes.append(data_sort[size*i])
-# threshes.append(data_sort[-1] + 1)
-
-# data_flat = data.flatten()
-# indexes = []
-# for i in range(len(threshes)-1):
-#     low = np.where(data_flat >= threshes[i])[0]
-#     high = np.where(data_flat < threshes[i+1])[0]
-#     indexes.append(np.intersect1d(low, high))
-
[…]

### 2022-12-13 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We include the first 4 radial terms, ignoring global piston, tip and tilt for a total of 7 terms.
-Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of \variable{output/rms_opd_in.txt}\,nm RMS total \ac{opd}.
+Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of
+\textcolor{red}{showyourwork variable}
+% \variable{output/rms_opd_in.txt}\,nm 
+RMS total \ac{opd}.
 These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}, where we see this induces a large amount of distortion in the \ac{psf}.
@@
 These parameters are recovered to the expected distribution when performed without modelling the \ac{prf}.
-Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of \variable{output/rms_opd_resid.txt} nm RMS.
+Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of
+\textcolor{red}{showyourwork variable}
+% \variable{output/rms_opd_resid.txt} nm RMS. 
 Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram.
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figs/ff.pdf}
+    % \includegraphics[width=0.8\textwidth]{figs/ff.pdf}
+    \includegraphics[width=0.8\textwidth]{figures/ff.pdf}
     \caption{This figure shows the PRF values after optimisation.
@@
     \label{fig:flat_field}
-    % \script{gen_FF.py}
+    \script{plot_FF.py}
 \end{figure*}

### 2022-12-15 src/scripts/calc_errors.py
Things might actually work this time

@@ -14,8 +14,9 @@ except FileExistsError:
 
 # Load model
 tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
-losses = np.load(paths.data / 'optimise/losses.npy')
+# models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
+final_model = p.load(open(paths.data / 'optimise/final_model.p', 'rb'))
+# losses = np.load(paths.data / 'optimise/losses.npy')
 data = np.load(paths.data / "make_model_and_data/data.npy")
 
 positions = 'MultiPointSource.position'
@@ -48,7 +49,8 @@ def hessian(f, x):
     return np.stack([hvp(e) for e in basis]).reshape(x.shape + x.shape)
 
 # model = tel.set(['observation'], [None])
-model = tel
+# model = tel
+model = final_model
 fn = lambda X: dl.utils.poisson_log_likelihood(X, data, model, perturb, 'model')
 
 H = hessian(fn, X)

### 2022-12-15 src/scripts/optimise.py

@@ -93,6 +93,7 @@ with tqdm(range(100), desc='Gradient Descent') as t:
 # Save model and losses
 np.save(paths.data / 'optimise/losses', np.array(losses))
 # p.dump(models_out, open(paths.data / "optimise/models_out.p", 'wb'))
+p.dump(models_out[-1], open(paths.data / "optimise/final_model.p", 'wb'))
 
 # Get final PSFs
 psfs_out = models_out[-1].model()

### 2022-12-15 src/scripts/plot_aberrations.py

@@ -13,7 +13,7 @@ plt.rcParams['figure.dpi'] = 120
 
 # Load model
 tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
+# models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
 # losses = np.load(paths.data / 'optimise/losses.npy')
 # data = np.load(paths.data / "make_model_and_data/data.npy")
 # psfs_out = np.load(paths.data / "optimise/final_psfs.npy")
@@ -47,7 +47,8 @@ r_residuals = r2a(r_residuals_rads)
 # Plot
 # OPDs
 true_opd = tel.ApplyBasisOPD.get_total_opd()
-opds_found = np.array([model.ApplyBasisOPD.get_total_opd() for model in models_out])
+
+opds_found = np.array([tel.set(zernikes, z).ApplyBasisOPD.get_total_opd() for z in zernikes_found])
 found_opd = opds_found[-1]
 opd_residuls = true_opd - opds_found
 opd_rmse_nm = 1e9*np.mean(opd_residuls**2, axis=(-1,-2))**0.5
@@ -57,7 +58,8 @@ vmax = np.max(np.array([true_opd, found_opd]))
 
 # Coefficients
 true_coeff = tel.get(zernikes)
-found_coeff = models_out[-1].get(zernikes)
+# found_coeff = models_out[-1].get(zernikes)
+found_coeff = zernikes_found[-1]
 index = np.arange(len(true_coeff))+4
 
 Nzern = len(true_coeff)

### 2022-12-15 src/scripts/plot_astro_params.py

@@ -11,7 +11,7 @@ plt.rcParams['figure.dpi'] = 120
 
 # Load model
 tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
+# models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
 # losses = np.load(paths.data / 'optimise/losses.npy')
 # data = np.load(paths.data / "make_model_and_data/data.npy")
 # psfs_out = np.load(paths.data / "optimise/final_psfs.npy")

### 2022-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of
-\textcolor{red}{showyourwork variable}
-% \variable{output/rms_opd_in.txt}\,nm 
+% \textcolor{red}{showyourwork variable}
+\variable{output/rms_opd_in.txt}\,nm 
 RMS total \ac{opd}.
@@
 This is necessary to circumvent the high degree of covariance that each individual pixel experiences with the parameters of the \ac{psf}.
-To achieve this the learning rate of the \ac{prf} is set to zero for the first 50 epochs of optimisation.
+To achieve this the learning rate of the \ac{prf} is set to zero for the first 50 epochs of optimisation, after which the \ac{prf} learning rate is initialised while the position, flux and aberrations parameters are frozen for 25 epochs before being allowed to vary again.
 
@@
 Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of
-\textcolor{red}{showyourwork variable}
-% \variable{output/rms_opd_resid.txt} nm RMS. 
+% \textcolor{red}{showyourwork variable}
+\variable{output/rms_opd_resid.txt} nm RMS. 
 Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram.

### 2022-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 This operation is then performed over 15 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm..
-This total model requires the calculation of 30 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 CPU.
+This total model requires the calculation of 30 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of $\sim$ 1.25 seconds to evaluate on an Apple M1 CPU.
 The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise with mean 10 and deviation 2.5 Using this image we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities.
@@
 
-The optimisation is performed for a total of 100 epochs, with each gradient calculation taking \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 Max CPU, leading to a total optimisation time of \textcolor{red}{showyourwork variable} minutes.
+The optimisation is performed for a total of 100 epochs, with each gradient calculation taking $\sim$ 2.75 seconds to evaluate on an Apple M1 Max CPU, leading to a total optimisation time of $\sim$ 4.5 minutes.
 
@@
 Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}.
-The 1-$sigma$ error bars indicate a poor recovery, however these are only marginalised over positions, fluxes and aberration, not the \ac{prf} which is the dominant noise source.
+The 1-$sigma$ error bars
+are generated by harnessing the differentiablility of the model by calculating the covaraince matrix under the Laplace approximation, defined as the negative matrix inverse of the hessian of the likelihood function evaluated at the maximum likelihood estimate, ie the final recovered values.
+These indicate a poor recovery, however they are only marginalised over positions, fluxes and aberration, not the \ac{prf} which is the dominant noise source as its high dimensionality makes marginalisation computationally intractable.
 These parameters are recovered to the expected distribution when performed without modelling the \ac{prf}.

### 2022-12-16 src/scripts/calc_errors.py
Hopefully final build before submission

@@ -1,9 +1,8 @@
 import jax.numpy as np
-import jax.random as jr
 from jax import jvp, grad, jit, linearize
 import dLux as dl
 import paths
-import dill as p
+import pickle as p
 from tqdm import tqdm
 
 import os
@@ -14,16 +13,12 @@ except FileExistsError:
 
 # Load model
 tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-# models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
 final_model = p.load(open(paths.data / 'optimise/final_model.p', 'rb'))
-# losses = np.load(paths.data / 'optimise/losses.npy')
 data = np.load(paths.data / "make_model_and_data/data.npy")
 
 positions = 'MultiPointSource.position'
 fluxes = 'MultiPointSource.flux'
 zernikes = 'ApplyBasisOPD.coefficients'
-flatfield = 'ApplyPixelResponse.pixel_response'
-parameters = [positions, fluxes, zernikes, flatfield]
 
 def perturb(X, model):
     """
@@ -34,11 +29,6 @@ def perturb(X, model):
     model = model.add(zernikes, X[3*Nstars:])
     return model
 
-Nstars = len(tel.get(positions))
-Nzern  = len(tel.get(zernikes))
-X = np.zeros(3*Nstars + Nzern)
-
-
 def hvp(f, primals, tangents):
     return jvp(grad(f), primals, tangents)[1]
 
@@ -48,20 +38,12 @@ def hessian(f, x):
     basis = np.eye(np.prod(np.array(x.shape))).reshape(-1, *x.shape)
     return np.stack([hvp(e) for e in basis]).reshape(x.shape + x.shape)
 
-# model = tel.set(['observation'], [None])
-# model = tel
+Nstars = len(tel.get(positions))
+Nzern  = len(tel.get(zernikes))
+X = np.zeros(3*Nstars + Nzern)
+
 model = final_model
 fn = lambda X: dl.utils.poisson_log_likelihood(X, data, model, perturb, 'model')
-
-H = hessian(fn, X)
-cov_mat = -np.linalg.inv(H)
+cov_mat = -np.linalg.inv(hessian(fn, X))
 
 np.save(paths.data / 'calc_errors/cov_mat', cov_mat)
-
-
-
-# from dLux.utils import bayes
[…]

### 2022-12-16 src/scripts/divergence.py

@@ -5,7 +5,7 @@ import equinox as eqx
 import optax
 import dLux as dl
 import paths
-import dill as p
+import pickle as p
 from tqdm import tqdm
 
 import os
@@ -14,47 +14,57 @@ try:
 except FileExistsError:
     pass
 
+tel = p.load
+
 # Define wavelengths
-wavels = 1e-9 * np.linspace(545, 645, 3)
+# wavels = 1e-9 * np.linspace(545, 645, 3)
+
+# Load parameters
+wavels = np.load(paths.data / 'make_model_and_data/wavelengths.npy')
+tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
 
 # Basic Optical Parameters
-aperture = 0.5
-wf_npix = 512
+aperture = tel.CreateWavefront.diameter
+wf_npix = tel.CreateWavefront.npixels
 
 # Detector Parameters
 det_npix = 100
 sampling_rate = 3
 det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
 
-# Load mask
-# raw_mask = dl.utils.phase_to_opd(np.load("mask.npy"), wavels.mean())
-raw_mask = dl.utils.phase_to_opd(np.load(paths.data / "mask.npy"), wavels.mean())
-mask = dl.utils.scale_array(raw_mask, wf_npix, 0)
+# Update detecotr parameters
+tel = tel.set(["AngularMFT.npixels_out", "AngularMFT.pixel_scale_out"], [det_npix, det_pixsize])
 
-# Zernike Basis
-zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
-coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), [len(zern_basis)])
+# # Load mask
+# raw_mask = dl.utils.phase_to_opd(np.load(paths.data / "mask.npy"), wavels.mean())
+# mask = dl.utils.scale_array(raw_mask, wf_npix, 0)
 
-# Define Optical Configuration
-optical_layers = [
-    dl.CreateWavefront    (wf_npix, aperture),
-    dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
-    dl.ApplyBasisOPD      (zern_basis, coeffs),
-    dl.AddOPD             (mask),
-    dl.NormaliseWavefront (),
[…]

### 2022-12-16 src/scripts/make_model_and_data.py

@@ -4,7 +4,7 @@ import jax.random as jr
 import equinox as eqx
 import dLux as dl
 from dLux.utils import arcseconds_to_radians as a2r
-import dill as p
+import pickle as p
 import paths
 
 import os
@@ -13,7 +13,7 @@ try:
 except FileExistsError:
     pass
 
-'''Create Optics, Detector, Source & Instrument'''
+'''Construct Optics and Detector'''
 # Define wavelengths
 wavels = 1e-9 * np.linspace(545, 645, 3)
 np.save(paths.data / 'make_model_and_data/wavelengths.npy', wavels)
@@ -28,7 +28,6 @@ sampling_rate = 20
 det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
 
 # Load mask
-# raw_mask = dl.utils.phase_to_opd(np.load("mask.npy"), wavels.mean())
 raw_mask = dl.utils.phase_to_opd(np.load(paths.data / "mask.npy"), wavels.mean())
 mask = dl.utils.scale_array(raw_mask, wf_npix, 0)
 
@@ -50,21 +49,20 @@ optics = dl.Optics(optical_layers)
 
 # Pixel response
 pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(1), [det_npix, det_npix])
-# pix_response = np.ones([det_npix, det_npix])
 counts, bins = np.histogram(pix_response.flatten(), bins=50)
 np.save(paths.data / "make_model_and_data/pixel_response_counts.npy", counts)
 np.save(paths.data / "make_model_and_data/pixel_response_bins.npy", bins)
 
-# Detector Noise
-bg_mean = 10
-
 # Create detector layers
+bg_mean = 10
 detector_layers =[dl.ApplyPixelResponse(pix_response),
                   dl.AddConstant(bg_mean)]
 
 # Create Detector object
 detector = dl.Detector(detector_layers)
 
+
+"""Construct sources"""
 # Multiple sources to observe
 Nstars = 20
 flux = 1e8
@@ -74,23 +72,17 @@ true_fluxes = flux + (flux/10)*jr.normal(jr.PRNGKey(2), (Nstars,))
 r_max = 4.5
 true_positions = a2r(jr.uniform(jr.PRNGKey(4), (Nstars, 2), minval=-r_max, maxval=r_max))
 
-# Uniform
-# Nstars = 5**2
-# vals = a2r(np.linspace(-4, 4, int(Nstars**0.5)))
[…]

### 2022-12-16 src/scripts/optimise.py

@@ -1,10 +1,9 @@
 import jax.numpy as np
-import jax.random as jr
 from jax.scipy.stats import poisson
 import equinox as eqx
 import optax
 import paths
-import dill as p
+import pickle as p
 from tqdm import tqdm
 
 import os
@@ -17,12 +16,20 @@ except FileExistsError:
 model = p.load(open(paths.data / 'make_model_and_data/model.p', 'rb'))
 data = np.load(paths.data / "make_model_and_data/data.npy")
 
-positions = 'MultiPointSource.position'
-fluxes = 'MultiPointSource.flux'
-zernikes = 'ApplyBasisOPD.coefficients'
-flatfield = 'ApplyPixelResponse.pixel_response'
-parameters = [positions, fluxes, zernikes, flatfield]
+def time(fn):
+    """Time a function by evaluating it 10 times"""
+    import timeit
+    return timeit.timeit(lambda : fn(), number=10)
 
+# Profile
+@eqx.filter_jit
+def jitted_model(model):
+    return model.model()
+_ = jitted_model(model)
+print(time(lambda: jitted_model(model).block_until_ready())/10)
+
+
+"""Define optimisation schedules and hyper parameters"""
 # Optimisation hyper parameters
 b1 = .75 # Momentum -> Higer = more momentum
 # b2 = 0.5 # Acceleration -> Higer = more momentum
@@ -57,9 +64,16 @@ FF_sched = optax.piecewise_constant_schedule(init_value=ff_lr*1e-8,
                              boundaries_and_scales={ff_start : int(1e8)})
 FF_optimiser = optax.adam(FF_sched, b1=b1)
 
-
 # Combine the optimisers into a list
 optimisers = [pos_optimiser, flux_optimiser, coeff_optimiser, FF_optimiser]
+
+"""Define paths to parameters and define losses"""
+# Define optimisation parameters
+positions = 'MultiPointSource.position'
+fluxes = 'MultiPointSource.flux'
+zernikes = 'ApplyBasisOPD.coefficients'
+flatfield = 'ApplyPixelResponse.pixel_response'
+parameters = [positions, fluxes, zernikes, flatfield]
 
 # Generate out optax optimiser, and also get our args
 optim, opt_state, args = model.get_optimiser(parameters, optimisers, get_args=True)
@@ -76,8 +90,12 @@ def log_prior(model, ff_mean=1., ff_std=0.05):
 def loss_fn(model, data):
     return log_prior(model) + log_like(model, data)
 
[…]

### 2022-12-16 src/scripts/plot_FF.py

@@ -1,7 +1,7 @@
 import jax.numpy as np
 from dLux.utils import radians_to_arcseconds as r2a
 import paths
-import dill as p
+import pickle as p
 import matplotlib.pyplot as plt
 
 plt.rcParams['image.cmap'] = 'inferno'
@@ -16,7 +16,6 @@ colours = np.load(paths.data / 'optimise/colours.npy')
 plt.figure(figsize=(10, 4))
 plt.suptitle("Pixel Response Function Recovery", size=15)
 
-# print("Starting Scatter Plot")
 ax = plt.subplot(1, 2, 1)
 plt.scatter(pr_true_sort, pr_found_sort, c=colours*1e-3, alpha=0.5, rasterized=True)
 cbar = plt.colorbar()
@@ -24,7 +23,6 @@ cbar.set_label("Counts (Photons $x10^3$)")
 plt.title("PRF Correlation")
 plt.ylabel("Recovered")
 plt.xlabel("True")
-# print("Finished Scatter Plot")
 
 xlims = ax.get_xlim()
 ylims = ax.get_ylim()
@@ -32,13 +30,9 @@ ax.set_xlim(xlims)
 ax.set_ylim(ylims)
 plt.plot(np.linspace(0.7, 1.3), np.linspace(0.7, 1.3), c='k', alpha=0.5)
 
-# thresh_indx = np.where(fmask)
-# res = (pix_response - flatfields_found[-1])[thresh_indx].flatten()
-
 counts = np.load(paths.data / "optimise/pixel_response_resid_counts.npy")
 bins = np.load(paths.data / "optimise/pixel_response_resid_bins.npy")
 ax2 = plt.subplot(1, 2, 2)
-# plt.hist(res, bins=51)
 plt.hist(bins[:-1], bins, weights=counts)
 plt.title("PRF Residual Histogram")
 plt.ylabel("Counts")
@@ -46,6 +40,56 @@ plt.xlabel("Residual")
 
 xlim = np.abs(np.array(ax2.get_xlim())).max()
 ax2.set_xlim(-xlim, xlim)
+
+
+# Plot
+# calculate the mask where there was enough flux to infer the flat field
+
+# data = np.load(paths.data / "make_model_and_data/data.npy")
+# flatfields_found = np.load(paths.data / "optimise/flatfields_found.npy")
+# tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
[…]

### 2022-12-16 src/scripts/plot_aberrations.py

@@ -1,7 +1,7 @@
 import jax.numpy as np
 from dLux.utils import radians_to_arcseconds as r2a
 import paths
-import dill as p
+import pickle as p
 import matplotlib.pyplot as plt
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
@@ -13,27 +13,15 @@ plt.rcParams['figure.dpi'] = 120
 
 # Load model
 tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-# models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
-# losses = np.load(paths.data / 'optimise/losses.npy')
-# data = np.load(paths.data / "make_model_and_data/data.npy")
-# psfs_out = np.load(paths.data / "optimise/final_psfs.npy")
+positions_found = np.load(paths.data / "optimise/positions_found.npy")
+fluxes_found = np.load(paths.data / "optimise/fluxes_found.npy")
+zernikes_found = np.load(paths.data / "optimise/zernikes_found.npy")
 
 positions = 'MultiPointSource.position'
 fluxes = 'MultiPointSource.flux'
 zernikes = 'ApplyBasisOPD.coefficients'
 flatfield = 'ApplyPixelResponse.pixel_response'
 parameters = [positions, fluxes, zernikes, flatfield]
-
-# # Get parameters
-# positions_found  = np.array([model.get(positions) for model in models_out])
-# fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
-# zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
-# flatfields_found = np.array([model.get(flatfield) for model in models_out])
-
-positions_found = np.load(paths.data / "optimise/positions_found.npy")
-fluxes_found = np.load(paths.data / "optimise/fluxes_found.npy")
-zernikes_found = np.load(paths.data / "optimise/zernikes_found.npy")
-# flatfields_found = np.load(paths.data / "optimise/flatfields_found.npy")
 
 # Get the residuals
 coeff_residuals = tel.get(zernikes) - zernikes_found
@@ -44,7 +32,6 @@ positions_residuals = tel.get(positions) - positions_found
 r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
 r_residuals = r2a(r_residuals_rads)
 
-# Plot
[…]

### 2022-12-16 src/scripts/plot_astro_params.py

@@ -1,7 +1,7 @@
 import jax.numpy as np
 from dLux.utils import radians_to_arcseconds as r2a
 import paths
-import dill as p
+import pickle as p
 import matplotlib.pyplot as plt
 
 plt.rcParams['image.cmap'] = 'inferno'
@@ -11,29 +11,15 @@ plt.rcParams['figure.dpi'] = 120
 
 # Load model
 tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-# models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
-# losses = np.load(paths.data / 'optimise/losses.npy')
-# data = np.load(paths.data / "make_model_and_data/data.npy")
-# psfs_out = np.load(paths.data / "optimise/final_psfs.npy")
-
-# Adding null line to attempt zenodo re-run
+positions_found = np.load(paths.data / "optimise/positions_found.npy")
+fluxes_found = np.load(paths.data / "optimise/fluxes_found.npy")
+zernikes_found = np.load(paths.data / "optimise/zernikes_found.npy")
 
 positions = 'MultiPointSource.position'
 fluxes = 'MultiPointSource.flux'
 zernikes = 'ApplyBasisOPD.coefficients'
 flatfield = 'ApplyPixelResponse.pixel_response'
 parameters = [positions, fluxes, zernikes, flatfield]
-
-# # Get parameters
-# positions_found  = np.array([model.get(positions) for model in models_out])
-# fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
-# zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
-# flatfields_found = np.array([model.get(flatfield) for model in models_out])
-
-positions_found = np.load(paths.data / "optimise/positions_found.npy")
-fluxes_found = np.load(paths.data / "optimise/fluxes_found.npy")
-zernikes_found = np.load(paths.data / "optimise/zernikes_found.npy")
-# flatfields_found = np.load(paths.data / "optimise/flatfields_found.npy")
 
 # Get the residuals
 coeff_residuals = tel.get(zernikes) - zernikes_found
@@ -77,13 +63,11 @@ ax2.set_xticks([])
 ax2.set_xlim(xlims)
 ax2.set_ylim(ylims)
 
[…]

### 2022-12-16 src/scripts/plot_data_resid.py

@@ -1,6 +1,6 @@
 import jax.numpy as np
 import paths
-import dill as p
+import pickle as p
 import matplotlib.pyplot as plt
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
@@ -10,9 +10,6 @@ plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
 # Load model
-# tel = p.load(open(paths.data / 'instrument.p', 'rb'))
-# model = p.load(open(paths.data / 'model.p', 'rb'))
-# source = p.load(open(paths.data / 'source.p', 'rb'))
 data = np.load(paths.data / "make_model_and_data/data.npy")
 final_psfs = np.load(paths.data / "optimise/final_psfs.npy")
 initital_psfs = np.load(paths.data / "make_model_and_data/initial_psfs.npy")

### 2022-12-16 src/scripts/plot_divergence.py

@@ -1,7 +1,7 @@
 import jax.numpy as np
 from dLux.utils import radians_to_arcseconds as r2a
 import paths
-import dill as p
+import pickle as p
 import matplotlib.pyplot as plt
 
 plt.rcParams['image.cmap'] = 'inferno'
@@ -9,31 +9,25 @@ plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
+# Load data
+fluxes_in = np.load(paths.data / "divergence/divergence_fluxes_in.npy")
+positions_found = np.load(paths.data / "divergence/positions_found.npy")
+fluxes_found = np.load(paths.data / "divergence/fluxes_found.npy")
+
 positions = 'PointSource.position'
 fluxes = 'PointSource.flux'
 zernikes = 'ApplyBasisOPD.coefficients'
 flatfield = 'ApplyPixelResponse.pixel_response'
 
-fluxes_in = np.load(paths.data / "divergence/divergence_fluxes_in.npy")
-# models_out = p.load(open(paths.data / 'divergence/divergence_models_out.p', 'rb'))
-
-positions_found = np.load(paths.data / "divergence/positions_found.npy")
-fluxes_found = np.load(paths.data / "divergence/fluxes_found.npy")
-# zernikes_found = np.load(paths.data / "optimise/zernikes_found.npy")
-# flatfields_found = np.load(paths.data / "optimise/flatfields_found.npy")
-
 # Positions
-# positions_found = np.array([model.get(positions) for model in models_out])
 x_res, y_res = (np.zeros(2)- positions_found[-1]).T
 r_res = np.hypot(x_res, y_res)
 
 # Fluxes
-# final_fluxes = models_out[-1].get(fluxes)
 final_fluxes = fluxes_found[-1]
 res = np.abs(fluxes_in - final_fluxes[:, 0])
 l1 = res/fluxes_in
 xs = fluxes_in
-# diffs = (fluxes_found[-1].get(fluxes) - models_out[-2].get(fluxes))/fluxes_in
 sigma = 1
 
 plt.figure(figsize=(10, 4))
@@ -47,7 +41,7 @@ plt.ylabel("Positional Residual (arcseconds)")
 
 plt.subplot(1, 2, 2)
[…]

### 2022-12-16 src/scripts/plot_hist.py

@@ -10,10 +10,10 @@ plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
 # Load model
-tel = p.load(open(paths.data / 'instrument.p', 'rb'))
-models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
+tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
+# models_out = p.load(open(paths.data / 'models_out.p', 'rb'))
 # losses = np.load(paths.data / 'losses.npy')
-data = np.load(paths.data / "data.npy")
+data = np.load(paths.data / "make_model_and_data/data.npy")
 # psfs_out = np.load(paths.data / "final_psfs.npy")
 
 positions = 'MultiPointSource.position'
@@ -30,6 +30,9 @@ positions_found = np.load(paths.data / "optimise/positions_found.npy")
 fluxes_found = np.load(paths.data / "optimise/fluxes_found.npy")
 zernikes_found = np.load(paths.data / "optimise/zernikes_found.npy")
 flatfields_found = np.load(paths.data / "optimise/flatfields_found.npy")
+
+# colours = np.load(paths.data / 'optimise/colours.npy')
+# print(colours)
 
 # Get the residuals
 coeff_residuals = tel.get(zernikes) - zernikes_found
@@ -60,11 +63,22 @@ for i in range(len(threshes)-1):
     high = np.where(data_flat < threshes[i+1])[0]
     indexes.append(np.intersect1d(low, high))
 
+import matplotlib
+cmap = matplotlib.cm.get_cmap('inferno')
+
+vmin, vmax = data.min(), data.max()
 prf_flat = (pix_response - flatfields_found[-1]).flatten()
 
 for i in range(len(indexes)):
     lab = "{} - {}".format(int(threshes[i]), int(threshes[i+1]))
-    plt.hist(prf_flat[indexes[i]], bins=50, alpha=0.85, label=lab)
+
+    # mean = (threshes[i] + threshes[i+1])/2
+    mean = threshes[i+1]
+    mean_norm = (mean - vmin)/(vmax - vmin)
+    cols = cmap(mean_norm)[:3]
+
+    plt.hist(prf_flat[indexes[i]], bins=50, alpha=0.85, label=lab, color=cols)
+# cbar = plt.colorbar()
 plt.legend()
 plt.xlim(-0.2, 0.2)
 plt.savefig(paths.figures / "hists.pdf", dpi=300)

### 2022-12-16 src/scripts/plot_optics.py

@@ -1,6 +1,6 @@
 import jax.numpy as np
 import paths
-import dill as p
+import pickle as p
 import dLux as dl
 import matplotlib.pyplot as plt
 from matplotlib.cm import get_cmap
@@ -13,6 +13,8 @@ plt.rcParams['figure.dpi'] = 120
 # Load model
 tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
 wavels = np.load(paths.data / 'make_model_and_data/wavelengths.npy')
+psf = np.load(paths.data / "make_model_and_data/plain_psf.npy")
+aberrated_psf = np.load(paths.data / "make_model_and_data/aberrated_psf.npy")
 
 # Pupil
 throughput = tel.CompoundAperture.get_aperture(npixels=tel.get('CreateWavefront.npixels'))
@@ -25,15 +27,6 @@ aberrated_pupil = pupil + opd
 
 # FF
 FF = tel.ApplyPixelResponse.pixel_response
-# psf_tel = tel.set(['detector'], [None])
-
-# # Get psfs
-# zernikes = 'ApplyBasisOPD.coefficients'
-# psf = psf_tel.set(zernikes, np.zeros(len(tel.get(zernikes)))).model(source=dl.PointSource(wavelengths=wavels))
-# aberrated_psf = psf_tel.model(source=dl.PointSource(wavelengths=wavels))
-
-psf = np.load(paths.data / "make_model_and_data/plain_psf.npy")
-aberrated_psf = np.load(paths.data / "make_model_and_data/aberrated_psf.npy")
 
 cmap = get_cmap("inferno").copy()
 cmap.set_bad('k',1.)

### 2022-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \author[a]{Peter G. Tuthill}
+\author[b]{Jordan Dennis}
 
@@
 In this series of papers, we present a new software package, \dlux, designed for fitting high-dimensional parametrized physical optics models to astronomical data.
-By using the Python library \textsc{Jax}\cite{jax}, we obtain \ac{gpu} hardware acceleration, as well as autodiff\cite{Margossian2018} features that enable high-dimensional optimization and inference with gradient descent or \ac{hmc}.
+By using the Python library \textsc{Jax}\cite{jax}, we obtain \ac{gpu} hardware acceleration, as well as automatic differentiation or `autodiff' \cite{Margossian2018} features that enable high-dimensional optimization and inference with gradient descent or \ac{hmc}.
 In the present paper, we focus on the particular problem of estimating the \ac{prf} and \ac{psf} simultaneously - i.e.

### 2022-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{xspace}
+\usepackage{lineno}
+\linenumbers
 %\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
@@
 \author[b,c]{Benjamin J. S. Pope}
+\author[b]{Jordan Dennis}
 \author[a]{Peter G. Tuthill}
-\author[b]{Jordan Dennis}
 
@@
 joint flat field calibration and phase retrieval.
-In Paper II, we will show how autodiff enables improvements to hardware \textit{design}, directly calculating and optimizing the Fisher information and therefore fundamental figures of merit of an optical system end-to-end.
+In Paper~II, we will show how autodiff enables improvements to hardware \textit{design}, directly calculating and optimizing the Fisher information and therefore fundamental figures of merit of an optical system end-to-end.
+And in Paper~III, we will use this for end-to-end deconvolution of high angular resolution imaging with an unknown \ac{psf}.
 We make \dlux available as open-source software on GitHub\footnote{\href{https://github.com/LouisDesdoigts/dLux}{github.com/LouisDesdoigts/dLux}}, and encourage interested readers to use and contribute to this package as a community resource.
@@
 Phase retrieval was memorably performed to infer and correct the serious aberration on the \textit{Hubble Space Telescope} mirror at launch \cite{hubble_phase_ret}.
-Earlier work in the vein of this paper has shown that the phase retrieval problem can be efficiently solved by taking a forwards model of physical optics, obtaining partial derivatives with autodiff, and optimizing its parameters by gradient descent \cite{jurling_fienup,phase_ret_and_design}.
+Earlier work in the vein of this paper has shown that the phase retrieval problem can be efficiently solved even in the case of detector nonlinearity by taking a forwards model of physical optics, obtaining partial derivatives with autodiff, and optimizing its parameters by gradient descent \cite{jurling_fienup,phase_ret_and_design}, including a conference presentation of an early version of the present work \cite{Desdoigts2022}.
 
@@
 In this paper we introduce a new optical simulation package, \dlux, available under an open-source BSD 3 license.
-It significantly extends the capabilities of \texttt{morphine} \cite{pope2021,phase_ret_and_design}, a previous \jax optics package from our team based on \texttt{poppy}, by taking full advantage of more-recent \jax features and libraries.
+It significantly extends the capabilities of \texttt{morphine} \cite{pope2021,phase_ret_and_design}, a previous \jax optics package from our team based on \texttt{poppy}, being rewritten from the ground up taking full advantage of more-recent \jax features and libraries.
 It is built in an object-oriented \jax framework called \equinox \cite{kidger2021equinox}, with an optical system consisting of an \equinox module wrapping a stack of `layers' which are themselves \equinox modules applying phase or amplitude screens, or performing Fourier or Fresnel transforms on a wavefront.
@@
 WaveDiff in TensorFlow has been used for detector calibration \cite{liaudat2021} and \ac{psf} modelling \cite{Liaudat2022}. 
-We believe that \dlux fills a unique role in this ecosystem as the only fully-generalised modelling framework for physical optics.
+We believe that \dlux fills a unique niche in this ecosystem as the most general modelling framework for physical optics.
 This has driven a design focus on flexibility and ease of open-source contribution and development.
-The goal is to provide a general object-oriented framework that enables all those working within optics to harness the advantages of autodiff and hardware acceleration provided by \jax with its user-friendly Numpy-like API.
+The goal is to provide a general object-oriented framework that enables a new user to harness the advantages of autodiff and hardware acceleration provided by \jax with its user-friendly \textsc{NumPy}-like API.
 
@@
 
-The optimisation is performed for a total of 100 epochs, with each gradient calculation taking $\sim$ 2.75 seconds to evaluate on an Apple M1 Max CPU, leading to a total optimisation time of $\sim$ 4.5 minutes.
+The optimisation is performed for a total of 100 epochs, with each value-and-gradient calculation taking $\sim$ 2.75 seconds to evaluate on an Apple~M1~Max CPU, for a total optimisation time of $\sim$ 4.5 minutes.
 
@@
 Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}.
-The 1-$sigma$ error bars
-are generated by harnessing the differentiablility of the model by calculating the covaraince matrix under the Laplace approximation, defined as the negative matrix inverse of the hessian of the likelihood function evaluated at the maximum likelihood estimate, ie the final recovered values.
-These indicate a poor recovery, however they are only marginalised over positions, fluxes and aberration, not the \ac{prf} which is the dominant noise source as its high dimensionality makes marginalisation computationally intractable.
+The 1-$\sigma$ error bars
+are generated by autodiff, calculating the covariance matrix under the Laplace approximation as the negative matrix inverse of the Hessian of the likelihood function evaluated at the maximum likelihood estimate.
+These errorbars are clearly an underestimate: this is understandable as they are marginalised over positions, fluxes and aberration, but not the \ac{prf} residuals which are the dominant noise source.
+The million-dimensional Hessian required to estimate these uncertainties is at present intractable even in \jax.
 These parameters are recovered to the expected distribution when performed without modelling the \ac{prf}.
@@
 \end{figure*}
-
-\section{Discussion}
 

### 2023-04-13 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
-%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
+%\documentclass[nshocompress]{spie}  %>>> to avoid compression of citations
 
@@
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage{wrapfig}
 
@@
 % phase retrieval and detector calibration are important limiting factors in many space missions
-The sensitivity limits of space telescopes are imposed by uncalibrated errors in the point spread function, background light, and detector sensitivity. 
+The sensitivity limits of space telescopes are imposed by uncalibrated errors in the point spread function, photon-noise, background light, and detector sensitivity. 
 %these are hard to calibrate
@@
 
-In this series of papers, we present a new software package, \dlux, designed for fitting high-dimensional parametrized physical optics models to astronomical data.
+In this series of papers, we present a new software package, \dlux, for fitting high-dimensional parametrized physical optics models to astronomical data.
 By using the Python library \textsc{Jax}\cite{jax}, we obtain \ac{gpu} hardware acceleration, as well as automatic differentiation or `autodiff' \cite{Margossian2018} features that enable high-dimensional optimization and inference with gradient descent or \ac{hmc}.
@@
 In Paper~II, we will show how autodiff enables improvements to hardware \textit{design}, directly calculating and optimizing the Fisher information and therefore fundamental figures of merit of an optical system end-to-end.
-And in Paper~III, we will use this for end-to-end deconvolution of high angular resolution imaging with an unknown \ac{psf}.
+And in subsequent papers, we will use this for end-to-end deconvolution of high angular resolution imaging with an unknown \ac{psf}.
 We make \dlux available as open-source software on GitHub\footnote{\href{https://github.com/LouisDesdoigts/dLux}{github.com/LouisDesdoigts/dLux}}, and encourage interested readers to use and contribute to this package as a community resource.
@@
 The \ac{psf} depends on the successive planes through which light passes from the entrance pupil of the telescope through to the detector.
-This can be calculated from physical optics: for a simple camera bringing light from the pupil to focus at a detector plane, the \ac{psf} is the Fourier transform of the input wavefront, the regime of Fraunhofer diffraction.
+This can be calculated from physical optics: for a simple camera bringing light from the pupil to focus at a detector plane (the regime of Fraunhofer diffraction), the \ac{psf} is the Fourier transform of the input wavefront.
 In the more general Fresnel regime where the detector is slightly out of focus, there are additional quadratic phase factors before and after this Fourier transform.
@@
 Changes in telescope pointing or variations in the \ac{psf} with focus couple to these inter- and intra-pixel variations in the flat field to produce changes in overall measured flux.
-This was particularly severe for the K2 mission where periodic thruster firings introduced a saw-tooth systematic on 6 hour timescales, comparable to the timescales of exoplanet transits and asteroseismic signals of interest.
+This was particularly severe for the K2 mission where periodic thruster firings introduced a saw-tooth systematic on 6 hour timescales, comparable to the durations of exoplanet transits and asteroseismic signals of interest.
 
@@
 It is built in an object-oriented \jax framework called \equinox \cite{kidger2021equinox}, with an optical system consisting of an \equinox module wrapping a stack of `layers' which are themselves \equinox modules applying phase or amplitude screens, or performing Fourier or Fresnel transforms on a wavefront.
+This allows us to re-frame the idea of a computational optical model by analogy to a `parametrized neural network', with each operation performed on the wavefront analogous to a neural network layer encoding the physics of the transformation.
+The object-oriented nature of \equinox allows for larger and more complex models to be build and seamlessly integrated than would be possible within the purely functional framework default to \jax.
+\equinox is also built to be directly integrated into the \jax framework, allowing for all \jax function to seamlessly interface with any \equinox models and vice-versa.
+Talk about zodiax too?
 We use the optimization library \optax \cite{optax2020github} for gradient-descent optimization, and the probabilistic programming library \numpyro \cite{Phan2019} for \ac{hmc}.
@@
 Clearly these aberrations have a large effect on the \ac{psf} and would make recovering information very difficult without appropriate calibration.
-Right panel: The top plot shows the histogram of the PRF that is applied in the focal plane.
+Right panel: The top plot shows the histogram of the \ac{prf} that is applied in the focal plane.
 These values have a large spread and would greatly affect any results without calibration.
-The bottom plot shows the full PRF across the whole detector, with a small zoomed region used to show the fine detail that can not be seen when examining the full detector.}
+The bottom plot shows the full \ac{prf} across the whole detector, with a small zoomed region used to show the fine detail that can not be seen when examining the full detector.}
     \label{fig:optics}
@@
 Here we show how to create this optical system within the $\partial$Lux framework:
-
-\newpage
 
@@
 # Zernike Basis (ignore piston, tip, tilt)
-zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
-coeffs = 2e-8 * jr.normal(jr.PRNGKey(1), (len(zern_basis),))
+noll_inds = np.arange(4, 11)
+coeffs = 2e-8 * jr.normal(jr.PRNGKey(1), (len(noll_inds),))
 
 # Define Optical Configuration
-optical_layers = [
+layers = [
     dl.CreateWavefront    (wf_npix, aperture),
-    dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
-    dl.ApplyBasisOPD      (zern_basis, coeffs),
+    dl.ApertureFactory    (wf_npix, secondary_ratio=0.2),
+    dl.AberrationFactory  (wf_npix, zernikes=noll_inds, coefficients=coeffs),
     dl.AddOPD             (np.load("mask.npy")),
@@
 # Create Optics object
-optics = dl.Optics(optical_layers)
+optics = dl.Optics(layers)
 \end{lstlisting}
 
-The first three layers simply modify the amplitude of the wavefront, representing its response to the opaque components of the optics.
-
-Next the wavefront encounters the binary phase plate diffractive pupil from Toliman \cite{tuthill2018}, with a half-wave total \ac{opd} between the binary regions.
-We use this as an information-preserving way to spread the \ac{psf} across a large number of pixels which encodes the PRF information for a large number of pixels simultaneously.
-Alternative methods such as defocusing or using a diffuser \cite{Stefansson2017} come at the cost of eliminating Fourier information which constrain the wavefront and positions.
-An important consideration which is alleviated by the Toliman pupil is that focal-plane wavefront sensing suffers a sign ambiguity for all phase modes that are inversion-symmetric about the origin, and for unambiguous phase retrieval we therefore need an asymmetric or odd-mode symmetric pupil \cite{Martinache2013}.
-This pupil and corresponding aberrated \ac{psf} are shown in the left panel of Figure \ref{fig:optics}.
+The first layer creates our wavefront and the second modifies the amplitude to model its response to the transmission of the aperture.
 
@@
 We include the first 4 radial terms, ignoring global piston, tip and tilt for a total of 7 terms.
-Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of
-% \textcolor{red}{showyourwork variable}
-\variable{output/rms_opd_in.txt}\,nm 
-RMS total \ac{opd}.
+Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm.
+%, for a total of % \textcolor{red}{showyourwork variable} % \variable{output/rms_opd_in.txt}\,nm 55.8\,nm RMS total \ac{opd}.
 These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}, where we see this induces a large amount of distortion in the \ac{psf}.
 
+Next the wavefront encounters the binary phase plate diffractive pupil from Toliman \cite{tuthill2018}, with a half-wave total \ac{opd} between the binary regions.
+We use this as an information-preserving way to spread the \ac{psf} across a large number of pixels which encodes the \ac{prf} information for a large number of pixels simultaneously.
+Alternative methods such as defocusing or using a diffuser \cite{Stefansson2017} come at the cost of eliminating Fourier information which constrain the wavefront and positions.
+An important consideration which is alleviated by the Toliman pupil is that focal-plane wavefront sensing suffers a sign ambiguity for all phase modes that are inversion-symmetric about the origin, and for unambiguous phase retrieval we therefore need an asymmetric or odd-mode symmetric pupil \cite{Martinache2013}.
+This pupil and corresponding aberrated \ac{psf} are shown in the left panel of Figure~\ref{fig:optics}.
+
+\newpage
+
 This is then propagated to the $1024\times1024$ pixel focal plane by a two-sided matrix Fourier transform \cite{Soummer2007,Martinache2020}.
-The detector PRF is modelled by multiplying the resulting \ac{psf} by a pixel sensitivity map which is drawn from a normal distribution with mean 0 and standard deviation 0.05.
+The detector \ac{prf} is modelled by multiplying the resulting \ac{psf} by a pixel sensitivity map which is drawn from a normal distribution with mean 0 and standard deviation 0.1.
 The histogram of pixel sensitivities and total pixel sensitivity map is shown in the right panel of Figure~\ref{fig:optics}.
 Here we show how to create this within the $\partial$Lux framework and create an instrument object to model this:
-\newpage
+
 \begin{lstlisting}[language=Python,frame=single]
 # Pixel response
-pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(2), [det_npix, det_npix])
+pix_response = 1 + 0.1*jr.normal(jr.PRNGKey(2), [det_npix, det_npix])
 
@@
 # Combine into instrument
-tel = dl.Instrument(optics=optics, detector=detector)
+tel = dl.Instrument(optics, detector=detector)
 \end{lstlisting}
 
-This operation is then performed over 15 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm..
-This total model requires the calculation of 30 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of $\sim$ 1.25 seconds to evaluate on an Apple M1 CPU.
-The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise with mean 10 and deviation 2.5 Using this image we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities.
-This gives us 30 stellar positional parameters (15 RA and DEC each), 15 stellar flux parameters, 7 optical aberration parameters, and 1,048,576 detector parameters: the dimensionality of the problem is vastly dominated by the \ac{prf} parameters, which is why this problem is made tractable only by using autodiff.
+This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm.
+This total model requires the calculation of 60 individual \ac{psf}s over a $1024\times1024$ detector from a $512\times512$ pixel wavefront.
+This calculation is vastly dominated by the propagation of the wavefront using a 2-sided MFT, which requires two matrix multiplications for each individual \ac{psf} of sizes $[N_0 \times N_i] \cdot [N_i \times N_i]$ and $[N_0 \times N_i] \cdot [N_i \times N_o]$ where $N_i = 512$ is the input size of the wavefront and $N_o = 1024$ is the output size of the wavefront.
+This takes a total of $\sim$ 1.25 seconds to evaluate on an Apple M1 CPU, with all operations performed using FP32 precision default to \jax.
+FP32 was chosen as this problem does not require precision over a large dynamic range, however FP64 calculations can be performed as required.
+We also apply 4 integer-pixel dithers of +- 20 pixels in the (x,y) directions, equivalent to a 1 $\lambda/D$ dither for a total of 5 images with the total flux spread evenly across each image.
+This helps the over-parameterisation consequential from having a free parameter for each pixel.
+Using pixel-integer dithers similarly allows for a single PSF calculation on a slightly larger detector plane (1064x1064 in this case) and then sampling the smaller 1024x1024 regions to produce the full set of 5 images.
+
+The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s.% and additive Gaussian noise to every pixel as detector read-noise with mean 10 and deviation 2.5
+Using this image we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities.
+This gives us 40 stellar positional parameters (20 RA and DEC each), 20 stellar flux parameters, 7 optical aberration parameters, and 1,048,576 detector parameters: the dimensionality of the problem is vastly dominated by the \ac{prf} parameters, which is why this problem is made tractable only by using autodiff.
+
+% \newpage
 
@@
 \label{sec:phaseretrieval}
+
+To recover these parameters we initialise a naive unaberrated optical model with a uniform \ac{prf}.
+Stellar positions were perturbed by adding a random value drawn from a normal distribution mean 0 deviation 1 pixel and fluxes were perturbed by multiplying their value by a random value value drawn from a normal distribution mean 1 deviation 0.1.
+We will use this model to recover the true values.
+Figure~\ref{fig:data_residual} shows the data and the initial and final residuals.
 
@@
     \includegraphics[width=\textwidth]{figures/data_resid.pdf}
-    \caption{Left: single image of the full four that is used to create the full data set.
-This image has had both photon and detector noise applied.
+    \caption{Left: The sum of the images from which the \ac{prf} is recovered.
 By eye it is clearly difficult to disentangle the astrophysical information.
-The large number of overlapping \ac{psf}s is chosen in order to spread light across the majority of the detector so that we encode the PRF information for as much of the detector as possible.
-Middle: the residual of this single image from the data and the initial uncalibrated model.
+The large number of overlapping \ac{psf}s is chosen in order to spread light across the majority of the detector so that we encode the \ac{prf} information for as much of the detector as possible.
+Middle: The sum of the residual of the data and the initial uncalibrated model.
 Clearly these residuals are large, showing that there is a large amount of calibration required.
-Right: this same residual after the model has been optimised.
-The residual values are much smaller and at the noise floor.
+Right: This same residual after the model has been optimised.
+The residual values are much smaller with no discernable structure remaining.
 A small zoomed region is shown so that the individual pixel-level residuals can be seen.}
@@
 
-To recover these parameters we initialise a naive unaberrated optical model with a uniform \ac{prf}.
-Stellar positions and fluxes are randomly perturbed from their true values.
-We will use this model to recover the true values.
-Figure~\ref{fig:data_residual} shows the data and the initial and final residuals.
-
 % Describe gradient descent implementation & scheduling
 % Takes X time on Y machine with Z precision on params (numbers from ShowYourWork)
-In order to recover these input aberrations, we perform gradient descent using Adam \cite{Kingma2014} as implemented in \optax \cite{optax2020github}, minimizing the posterior of a per-pixel Poisson log-likelihood and a $\chi^2$ log prior on the PRF values, matching the true distribution of mean 1 deviation 0.05.
+In order to recover these input aberrations, we perform gradient descent using Adam \cite{Kingma2014} as implemented in \optax \cite{optax2020github}, minimizing the posterior of a per-pixel Poisson log-likelihood with a $\chi^2$ log prior on the \ac{prf} values, matching the true distribution of mean 1 deviation 0.1.
 We take the gradient of this loss function with respect to the position and flux of all stars, the response in all pixels, and the Zernike coefficients.
@@
 This is necessary to circumvent the high degree of covariance that each individual pixel experiences with the parameters of the \ac{psf}.
-To achieve this the learning rate of the \ac{prf} is set to zero for the first 50 epochs of optimisation, after which the \ac{prf} learning rate is initialised while the position, flux and aberrations parameters are frozen for 25 epochs before being allowed to vary again.
+To achieve this the learning rate of the \ac{prf} is set to zero for the first 100 epochs of optimisation, after which the \ac{prf} learning rate is initialised while the position, flux and aberrations parameters are frozen for the remaining 50 epochs.
+This helps to circumvent the global covariance between the mean \ac{prf} and flux which can result in an overestimation of the flux parameters as the model tries to fit to the noise.
 
-The optimisation is performed for a total of 100 epochs, with each value-and-gradient calculation taking $\sim$ 2.75 seconds to evaluate on an Apple~M1~Max CPU, for a total optimisation time of $\sim$ 4.5 minutes.
+The optimisation is performed for a total of 150 epochs, with each value-and-gradient calculation taking $\sim$ 2.75 seconds to evaluate on an Apple~M1~Max CPU, for a total optimisation time of $\sim$ 6.5 minutes.
+Faster convergence can be achieved by careful tuning of the optimisation hyper-parameters such as learning rate and momentum, however we take a slower approach in this work in order to get good performance across a range of fluxes and PRFs which is analysed in Section~\ref{sec:noise}.
 
-This optimisation strategy performs well, with all of the parameters being recovered down to the noise level.
-This can be seen in Figure~\ref{fig:data_residual} with the right panel showing a sample residual of the data and model after optimisation, which is clearly only noise.
-Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}.
-The 1-$\sigma$ error bars
-are generated by autodiff, calculating the covariance matrix under the Laplace approximation as the negative matrix inverse of the Hessian of the likelihood function evaluated at the maximum likelihood estimate.
-These errorbars are clearly an underestimate: this is understandable as they are marginalised over positions, fluxes and aberration, but not the \ac{prf} residuals which are the dominant noise source.
-The million-dimensional Hessian required to estimate these uncertainties is at present intractable even in \jax.
-These parameters are recovered to the expected distribution when performed without modelling the \ac{prf}.
+This optimisation strategy performs well, with Figure~\ref{fig:data_residual} showing the final residual image with no visually discernable structure.
+A further analysis of parameter recovery over a range of fluxes and PRFs is shown in Section~\ref{sec:noise}.
+Correlations and residual plots of the positions, fluxes and aberrations after the optimisation are shown in Figure~\ref{fig:astro_params} and Figure~\ref{fig:aberrations}.
 Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of
 % \textcolor{red}{showyourwork variable}
-\variable{output/rms_opd_resid.txt} nm RMS. 
+% \variable{output/rms_opd_resid.txt} 
+0.0499\,nm RMS. \textcolor{red}{update with final values}
 Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram.
-The points of the correlation plot are colour-coded with the total counts for each pixel, showing as expected that pixels with higher photon counts having a better recovery of the PRF since these pixels have a greater \ac{snr}.
+The points of the correlation plot are colour-coded with the total counts for each pixel, showing as expected that pixels with higher photon counts having a better recovery of the \ac{prf} since these pixels have a greater \ac{snr}.
 Overall the parameters are very well recovered down to the noise level.
-
-The optical propagation algorithm employs two-sided matrix Fourier transforms \cite{Soummer2007}, which can be greatly improved if deployed onto GPU hardware.
-In our a test case, we initialise from a very naive state far from the truth, requiring a larger number of optimisation epochs than would be required in a real-world example.
-Furthermore as a calibration algorithm across a whole detector the complexity of the model is greatly increased, applying this to real data would likely be focused on small detector regions resulting in less computationally intensive models.
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figures/astro_params.pdf}
+    \includegraphics[width=\textwidth]{figures/astro_params.pdf}
     % \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
@@
     % \includegraphics[width=0.8\textwidth]{figs/ff.pdf}
-    \includegraphics[width=0.8\textwidth]{figures/ff.pdf}
-    \caption{This figure shows the PRF values after optimisation.
+    \includegraphics[width=\textwidth]{figures/ff.pdf}
+    \caption{\ac{prf} values after optimisation.
 Left: Correlation between the pixel-level true and recovered values.
@@
 
+We can also leverage the differentiability of these models in order to estimate the errors in these parameters under the Laplace approximation \cite{bard1974nonlinear}, by taking the inverse Hessian of the Poisson likelihood.
+Using this method marginalised errors can be calculated incorporating parameter covariance.
+There are some limitations using this method as it requires the calculation of the covariance matrix over all parameters - which in this model is over a million, giving a covariance matrix of over a computationally-intractable trillion values.
+Instead we calculate the errors marginalising \textit{only} over stellar positions, fluxes and optical aberrations, with pixel sensitivities held constant.
+This provides us with an accurate calculation of the true errors when photon noise is dominant, but it will remain an underestimate when the \ac{prf} is the dominant noise source.
+We chose to estimate our errors in this manner as opposed to the conventional photon-noise scaling relations in order to better capture covariance between parameters.
+In Section~\ref{sec:noise} we provide a comprehensive analysis of our results with respect to these photon-noise scaling relations.
+
+% The 1-$\sigma$ error bars are generated by autodiff, calculating the covariance matrix under the Laplace approximation as the negative matrix inverse of the Hessian of the likelihood function evaluated at the maximum likelihood estimate.
+These errorbars are clearly an underestimate: this is understandable as they are marginalised over positions, fluxes and aberration, but not the \ac{prf} residuals which are the dominant noise source.
+The million-dimensional Hessian required to estimate these uncertainties is at present intractable even in \jax.
+These parameters are recovered to the expected distribution when performed without modelling the \ac{prf}.
+
+% The optical propagation algorithm employs two-sided matrix Fourier transforms \cite{Soummer2007}, which can be greatly improved if deployed onto GPU hardware.
+In our a test case, we initialise from a very naive state far from the truth, requiring a larger number of optimisation epochs than would be required in a real-world example.
+Furthermore as a calibration algorithm across a whole detector the complexity of the model is greatly increased, applying this to real data would likely be focused on small detector regions resulting in less computationally intensive models.
+
 % poor model
@@
 Similarly since we are working with direct image data as opposed to reduced data-products, miss-calibrations of the system are simple to diagnose through direct examination of the residuals of the data and model.
-Figure~\ref{fig:data_residual} demonstrates that in this example we can be sure of a good fit because the final residual is pure noise.
+Figure~\ref{fig:data_residual} demonstrates that in this example we can be sure of a good fit because the final residual contains no visually identifiable structure.
 
-% \subsection{Limitations}
-This algorithm demonstrates promising performance applied to complex high-dimensional optimisation problems as shown in Section~\ref{sec:phaseretrieval}.
+% % \subsection{Limitations}
+% This algorithm demonstrates promising performance applied to complex high-dimensional optimisation problems as shown in Section~\ref{sec:phaseretrieval}.
 However, this method does have its limitations.
@@
 
-\begin{figure*}
-    \centering
-    % \includegraphics[width=0.8\textwidth]{figs/divergence.pdf}
-    \includegraphics[width=0.8\textwidth]{figures/divergence.pdf}
-    \caption{Parameter recovery of a single star over varying \ac{snr}. Left: Absolute positional residuals vs total flux. Right: Fractional flux error vs total flux.}
-    \label{fig:diverge}
-    \script{plot_divergence.py}
-\end{figure*}
+% \begin{figure*}
+%     \centering
+%     % \includegraphics[width=0.8\textwidth]{figs/divergence.pdf}
+%     \includegraphics[width=0.8\textwidth]{figures/divergence.pdf}
+%     \caption{Parameter recovery of a single star over varying \ac{snr}. Left: Absolute positional residuals vs total flux. Right: Fractional flux error vs total flux.}
+%     \label{fig:diverge}
+%     \script{plot_divergence.py}
+% \end{figure*}
 
-% Without the ability to recover the \ac{psf} model accurately, pixel level information will be dominated by the noise processes and the PRF will not be sufficiently encoded.
-This can be directly seen in Figure~\ref{fig:flat_field}, with the PRF residual having greater variance in for low-flux pixels.
+% Without the ability to recover the \ac{psf} model accurately, pixel level information will be dominated by the noise processes and the \ac{prf} will not be sufficiently encoded.
+This can be directly seen in Figure~\ref{fig:flat_field}, with the \ac{prf} residual having greater variance in for low-flux pixels.
 This however has a simple solution of observing brighter targets for calibration purposes or observing chosen targets for longer, until a sufficient amount of information is encoded into the data.
@@
 % Dithering or N stars requirement
+
+\section{Noise \& Performance Analysis}
+\label{sec:noise}
+
+This section examines how well this method performs across a range of both \ac{prf} and flux values, which are our two dominant noise sources.
+Recovering parameters in a Poisson-noise-dominated regime for simple optical systems should have the following relationships:
+
+$$\sigma_{\text{flux}} = 1/\sqrt{N_\text{phot}} \ \ \text{(photons)}$$
+
+$$\sigma_{\text{position}} = \frac{1}{\pi} \sqrt{2 / N_{\text{phot}}} \frac{\lambda}{D} \ \ \text{(radians)}$$
+
+$$\sigma_{\text{zernike}} = 1 / \sqrt N_{\text{phot}} \ \ \text{(RMS radians)}$$
+
+Comparing the errors in the recovered parameters across a range of fluxes provides a good way to examine when we are in the photon noise dominated regime.
+After optimisation the errors are calculated for each parameter and compared to these expected values.
+ie:
+
+$$\sigma_{\text{relative}} = \sum_{i=1}^{N} \frac{1}{N} \frac{\sigma_{\text{i recovered}}}{\sigma_{\text{i expected}}}$$
+
+The top-left panel of Figure~\ref{fig:noise} shows the weighted Pearson correlation coefficient~\cite{Costa2011} as a function of both flux and \ac{prf}.
+The transition where the \ac{prf} is recoverable denotes two regimes: one where photon noise dominates and another where \ac{prf} noise dominates.
+We have distinct results in these different regimes.
+Astrophysical and optical parameters are recovered to the photon noise expectation, while there is poor recovery of the \ac{prf} when photon noise is the dominant source and vice versa when the \ac{prf} is the dominant noise source.
+The top-right and bottom panels of Figure~\ref{fig:noise} show how $\sigma_{\text{relative}}$ changes as a function of both flux and \ac{prf}.
+These $\sigma_{\text{expected}}$ values are expected to be underestimates of the true values because they are not marginalised over all of the fitted parameters in the forwards model.
+By examining Figure~\ref{fig:noise}, we can see that there is a transition between two regimes: photon noise dominated and \ac{prf} noise dominated.
+In the photon noise dominated regime, we can see that the parameters are recovered to the level expected by these relationships and diverge from these once the \ac{prf} becomes the primary noise source.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=1.0\textwidth]{figs/noise_performance.pdf}
+    % \includegraphics[width=0.8\textwidth]{figures/noise_performance.pdf}
+    \caption{Recovered parameter error relative to photon noise expectation. Left: Mean relative per-mode Zernike error. Middle: Mean relative flux error. Right: Mean relative positional error.}
+    \label{fig:noise}
+\end{figure*}
+
+% \textcolor{red}{The paragraph below has been moved into optimisation in order to describe the error bar calcualtion process}
+% We can leverage the differentiability of these models in order to get a better estimate of these errors under the Laplace approximation (\textcolor{red}{citiation needed}) using a poissonian likelihood.
+Using this method marginalised errors can be calculated incorporating parameter covariance.
+There are some limitations using this method as it requires the calculation of the covariance matrix over all parameters, which in this model is over a million parameters, giving a covariance matrix of over a trillion values which is computationally intractable.
+Instead we calculate the errors using this method, marginalising over stellar positions, fluxes and optical aberrations.
+This provides us with an accurate calculation of the true errors in the photon-noise dominated regime, but it will remain an underestimate when the \ac{prf} is the dominant noise source.
+
+% Figure~\ref{fig:noise_laplace} again shows the $\sigma_{\text{relative}}$, but taking the $\sigma_{\text{expected}}$ from the calculated covariance matrix.
+This broadly shows the same result but postulates that the parameters are recovered overall closer to the expected noise value.
+
+% \begin{figure*}
+%     \centering
+%     \includegraphics[width=1.0\textwidth]{figs/noise_performance_laplace.pdf}
+%     % \includegraphics[width=0.8\textwidth]{figures/noise_performance.pdf}
+%     \caption{Recovered parameter error relative to covariance matrix estimation under the Laplace approximation.
+Left: Mean relative per-mode Zernike error.
+Middle: Mean relative flux error.
+Right: Mean relative positional error.}
+%     \label{fig:noise_laplace}
+% \end{figure*}
+
+It is important to note in this section that there is a degree of scatter in the recovered results.
+Due to the gradient descent optimisation strategy there is a degree of sensitivity to the optimisation hyper-parameters, such as learning rates which must be tuned based on the flux, momentum, and distance to the noise floor.
+These parameters are simple enough to tune for individual data sets with similar scales, but it is non-trivial to automate this process for data with a vast range of scales like those being analysed here.
+The hyper-parameters chosen were tuned on the 1e8 flux data set and roughly scaled to the work with the other data sets, which is why we observe the best recovery for that set of parameters.
+These figures are designed to demonstrate the overall trends in the different noise regimes, rather than to be a precise analysis of the fundamental limits.
 
@@
 Prior knowledge of a structured \ac{psf}, such as with the diffractive pupil architecture of Toliman \cite{Guyon2012,Guyon2013}, may further improve this deblending and retrieval.
+
+% Adding for reviewer #1 comment on intra-pixel sensitivity
+High-precision photometric measurements often need to also consider variations in the intra-pixel sensitivity.
+This work only examined the inter-pixel sensitivity for simplicity, however the \ac{psf}s examined here were sampled at 10x the Nyquist limit, well beyond the typical 1-1.5x Nyquist used in practice.
+The ability to recover sensitivity variations over these sub-fringe scales show that this method can be directly applied to recovery of both inter and intra-pixel sensitivity fluctuations.
 

### 2023-04-13 src/scripts/analyse.py
General restructure for revisions

@@ -1,37 +1,0 @@
-import os
-import time as t
-
-# t0 = t.time()
-# print("\nRunning plot_progress.py")
-# os.system('python plot_progress.py')
-# print("Done in {:.2f} seconds".format(t.time() - t0))
-
-t0 = t.time()
-print("\nRunning plot_astro_params.py")
-os.system('python plot_astro_params.py')
-print("Done in {:.2f} seconds".format(t.time() - t0))
-
-t0 = t.time()
-print("\nRunning plot_aberrations.py")
-os.system('python plot_aberrations.py')
-print("Done in {:.2f} seconds".format(t.time() - t0))
-
-t0 = t.time()
-print("\nRunning plot_data_resid.py")
-os.system('python plot_data_resid.py')
-print("Done in {:.2f} seconds".format(t.time() - t0))
-
-t0 = t.time()
-print("\nRunning plot_FF.py")
-os.system('python plot_FF.py')
-print("Done in {:.2f} seconds".format(t.time() - t0))
-
-t0 = t.time()
-print("\nRunning plot_optics.py")
-os.system('python plot_optics.py')
-print("Done in {:.2f} seconds".format(t.time() - t0))
-
-t0 = t.time()
-print("\nRunning plot_divergence.py")
-os.system('python plot_divergence.py')
-print("Done in {:.2f} seconds".format(t.time() - t0))

### 2023-04-13 src/scripts/calc_errors.py

@@ -1,49 +1,0 @@
-import jax.numpy as np
-from jax import jvp, grad, jit, linearize
-import dLux as dl
-import paths
-import pickle as p
-from tqdm import tqdm
-
-import os
-try:
-    os.mkdir(paths.data / "calc_errors")
-except FileExistsError:
-    pass
-
-# Load model
-tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-final_model = p.load(open(paths.data / 'optimise/final_model.p', 'rb'))
-data = np.load(paths.data / "make_model_and_data/data.npy")
-
-positions = 'MultiPointSource.position'
-fluxes = 'MultiPointSource.flux'
-zernikes = 'ApplyBasisOPD.coefficients'
-
-def perturb(X, model):
-    """
-    Perturbs the values of the model
-    """
-    model = model.add(positions, X[:2*Nstars].reshape((Nstars, 2)))
-    model = model.add(fluxes, X[2*Nstars:3*Nstars])
-    model = model.add(zernikes, X[3*Nstars:])
-    return model
-
-def hvp(f, primals, tangents):
-    return jvp(grad(f), primals, tangents)[1]
-
-def hessian(f, x):
-    _, hvp = linearize(grad(f), x)
-    hvp = jit(hvp)  # seems like a substantial speedup to do this
-    basis = np.eye(np.prod(np.array(x.shape))).reshape(-1, *x.shape)
-    return np.stack([hvp(e) for e in basis]).reshape(x.shape + x.shape)
-
-Nstars = len(tel.get(positions))
-Nzern  = len(tel.get(zernikes))
-X = np.zeros(3*Nstars + Nzern)
-
-model = final_model
-fn = lambda X: dl.utils.poisson_log_likelihood(X, data, model, perturb, 'model')
-cov_mat = -np.linalg.inv(hessian(fn, X))
-
-np.save(paths.data / 'calc_errors/cov_mat', cov_mat)

### 2023-04-13 src/scripts/divergence.py

@@ -1,202 +1,0 @@
-import jax.numpy as np
-import jax.random as jr
-from jax.scipy.stats import poisson
-import equinox as eqx
-import optax
-import dLux as dl
-import paths
-import pickle as p
-from tqdm import tqdm
-
-import os
-try:
-    os.mkdir(paths.data / "divergence")
-except FileExistsError:
-    pass
-
-tel = p.load
-
-# Define wavelengths
-# wavels = 1e-9 * np.linspace(545, 645, 3)
-
-# Load parameters
-wavels = np.load(paths.data / 'make_model_and_data/wavelengths.npy')
-tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-
-# Basic Optical Parameters
-aperture = tel.CreateWavefront.diameter
-wf_npix = tel.CreateWavefront.npixels
-
-# Detector Parameters
-det_npix = 100
-sampling_rate = 3
-det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
-
-# Update detecotr parameters
-tel = tel.set(["AngularMFT.npixels_out", "AngularMFT.pixel_scale_out"], [det_npix, det_pixsize])
-
-# # Load mask
-# raw_mask = dl.utils.phase_to_opd(np.load(paths.data / "mask.npy"), wavels.mean())
-# mask = dl.utils.scale_array(raw_mask, wf_npix, 0)
-
-# # Zernike Basis
-# zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
-# coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), [len(zern_basis)])
-
-# # Define Optical Configuration
-# optical_layers = [
-#     dl.CreateWavefront    (wf_npix, aperture),
-#     dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
-#     dl.ApplyBasisOPD      (zern_basis, coeffs),
-#     dl.AddOPD             (mask),
-#     dl.NormaliseWavefront (),
-#     dl.AngularMFT         (det_npix, det_pixsize)]
-
-# # Create Optics object
-# optics = dl.Optics(optical_layers)
-
-# Pixel response
-pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(0), [det_npix, det_npix])
-
-# Create Detector object
[…]

### 2023-04-13 src/scripts/lib.py

@@ -1,0 +1,444 @@
+import os
+import time as t
+import paths
+import jax.numpy as np
+import jax.random as jr
+from jax.scipy.stats import poisson
+import equinox as eqx
+import optax
+import paths
+from tqdm import tqdm
+import zodiax as zdx
+from zodiax.experimental.serialisation import serialise, deserialise
+from observation import MultiImage, IntegerDither
+import dLux.utils as dlu
+import dLux as dl
+# from lib import mkdir
+import sys
+
+# Core jax
+import jax.numpy as np
+import jax.random as jr
+import dLux as dl
+import dLux.utils as dlu
+import paths
+import json
+import time
+import os
+import sys
+from zodiax.experimental.serialisation import serialise
+
+from jax import jvp, grad, jit, linearize
+
+
+def mkdir(path):
+    try:
+        os.mkdir(path)
+    except FileExistsError:
+        pass
+
+def format_time(seconds):
+    """Format time in seconds to hours, minutes and seconds"""
+    hours = int(seconds // 3600)
+    seconds -= hours * 3600
+    minutes = int(seconds // 60)
+    seconds -= minutes * 60
+    if minutes == 0:
+        return f"{int(seconds):02d} sec"
+    return f"{minutes:02d} min {int(seconds):02d} sec"
+
+def run_script(script_name, sub_dir, make_dirs=False):
+    t0 = t.time()
+    print(f"Running {script_name}.py ")
+    if make_dirs:
+        mkdir(paths.data / script_name)
+        mkdir(paths.data / script_name / sub_dir)
+    os.system(f'python {script_name}.py {sub_dir}')
+    print(f"Executed in {format_time(t.time() - t0)}")
+
+def weighted_pearson_corr_coeff(x, y, weights):
+    """
+    Calculates the weighted Pearson correlation coefficient between two arrays of data.
+    
+    Parameters:
+    x (numpy array): Array of data for the first variable
+    y (numpy array): Array of data for the second variable
+    weights (numpy array): Array of weights, one for each observation
+    
+    Returns:
+    w (float): Weighted Pearson correlation coefficient
+    """
[…]

### 2023-04-13 src/scripts/make_model_and_data.py

@@ -1,128 +1,0 @@
-# Core jax
-import jax.numpy as np
-import jax.random as jr
-import equinox as eqx
-import dLux as dl
-from dLux.utils import arcseconds_to_radians as a2r
-import pickle as p
-import paths
-
-import os
-try:
-    os.mkdir(paths.data / "make_model_and_data")
-except FileExistsError:
-    pass
-
-'''Construct Optics and Detector'''
-# Define wavelengths
-wavels = 1e-9 * np.linspace(545, 645, 3)
-np.save(paths.data / 'make_model_and_data/wavelengths.npy', wavels)
-
-# Basic Optical Parameters
-aperture = 0.5
-wf_npix = 512
-
-# Detector Parameters
-det_npix = 1024
-sampling_rate = 20
-det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)
-
-# Load mask
-raw_mask = dl.utils.phase_to_opd(np.load(paths.data / "mask.npy"), wavels.mean())
-mask = dl.utils.scale_array(raw_mask, wf_npix, 0)
-
-# Zernike Basis
-zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
-coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), [len(zern_basis)])
-
-# Define Optical Configuration
-optical_layers = [
-    dl.CreateWavefront    (wf_npix, aperture),
-    dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
-    dl.ApplyBasisOPD      (zern_basis, coeffs),
-    dl.AddOPD             (mask),
-    dl.NormaliseWavefront (),
-    dl.AngularMFT         (det_npix, det_pixsize)]
-
-# Create Optics object
-optics = dl.Optics(optical_layers)
-
-# Pixel response
-pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(1), [det_npix, det_npix])
-counts, bins = np.histogram(pix_response.flatten(), bins=50)
-np.save(paths.data / "make_model_and_data/pixel_response_counts.npy", counts)
-np.save(paths.data / "make_model_and_data/pixel_response_bins.npy", bins)
-
-# Create detector layers
-bg_mean = 10
-detector_layers =[dl.ApplyPixelResponse(pix_response),
-                  dl.AddConstant(bg_mean)]
-
-# Create Detector object
[…]

### 2023-04-13 src/scripts/observation.py

@@ -1,0 +1,72 @@
+import jax.numpy as np
+from jax import tree_map
+from dLux.observations import AbstractObservation
+from dLux import model
+
+class MultiImage(AbstractObservation):
+    N : int
+
+    def __init__(self, N):
+        super().__init__(name='observation')
+        self.N = int(N)
+    
+    def observe(self, instrument, **kwargs):
+        """
+        Models the instrument sources through the instrument optics and
+        then returns self.N images of the scene with the source flux 
+        distributed across the images.
+        """
+        # if 'individual_psfs' in kwargs and kwargs['individual_psfs']:
+        #     source = instrument.scene.sources[0]
+        #     wavelengths = source.get_wavelengths()
+        #     weights = source.get_weights()
+        #     fluxes = source.get_flux()
+        #     positions = source.get_position()
+
+        #     psfs = np.array([instrument.optics.propagate_multi(wavelengths, 
+        #         positions[i], weights) for i in range(len(fluxes))])
+        #     psfs *= fluxes
+
+        #     psfs /= self.N
+
+        psf = model(instrument.optics, sources=instrument.sources)
+        psf /= self.N
+        if instrument.detector is not None:
+            image = instrument.detector.apply_detector(psf)
+        else:
+            image = psf
+        return np.tile(image, (self.N, 1, 1))
+
+
+class IntegerDither(AbstractObservation):
+    dithers : list[int]
+    padding : int
+    N : int
+
+    def __init__(self, dithers):
+        super().__init__(name='observation')
+        self.dithers = dithers
+        self.N = len(dithers)
+        self.padding = int(np.array(dithers).max())
+    
+    def observe(self, instrument, **kwargs):
+        """
+        Models the instrument sources through the instrument optics and
+        then returns self.N images of the scene with the source flux 
+        distributed across the images.
+        """
+        padded_model = instrument.add('AngularMFT.npixels_out', 
+            2*self.padding)
+        c = padded_model.AngularMFT.npixels_out//2
+        s = instrument.AngularMFT.npixels_out//2
+
+        source = instrument.sources
+        psf = padded_model.optics.model(source)
+        psfs_out = []
+        for i, dither in enumerate(self.dithers):
+            x, y = dither
+            psfs_out.append(psf[c+y-s:c+y+s, c+x-s:c+x+s]/self.N)
+
+        apply_detector = lambda psf: instrument.detector.apply_detector(psf)
+        psfs = np.array(tree_map(apply_detector, psfs_out))
+        return psfs

### 2023-04-13 src/scripts/optimise.py

@@ -1,189 +1,0 @@
-import jax.numpy as np
-from jax.scipy.stats import poisson
-import equinox as eqx
-import optax
-import paths
-import pickle as p
-from tqdm import tqdm
-
-import os
-try:
-    os.mkdir(paths.data / "optimise")
-except FileExistsError:
-    pass
-
-# Load model
-model = p.load(open(paths.data / 'make_model_and_data/model.p', 'rb'))
-data = np.load(paths.data / "make_model_and_data/data.npy")
-
-def time(fn):
-    """Time a function by evaluating it 10 times"""
-    import timeit
-    return timeit.timeit(lambda : fn(), number=10)
-
-# Profile
-@eqx.filter_jit
-def jitted_model(model):
-    return model.model()
-_ = jitted_model(model)
-print(time(lambda: jitted_model(model).block_until_ready())/10)
-
-
-"""Define optimisation schedules and hyper parameters"""
-# Optimisation hyper parameters
-b1 = .75 # Momentum -> Higer = more momentum
-# b2 = 0.5 # Acceleration -> Higer = more momentum
-
-# Position
-pos_lr, pos_start, pos_stop, pos_restart = 1e-8, 0, 50, 75
-pos_sched = optax.piecewise_constant_schedule(init_value=pos_lr*1e-8,
-                             boundaries_and_scales={pos_start   : int(1e8),
-                                                    pos_stop    : int(1e-8),
-                                                    pos_restart : int(1e8)})
-pos_optimiser   = optax.adam(pos_sched, b1=b1)
-
-# Flux
-flux_lr, flux_start, flux_stop, flux_restart = 2.5e6, 0, 50, 75
-flux_sched = optax.piecewise_constant_schedule(init_value=flux_lr*1e-8,
-                             boundaries_and_scales={flux_start   : int(1e8),
-                                                    flux_stop    : int(1e-8),
-                                                    flux_restart : int(1e8)})
-flux_optimiser = optax.adam(flux_sched, b1=b1)
-
-# Zernikes
-coeff_lr, coeff_start, coeff_stop, coeff_restart = 5e-9, 0, 50, 75
-coeff_sched = optax.piecewise_constant_schedule(init_value=coeff_lr*1e-8,
-                             boundaries_and_scales={coeff_start   : int(1e8),
[…]

### 2023-04-13 src/scripts/plot_FF.py

@@ -1,26 +1,54 @@
 import jax.numpy as np
 from dLux.utils import radians_to_arcseconds as r2a
 import paths
-import pickle as p
 import matplotlib.pyplot as plt
+from zodiax.experimental.serialisation import serialise, deserialise
+from observation import MultiImage
+import jax.scipy as jsp
+from lib import weighted_pearson_corr_coeff
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
-pr_true_sort = np.load(paths.data / 'optimise/true_prf_sorted.npy')
-pr_found_sort = np.load(paths.data / 'optimise/found_prf_sorted.npy')
-colours = np.load(paths.data / 'optimise/colours.npy')
 
-plt.figure(figsize=(10, 4))
+flux = 1e8
+PRFdev = 1e-1
+sub_dir = f"flux_{flux:.0e}_PRFdev_{PRFdev:.0e}"
+
+# Load data
+tel = deserialise(paths.data / f'make_model_and_data/{sub_dir}/instrument.zdx')
+model = deserialise(paths.data / f'optimise/{sub_dir}/final_model.zdx')
+data = np.load(paths.data / f'make_model_and_data/{sub_dir}/data.npy')
+params = deserialise(paths.data / f"make_model_and_data/{sub_dir}/params.zdx")
+
+# PRFs
+pr_true_sort = np.load(paths.data / f'process/{sub_dir}/true_prf_sorted.npy')
+pr_found_sort = np.load(paths.data / f'process/{sub_dir}/found_prf_sorted.npy')
+colours = np.load(paths.data / f'process/{sub_dir}/colours.npy')
+
+# Histograms
+counts = np.load(paths.data / f"process/{sub_dir}/PRF_found_counts.npy")
+bins = np.load(paths.data / f"process/{sub_dir}/PRF_found_bins.npy")
+resid_counts = np.load(paths.data / f"process/{sub_dir}/PRF_resid_counts.npy")
+resid_bins = np.load(paths.data / f"process/{sub_dir}/PRF_resid_bins.npy")
+
+# Correlation coefficient
+true_prf = model.PRF.pixel_response
[…]

### 2023-04-13 src/scripts/plot_aberrations.py

@@ -4,52 +4,40 @@ import paths
 import pickle as p
 import matplotlib.pyplot as plt
 from mpl_toolkits.axes_grid1 import make_axes_locatable
+from zodiax.experimental.serialisation import serialise, deserialise
+from observation import MultiImage
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
+flux = 1e8
+PRFdev = 1e-1
+sub_dir = f"flux_{flux:.0e}_PRFdev_{PRFdev:.0e}"
 
 # Load model
-tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-positions_found = np.load(paths.data / "optimise/positions_found.npy")
-fluxes_found = np.load(paths.data / "optimise/fluxes_found.npy")
-zernikes_found = np.load(paths.data / "optimise/zernikes_found.npy")
-
-positions = 'MultiPointSource.position'
-fluxes = 'MultiPointSource.flux'
-zernikes = 'ApplyBasisOPD.coefficients'
-flatfield = 'ApplyPixelResponse.pixel_response'
-parameters = [positions, fluxes, zernikes, flatfield]
-
-# Get the residuals
-coeff_residuals = tel.get(zernikes) - zernikes_found
-flux_residuals = tel.get(fluxes) - fluxes_found
-
-scaler = 1e3
-positions_residuals = tel.get(positions) - positions_found
-r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
-r_residuals = r2a(r_residuals_rads)
+tel = deserialise(paths.data / f'make_model_and_data/{sub_dir}/instrument.zdx')
+model = deserialise(paths.data / f'optimise/{sub_dir}/final_model.zdx')
 
 # OPDs
-true_opd = tel.ApplyBasisOPD.get_total_opd()
-
-opds_found = np.array([tel.set(zernikes, z).ApplyBasisOPD.get_total_opd() for z in zernikes_found])
-found_opd = opds_found[-1]
-opd_residuls = true_opd - opds_found
+true_opd = tel.Aberrations.get_opd()
+found_opd = model.Aberrations.get_opd()
[…]

### 2023-04-13 src/scripts/plot_astro_params.py

@@ -1,49 +1,43 @@
 import jax.numpy as np
 from dLux.utils import radians_to_arcseconds as r2a
 import paths
-import pickle as p
 import matplotlib.pyplot as plt
+from zodiax.experimental.serialisation import serialise, deserialise
+from observation import MultiImage
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
+flux = 1e8
+PRFdev = 1e-1
+sub_dir = f"flux_{flux:.0e}_PRFdev_{PRFdev:.0e}"
+
 # Load model
-tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-positions_found = np.load(paths.data / "optimise/positions_found.npy")
-fluxes_found = np.load(paths.data / "optimise/fluxes_found.npy")
-zernikes_found = np.load(paths.data / "optimise/zernikes_found.npy")
+tel = deserialise(paths.data / f'make_model_and_data/{sub_dir}/instrument.zdx')
+model = deserialise(paths.data / f'optimise/{sub_dir}/final_model.zdx')
+errs = np.abs(np.diag(np.load(paths.data / f'calc_errors/{sub_dir}/cov_mat.npy')))**0.5
 
-positions = 'MultiPointSource.position'
-fluxes = 'MultiPointSource.flux'
-zernikes = 'ApplyBasisOPD.coefficients'
-flatfield = 'ApplyPixelResponse.pixel_response'
-parameters = [positions, fluxes, zernikes, flatfield]
-
-# Get the residuals
-coeff_residuals = tel.get(zernikes) - zernikes_found
-flux_residuals = tel.get(fluxes) - fluxes_found
-
-scaler = 1e3
-positions_residuals = tel.get(positions) - positions_found
-r_residuals_rads = np.hypot(positions_residuals[:, :, 0], positions_residuals[:, :, 1])
-r_residuals = r2a(r_residuals_rads)
-
-# Errors
+# Define Parameters
+positions  = 'Source.position'
+fluxes     = 'Source.flux'
[…]

### 2023-04-13 src/scripts/plot_data_resid.py

@@ -3,22 +3,29 @@ import paths
 import pickle as p
 import matplotlib.pyplot as plt
 from mpl_toolkits.axes_grid1 import make_axes_locatable
+from zodiax.experimental.serialisation import serialise, deserialise
+from observation import MultiImage
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
+flux = 1e8
+PRFdev = 1e-1
+sub_dir = f"flux_{flux:.0e}_PRFdev_{PRFdev:.0e}"
+
 # Load model
-data = np.load(paths.data / "make_model_and_data/data.npy")
-final_psfs = np.load(paths.data / "optimise/final_psfs.npy")
-initital_psfs = np.load(paths.data / "make_model_and_data/initial_psfs.npy")
+data = np.load(paths.data / f"make_model_and_data/{sub_dir}/data.npy").sum(0)
+final_psfs = np.load(paths.data / f"process/{sub_dir}/final_psfs.npy").sum(0)
+initital_psfs = np.load(paths.data / f"process/{sub_dir}/initial_psfs.npy").sum(0)
 
 plt.figure(figsize=(15, 4.5))
 plt.suptitle("Data and Residuals", size=15)
 
 ax1 = plt.subplot(1, 3, 1)
-plt.title("Sample Data")
+rms_data = np.sqrt((data**2).mean())
+plt.title(f"Data, Mean counts: {data.mean():.0f}")
 plt.imshow(data*1e-3)
 plt.xlabel("Pixels")
 plt.ylabel("Pixels")
@@ -28,7 +35,8 @@ cbar = plt.colorbar(cax=cax)
 cbar.set_label("Counts $x10^3$")
 
 ax2 = plt.subplot(1, 3, 2)
-plt.title("Initial Residual")
+initial_rms = np.sqrt(((initital_psfs - data)**2).mean())
+plt.title(f"Initial Residual: {initial_rms:.0f} counts RMS")
 plt.imshow((initital_psfs - data)*1e-3)
 plt.xlabel("Pixels")
 plt.ylabel("Pixels")
@@ -38,14 +46,15 @@ cbar = plt.colorbar(cax=cax)
 cbar.set_label("Counts $x10^3$")
 
 ax3 = plt.subplot(1, 3, 3)
[…]

### 2023-04-13 src/scripts/plot_divergence.py

@@ -1,51 +1,0 @@
-import jax.numpy as np
-from dLux.utils import radians_to_arcseconds as r2a
-import paths
-import pickle as p
-import matplotlib.pyplot as plt
-
-plt.rcParams['image.cmap'] = 'inferno'
-plt.rcParams["font.family"] = "serif"
-plt.rcParams["image.origin"] = 'lower'
-plt.rcParams['figure.dpi'] = 120
-
-# Load data
-fluxes_in = np.load(paths.data / "divergence/divergence_fluxes_in.npy")
-positions_found = np.load(paths.data / "divergence/positions_found.npy")
-fluxes_found = np.load(paths.data / "divergence/fluxes_found.npy")
-
-positions = 'PointSource.position'
-fluxes = 'PointSource.flux'
-zernikes = 'ApplyBasisOPD.coefficients'
-flatfield = 'ApplyPixelResponse.pixel_response'
-
-# Positions
-x_res, y_res = (np.zeros(2)- positions_found[-1]).T
-r_res = np.hypot(x_res, y_res)
-
-# Fluxes
-final_fluxes = fluxes_found[-1]
-res = np.abs(fluxes_in - final_fluxes[:, 0])
-l1 = res/fluxes_in
-xs = fluxes_in
-sigma = 1
-
-plt.figure(figsize=(10, 4))
-plt.subplot(1, 2, 1)
-plt.title("Positional Residual")
-plt.scatter(fluxes_in, r2a(r_res))
-plt.semilogx()
-plt.axhline(0, c='k')
-plt.xlabel("Flux")
-plt.ylabel("Positional Residual (arcseconds)")
-
-plt.subplot(1, 2, 2)
-plt.title("Fractional Flux Error $|(\hat{f} - f)/\hat{f}|$")
-plt.errorbar(xs, l1, yerr=(sigma*(xs**0.5)/xs), fmt='o', capsize=5)
-plt.axhline(0, c='k')
-plt.semilogx()
-plt.xlabel("Flux")
-plt.ylabel("Fractional Error")
-
-plt.tight_layout()
-plt.savefig(paths.figures / "divergence.pdf", dpi=300)

### 2023-04-13 src/scripts/plot_hist.py

@@ -56,12 +56,23 @@ for i in range(nhists):
     threshes.append(data_sort[size*i])
 threshes.append(data_sort[-1] + 1)
 
+threshes = np.linspace(0, data.max(), nhists+1).astype(int)
+
+
 data_flat = data.flatten()
 indexes = []
 for i in range(len(threshes)-1):
     low = np.where(data_flat >= threshes[i])[0]
     high = np.where(data_flat < threshes[i+1])[0]
     indexes.append(np.intersect1d(low, high))
+
+bin_size = []
+for i in range(len(indexes)):
+    # print(len(indexes[i]))
+    # print(len(indexes[i])/2/5)
+    bin_size.append(int(np.minimum(np.ceil(len(indexes[i])/25), 50)))
+print(bin_size)
+
 
 import matplotlib
 cmap = matplotlib.cm.get_cmap('inferno')
@@ -72,12 +83,15 @@ prf_flat = (pix_response - flatfields_found[-1]).flatten()
 for i in range(len(indexes)):
     lab = "{} - {}".format(int(threshes[i]), int(threshes[i+1]))
 
-    # mean = (threshes[i] + threshes[i+1])/2
-    mean = threshes[i+1]
+    mean = (threshes[i] + threshes[i+1])/2
+    # mean = threshes[i+1]
+    # mean = threshes[i]
     mean_norm = (mean - vmin)/(vmax - vmin)
     cols = cmap(mean_norm)[:3]
 
-    plt.hist(prf_flat[indexes[i]], bins=50, alpha=0.85, label=lab, color=cols)
+    plt.hist(prf_flat[indexes[i]], bins=bin_size[i], alpha=0.5, label=lab, color=cols, density=True)
+    # plt.hist(prf_flat[indexes[i]], bins=50, alpha=0.85, label=lab, color=cols, density=True)
+    # plt.hist(prf_flat[indexes[i]], bins=50, alpha=0.85, density=True)
 # cbar = plt.colorbar()
 plt.legend()
 plt.xlim(-0.2, 0.2)

### 2023-04-13 src/scripts/plot_noise.py

@@ -1,0 +1,129 @@
+import jax.numpy as np
+from dLux.utils import radians_to_arcseconds as r2a
+import paths
+import jax.scipy as jsp
+import matplotlib.pyplot as plt
+from zodiax.experimental.serialisation import serialise, deserialise
+from observation import MultiImage
+import dLux as dl
+from lib import weighted_pearson_corr_coeff
+
+plt.rcParams['image.cmap'] = 'inferno'
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["image.origin"] = 'lower'
+plt.rcParams['figure.dpi'] = 120
+
+fluxes = [1e6, 1e7, 1e8, 1e9, 1e10, 1e11, 1e12]
+PRFdevs = [0.0001, 0.001, 0.01, 0.1]
+
+flux_err, pos_err, zern_err, corrs = [], [], [], []
+for i in range(len(fluxes)):
+    flux = fluxes[i]
+    flux_err_, pos_err_, zern_err_, corrs_ = [], [], [], []
+    for j in range(len(PRFdevs)):
+        PRFdev = PRFdevs[j]
+        sub_dir = f"flux_{fluxes[i]:.0e}_PRFdev_{PRFdevs[j]:.0e}"
+
+        data = np.load(paths.data / f"make_model_and_data/{sub_dir}/data.npy")
+        tel = deserialise(paths.data / f'make_model_and_data/{sub_dir}/instrument.zdx')
+        model = deserialise(paths.data / f'optimise/{sub_dir}/final_model.zdx')
+
+
+        # PRF Correlations
+        flatfield = 'PRF.pixel_response'
+        corr = weighted_pearson_corr_coeff(
+            model.get(flatfield), tel.get(flatfield), data.sum(0))
+        corrs_.append(corr)
+
+
+        # Fluxes
+        true_fluxes = tel.Source.flux
+        found_fluxes = model.Source.flux
+        expected_err = np.mean(np.sqrt(true_fluxes))
+        MAE = np.mean(np.abs(true_fluxes - found_fluxes)/np.sqrt(true_fluxes))
+        flux_err_.append(MAE)
+
+
+        # Positions
+        true_pos = tel.Source.position
+        found_pos = model.Source.position
+        mean_wl = tel.Source.spectrum.wavelengths.mean()
+        lamd = mean_wl/tel.CreateWavefront.diameter
[…]

### 2023-04-13 src/scripts/plot_optics.py

@@ -4,31 +4,39 @@ import pickle as p
 import dLux as dl
 import matplotlib.pyplot as plt
 from matplotlib.cm import get_cmap
+from zodiax.experimental.serialisation import serialise, deserialise
+from observation import MultiImage
 
 plt.rcParams['image.cmap'] = 'inferno'
 plt.rcParams["font.family"] = "serif"
 plt.rcParams["image.origin"] = 'lower'
 plt.rcParams['figure.dpi'] = 120
 
+
+flux = 1e8
+PRFdev = 1e-1
+sub_dir = f"flux_{flux:.0e}_PRFdev_{PRFdev:.0e}"
+
 # Load model
-tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-wavels = np.load(paths.data / 'make_model_and_data/wavelengths.npy')
-psf = np.load(paths.data / "make_model_and_data/plain_psf.npy")
-aberrated_psf = np.load(paths.data / "make_model_and_data/aberrated_psf.npy")
+tel = deserialise(paths.data / f'make_model_and_data/{sub_dir}/instrument.zdx')
+unit_psf = np.load(paths.data / f"process/{sub_dir}/unit_psf.npy")
+aberrated_psf = np.load(paths.data / f"process/{sub_dir}/aberrated_psf.npy")
+counts = np.load(paths.data / f"process/{sub_dir}/PRF_counts.npy")
+bins = np.load(paths.data / f"process/{sub_dir}/PRF_bins.npy")
 
 # Pupil
-throughput = tel.CompoundAperture.get_aperture(npixels=tel.get('CreateWavefront.npixels'))
+throughput = tel.CircularAperture.aperture
 mask = tel.AddOPD.opd
 pupil = mask.at[np.where(throughput==0.)].set(np.nan)
 
 # Aberrations
-opd = tel.ApplyBasisOPD.get_total_opd()
+opd = tel.Aberrations.get_opd()
 aberrated_pupil = pupil + opd
 
 # FF
-FF = tel.ApplyPixelResponse.pixel_response
-
-cmap = get_cmap("inferno").copy()
+FF = tel.PRF.pixel_response
+from matplotlib import colormaps as cm
+cmap = cm["inferno"]
 cmap.set_bad('k',1.)
 
 plt.figure(figsize=(15, 8))
[…]

### 2023-04-13 src/scripts/plot_progress.py

@@ -1,107 +1,0 @@
-import jax.numpy as np
-from dLux.utils import radians_to_arcseconds as r2a
-import paths
-import dill as p
-import matplotlib.pyplot as plt
-
-plt.rcParams['image.cmap'] = 'inferno'
-plt.rcParams["font.family"] = "serif"
-plt.rcParams["image.origin"] = 'lower'
-plt.rcParams['figure.dpi'] = 120
-
-
-# Load model
-tel = p.load(open(paths.data / 'make_model_and_data/instrument.p', 'rb'))
-models_out = p.load(open(paths.data / 'optimise/models_out.p', 'rb'))
-losses = np.load(paths.data / 'optimise/losses.npy')
-data = np.load(paths.data / "make_model_and_data/data.npy")
-# psfs_out = np.load(paths.data / "optimise/final_psfs.npy")
-
-positions = 'MultiPointSource.position'
-fluxes = 'MultiPointSource.flux'
-zernikes = 'ApplyBasisOPD.coefficients'
-flatfield = 'ApplyPixelResponse.pixel_response'
-parameters = [positions, fluxes, zernikes, flatfield]
-
-# Get parameters
-# positions_found  = np.array([model.get(positions) for model in models_out])
-# fluxes_found     = np.array([model.get(fluxes)    for model in models_out])
-# zernikes_found   = np.array([model.get(zernikes)  for model in models_out])
-# flatfields_found = np.array([model.get(flatfield) for model in models_out])
-positions_found = np.load(paths.data / "optimise/positions_found.npy")
-fluxes_found = np.load(paths.data / "optimise/fluxes_found.npy")
-zernikes_found = np.load(paths.data / "optimise/zernikes_found.npy")
-flatfields_found = np.load(paths.data / "optimise/flatfields_found.npy")
-
-# Get the residuals
-coeff_residuals     = tel.get(zernikes)  - zernikes_found
-flux_residuals      = tel.get(fluxes)    - fluxes_found
-flatfield_residuals = tel.get(flatfield) - flatfields_found
-
-scaler = 1e3
-positions_residuals = tel.get(positions) - positions_found
[…]

### 2023-04-13 src/scripts/run.py

@@ -1,22 +1,0 @@
-import os
-import time as t
-
-t0 = t.time()
-print("\nRunning make_model_and_data.py")
-os.system('python make_model_and_data.py')
-print("Done in {:.2f} seconds".format(t.time() - t0))
-
-t0 = t.time()
-print("\nRunning optimise.py")
-os.system('python optimise.py')
-print("Done in {:.2f} seconds".format(t.time() - t0))
-
-t0 = t.time()
-print("\nRunning calc_errors.py")
-os.system('python calc_errors.py')
-print("Done in {:.2f} seconds".format(t.time() - t0))
-
-t0 = t.time()
-print("\nRunning divergence.py")
-os.system('python divergence.py')
-print("Done in {:.2f} seconds".format(t.time() - t0))

### 2023-04-13 src/scripts/run_full.py

@@ -1,0 +1,44 @@
+import sys
+import os
+import time as t
+import paths
+from lib import mkdir, format_time, run_script
+from lib import make_model_and_data, optimise, process, calc_errors
+
+fluxes = [1e6, 1e7, 1e8, 1e9, 1e10, 1e11, 1e12]
+PRFdevs = [0.0001, 0.001, 0.01, 0.1]
+
+# fluxes = [1e8]
+# PRFdevs = [0.1]
+
+for i in range(len(fluxes)):
+    flux = fluxes[i]
+    for j in range(len(PRFdevs)):
+        PRFdev = PRFdevs[j]
+        sub_dir = f"flux_{fluxes[i]:.0e}_PRFdev_{PRFdevs[j]:.0e}"
+
+        # print()
+        print(f"Flux, PRFdev: {flux:.0e}, {PRFdev:.1}")
+
+        # make_model_and_data(flux, PRFdev)
+        # optimise(flux, PRFdev)
+        # process(flux, PRFdev)
+        # calc_errors(flux, PRFdev)
+
+        # # Calculation
+        # run_script('make_model_and_data', sub_dir, True)
+        # # run_script('optimise',            sub_dir, True)
+        # run_script('optimise2',           sub_dir, True)
+        # run_script('process',             sub_dir, True)
+        # run_script('calc_errors',         sub_dir, True)
+
+        # # Plotting
+        # run_script('plot_progress',     sub_dir)
+        # run_script('plot_astro_params', sub_dir)
+        # run_script('plot_aberrations',  sub_dir)
+        # run_script('plot_data_resid',   sub_dir)
+        # run_script('plot_optics',       sub_dir)
+        # run_script('plot_FF',           sub_dir)
+        
+        # mkdir(paths.figures / f"summary/Flux_{fluxes[i]:.0e}")
+        # run_script('plot_summary', sub_dir)

### 2023-04-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand\equinox{\texttt{equinox}\xspace}
+\newcommand\zodiax{\texttt{zodiax}\xspace}
 \newcommand\optax{\texttt{optax}\xspace}
@@
 % we introduce a new package dLux for differentiable, gpu-accelerated physical optics simulation
-We introduce a new Python package for physical optics simulation, \dlux, which uses the machine learning framework \jax to achieve GPU acceleration and automatic differentiation, and apply this to simulating astronomical imaging.
+We introduce a new Python package for physical optics simulation, \dlux, which uses the machine learning framework \jax to achieve GPU acceleration and automatic differentiation ('autodiff'), and apply this to simulating astronomical imaging.
 % we show we can do flat field and phase retrieval simultaneously
@@
 It significantly extends the capabilities of \texttt{morphine} \cite{pope2021,phase_ret_and_design}, a previous \jax optics package from our team based on \texttt{poppy}, being rewritten from the ground up taking full advantage of more-recent \jax features and libraries.
-It is built in an object-oriented \jax framework called \equinox \cite{kidger2021equinox}, with an optical system consisting of an \equinox module wrapping a stack of `layers' which are themselves \equinox modules applying phase or amplitude screens, or performing Fourier or Fresnel transforms on a wavefront.
+It is built in an object-oriented \jax framework called \zodiax which extends \equinox \cite{kidger2021equinox} for scientific programming.
+An optical system consisting of an \zodiax module wrapping a stack of `layers' which are themselves \zodiax modules applying phase or amplitude screens, or performing Fourier or Fresnel transforms on a wavefront.
 This allows us to re-frame the idea of a computational optical model by analogy to a `parametrized neural network', with each operation performed on the wavefront analogous to a neural network layer encoding the physics of the transformation.
-The object-oriented nature of \equinox allows for larger and more complex models to be build and seamlessly integrated than would be possible within the purely functional framework default to \jax.
-\equinox is also built to be directly integrated into the \jax framework, allowing for all \jax function to seamlessly interface with any \equinox models and vice-versa.
+The object-oriented nature of \zodiax allows for larger and more complex models to be built and seamlessly integrated than would be possible within the purely functional framework default to \jax.
+\zodiax is also built to be directly integrated into the \jax framework, allowing for all \jax functions to seamlessly interface with any \zodiax models and vice-versa.
 Talk about zodiax too?
@@
 To recover these parameters we initialise a naive unaberrated optical model with a uniform \ac{prf}.
-Stellar positions were perturbed by adding a random value drawn from a normal distribution mean 0 deviation 1 pixel and fluxes were perturbed by multiplying their value by a random value value drawn from a normal distribution mean 1 deviation 0.1.
+Stellar positions were perturbed by adding a random value drawn from a normal distribution mean 0 deviation 1 pixel and fluxes were perturbed by multiplying their value by a random value drawn from a normal distribution mean 1 deviation 0.1.
 We will use this model to recover the true values.
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{figs/noise_performance.pdf}
+    \includegraphics[width=1.0\textwidth]{figures/noise_performance.pdf}
     % \includegraphics[width=0.8\textwidth]{figures/noise_performance.pdf}
@@
 These parameters are simple enough to tune for individual data sets with similar scales, but it is non-trivial to automate this process for data with a vast range of scales like those being analysed here.
-The hyper-parameters chosen were tuned on the 1e8 flux data set and roughly scaled to the work with the other data sets, which is why we observe the best recovery for that set of parameters.
+The hyper-parameters chosen were tuned on the 1e8 flux data set and roughly scaled to work with the other data sets, which is why we observe the best recovery for that set of parameters.
 These figures are designed to demonstrate the overall trends in the different noise regimes, rather than to be a precise analysis of the fundamental limits.

### 2023-04-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Correlations and residual plots of the positions, fluxes and aberrations after the optimisation are shown in Figure~\ref{fig:astro_params} and Figure~\ref{fig:aberrations}.
-Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of
-% \textcolor{red}{showyourwork variable}
-% \variable{output/rms_opd_resid.txt} 
-0.0499\,nm RMS. \textcolor{red}{update with final values}
+Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of 0.0282\,nm RMS.
+
 Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram.

### 2023-04-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We make \dlux available as open-source software on GitHub\footnote{\href{https://github.com/LouisDesdoigts/dLux}{github.com/LouisDesdoigts/dLux}}, and encourage interested readers to use and contribute to this package as a community resource.
-This paper has been compiled using ShowYourWork \cite{Luger2021}, so that all figures link to the code used to produce them.
+This paper has been compiled using ShowYourWork \cite{Luger2021}, so that all figures link to the code used to produce themcan be found \href{https://github.com/LouisDesdoigts/instrumental_calibration/tree/main}{here}.
 
@@
 
-This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm.
+This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545\,nm, 595\,nm and 645\,nm.
 This total model requires the calculation of 60 individual \ac{psf}s over a $1024\times1024$ detector from a $512\times512$ pixel wavefront.
-This calculation is vastly dominated by the propagation of the wavefront using a 2-sided MFT, which requires two matrix multiplications for each individual \ac{psf} of sizes $[N_0 \times N_i] \cdot [N_i \times N_i]$ and $[N_0 \times N_i] \cdot [N_i \times N_o]$ where $N_i = 512$ is the input size of the wavefront and $N_o = 1024$ is the output size of the wavefront.
+This calculation is vastly dominated by the propagation of the wavefront using a 2-sided MFT, which requires two matrix multiplications for each individual \ac{psf} of sizes $[N_o \times N_i] \cdot [N_i \times N_i]$ and $[N_o \times N_i] \cdot [N_i \times N_o]$ where $N_i = 512$ is the input size of the wavefront and $N_o = 1024$ is the output size of the wavefront.
 This takes a total of $\sim$ 1.25 seconds to evaluate on an Apple M1 CPU, with all operations performed using FP32 precision default to \jax.

### 2023-04-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We make \dlux available as open-source software on GitHub\footnote{\href{https://github.com/LouisDesdoigts/dLux}{github.com/LouisDesdoigts/dLux}}, and encourage interested readers to use and contribute to this package as a community resource.
-This paper has been compiled using ShowYourWork \cite{Luger2021}, so that all figures link to the code used to produce themcan be found \href{https://github.com/LouisDesdoigts/instrumental_calibration/tree/main}{here}.
+This paper has been compiled using ShowYourWork \cite{Luger2021}, so that all figures link to the code used to produce them can be found \href{https://github.com/LouisDesdoigts/instrumental_calibration/tree/main}{here}.
 
